### 2015-03-19 paper_prni/prni.tex
first commit

%\documentclass[runningheads,a4paper]{llncs}
\documentclass[10pt, conference, compsocconf]{IEEEtran}
\usepackage{amssymb}
%\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{comment}
\usepackage{hyperref}
\usepackage{subfigure}
\usepackage{url}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{xcolor,colortbl}
\usepackage{float}
\usepackage{capt-of}
\usepackage{natbib}

%\newcommand{\keywords}[1]{\par\addvspace\baselineskip
%\noindent\keywordname\enspace\ignorespaces#1}

\def\bX{\mathbf{X}}
\def\bS{\mathbf{S}}
\def\bN{\mathbf{N}}
\def\bU{\mathbf{U}}
\def\cX{\mathcal{X}}  
\def\cG{\mathcal{G}}
\def\cT{\mathcal{T}}
\def\bx{\mathbf{x}}
\def\bu{\mathbf{u}}
\def\bs{\mathbf{s}}
\def\bn{\mathbf{n}}

\hyphenation{}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\definecolor{Gray}{gray}{0.92}
\newcolumntype{u}{>{\columncolor{Gray}}c}

\begin{document}
%
\title{Low-rank logistic regression for supervised network decomposition}
%
\author{
\IEEEauthorblockN{
Danilo Bzdok\IEEEauthorrefmark{1}
Olivier Grisel\IEEEauthorrefmark{1},
Michael Eickenberg\IEEEauthorrefmark{1},
Ga\"el Varoquaux\IEEEauthorrefmark{1},
Bertrand Thirion\IEEEauthorrefmark{1},
}
\IEEEauthorblockA{\IEEEauthorrefmark{1}INRIA Parietal,
Neurospin, b\^at 145, CEA Saclay, 91191 Gif sur Yvette, France\\
firstname.lastname@inria.fr
}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Corresponding author}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
Neuroimaging methods tend to focus on either structure discovery
or cognitive modulation. Predicting cognitive processes however hinges
on the neurobiological pertinence of the constructed feature space.
%
We therefore propose to the solve the unsupervised dimensionality reduction
and supervised task classification in
an identical optimization problem. Using
Human Connectome Project (HCP) data (n=498), optimal low-rank projections and
logistic-regression models are identified in a same gradient descent. Brain
network decompositions are thus exposed that explain task-descriminative spatial
patterns. This is compared against separate decomposition (i.e., ICA and
sparse PCA) and classification by logistic regression in independent data splits.
%
RESULTSoutperformes
%
We thus provide modes of brain signal variation extracted for
cognitive interpretability.
%
\end{abstract}

\begin{IEEEkeywords}
Dimensional reduction; brain networks; machine learning; brain imaging
\end{IEEEkeywords}

\section{Introduction}
%
Neuroimaging methods can be grouped by focus on discovering
neurobiological structure or discovering the neural correlates associated
with cognitive processes.
To reveal coherent spatial structures across time,
independent component analysis (ICA; \citep{beckmann2005}) is frequently used
to decompose BOLD signals into a
set of fluctuation patterns. 
The ensuing spatial patterns are believed to represent neural networks of
functionally interaction brain regions.
Similarly, sparse PCA (SPCA; \citep{varoqu2011}) has recently been used to
separate BOLD signals into parsimonious
network components with less distributed regions.
Rather than partially overlapping spatial patterns, most clustering
approaches used in neuroimaging (e.g., k-means, hierarchical, or ward clustering;
\citep{thirion14}) yield non-overlapping regions of coherent connectivity
or task activation.
Clustering of brain regions or the entire brain is increasingly used to 
locate functionally distinct neurobiological compartments \citep{behrens03}.
As another example, graph analyses \citep{bullmore09} use predefined
regions to study high-level network properties.
All these approaches are completely \textit{unsupervised} in
characterizing neurobiological structure without any relation to
psychological processes.
\linebreak

To investigate, however, the neural correlates underlying mental operations,
the general linear model (GLM; \citep{friston94}) is the prevalent approach.
The contribution of
individual voxels is estimated according to a design matrix of experimental
conditions. Psychophysiological interactions (PPI; \citep{friston97}),
on the other hand,
elucidate the functional interactions between brain regions as a function
of experimental conditions. As a final example, 
dynamic causal modeling (DCM; \citep{stephan04}) quantifies directed,
cognitive-task-driven influences between regions
by treating the brain as a nonlinear dynamic system with unknown
neuronal states. This second set of \textit{supervised} approaches
operates in single voxels or
investigator-provided region definitions to localize the neural underpinnings of
psychological processes.
\linebreak

Yet, special interest lies in the data-driven
discovery of neurobiological structure that explains differences
in cognitive conditions. Integrating unsupervised discovery and
supervised classification should preferentially identify
neurobiological structure that allows for the best predictive models. We
propose such an approach based on conjoint gradient-descent to optimize
linear decomposition and multi-task disambiguation.

\begin{wrapfigure}
  \includegraphics[width=0.48\textwidth]{fig001}
  \caption {\textbf{Integrated versus serial decomposition and classification}\\
  Depicts the out-of-sample performance of 18-task classification based on
	HCP task activity maps. Data reduction and classification
	was performed simultaneously (Low-rank logistic regression, red) or serially
	(ICA and SparsePCA, blue)}
\end{wrapfigure}

%
\section{Methods}
%

Data was drawn from 498 unrelated, healthy HCP participants.
All provided informed consent to the Washington
University in St. Louis institutional review board. HCP tasks 
were selected that feature known suitability as localizers
and reliability
across participants \citep{barch2013}.
Mostly block-design, but also event-related, paradigms were
administered on 1) working memory/cognitive control processing, 2)
incentive processing, 3) visual and somatosensory-motor processing,
4) language processing (semantic and phonological processing),
5) social cognition, 6) relational processing, and 7) emotional
processing. All data were acquired on the same Siemens Skyra 3T scanner.
Whole-brain EPI acquisitions were acquired with a
32 channel head coil (TR=720ms, TE=33.1ms, flip angle=52°, BW=2290Hz/Px,
in-plane FOV=$280\times180$mm, 72 slices, 2.0mm isotropic voxels).
The “minimally preprocessed” pipeline \citep{glass13} includes
gradient unwarping, motion correction, fieldmap-based EPI distortion
correction, brain-boundary-based registration of EPI to structural
T1-weighted scan, non-linear (FNIRT) registration into MNI space,
and grand-mean intensity normalization. Activation maps were spatially
smoothed by a Gaussian kernel of 4mm (FWHM). A general linear model (GLM) was
implemented by FILM from the FSL suite with model regressors from convolution
with a “canonical” hemodynamic response function and from temporal derivatives.
HCP tasks were conceived to modulate activation
in a maximum of different brain regions and neural systems. Indeed, at
least 70\% of the participants showed consistent brain activity in
contrasts from the task battery, which certifies excellent
coverage \citep{barch2013}.
In sum, the HCP task dataset incorporated 8650 first-level activity maps
from 18 diverse paradigms administered to 498 participants.
All maps were downsampled to a common 36x43x36 space of
5mm isotropic voxels and gray-matter masked (at least 10\%).
All analyses were based on task maps of
13,657 voxels representing Z values in gray matter.
\linebreak

Unsupervised and supervised learning were combined into a low-rank logistic
regression problem. The 13,657 z values from each activity map were
subject to a first linear
projection into Z latent components (i.e., 1, 5, 10, .., 100).
These hidden brain networks
loadings were subsequently projected into the 18 class space for multinomial
logistic regression. The goal was to find the two weight matrices
(input-to-hidden: 13,657 x Z and hidden-to-output: Z x 18) and their
corresponding bias vectors. Non-linearities were not applied on the
transformation results. Weights and biases were initialized by Gaussian
random noise. Gradient descent updated these matrices and vectors in each
iteration (100 samples per batch, 250 epochs). Using the chain rule, the
partial
derivates for the update were computed for the transformation into the
latent space and the subsequent transformation into the class space. We choose
the RMSprop algorithm \citep{dauphin15} with an initial learning rate of
0.01. Early stopping ensured that the best weight matrices/vectors were
retained in each epoch cycle. This was evaluated by the prediction
accuracy on a validation set (10\% of the training data) at each iteration.
\linebreak

This approach was benchmarked against independent dimensionality reduction and
learning of a classification function. Data reduction was performed on one
half of the data by ICA and SPCA.
ICA unmixed the
BOLD signals into separate spatial components
by minimizing their mutual information \citep{hyvarinen2000}.
This iterative blind source separation was
realized by a parallel FASTICA implementation (200 maximum iterations,
per-iteration tolerance of 0.0001,
initialized by a random mixing matrix, preliminary whitening).
SPCA separated the BOLD signals into
network components with few regions, which scales well to 
large datasets \citep{varoqu2011}.
This regression-type optimization problem constrained by
$\ell_1$-penalty in an implementation
without orthogonality assumptions
(1000 maximum iterations, per-iteration tolerance of 1 * 10\textsuperscript{-8}, sparsity
alpha=1, ridge-shrinkage at 0.01, Lasso path computed with coordinate
descent). Each linear decomposition revealed the specified number of
latent network components in one half the HCP data.
\linebreak
The extracted hidden network components were subsequently used
to reduce the remain half of task maps to a considerably smaller number of component
loadings. 13,657 voxels were thus condensed into 1, 5, 10, .., 100
measures of network involvement using ridge regression
(regularization alpha parameter=0.001, using
Cholesky solver). Multinomial logistic regression was finally performed on the
brain network loadings to learn classifying the 18 cognitive tasks.
Importantly, in all approaches, the final classification models were tested
on an unseen test set (10\% of data).
\linebreak

All analysis scripts that produced the results are accessible online
(\url{http://github.com/banilo/prni2015}).

\begin{figure*}
  \includegraphics[width=0.98\textwidth]{fig002}
  \caption {\textbf{Decomposition into 5 networks}}
\end{figure*}

\section{Results}
Low-rank regression outperformed serial ICA/SparsePCA and logistic regression.

\section{Discussion}
%

-hypothesis space includes sparse PCA and PCA but not ICA since no
linearity

- if linearity, then would be closer to the notion of 1-hidden layer neural
network rather than low-rank logistic regression

%
\paragraph{Acknowledgment}
{\small
Data were provided the Human Connectome Project.
}

%
% ---- Bibliography ----
%
\bibliographystyle{IEEEtran}
\bibliography{prni_refs}
%
\end{document}

### 2015-03-19 dnn.py

@@ -1,0 +1,865 @@
+"""
+A deep neural network with or w/o dropout in one file.
+"""
+
+import numpy
+import theano
+import sys
+import math
+from theano import tensor as T
+from theano import shared
+from theano.tensor.shared_randomstreams import RandomStreams
+from collections import OrderedDict
+
+BATCH_SIZE = 100  # default batch size
+L2_LAMBDA = 1.    # default L2 regularization parameter
+INIT_LR = 0.01    # initial learning rate, try making it larger
+
+def relu_f(vec):
+    """ Wrapper to quickly change the rectified linear unit function """
+    return (vec + abs(vec)) / 2.
+
+
+def softplus_f(v):
+    return T.log(1 + T.exp(v))
+
+
+def dropout(rng, x, p=0.5):
+    """ Zero-out random values in x with probability p using rng """
+    if p > 0. and p < 1.:
+        seed = rng.randint(2 ** 30)
+        srng = theano.tensor.shared_randomstreams.RandomStreams(seed)
+        mask = srng.binomial(n=1, p=1.-p, size=x.shape,
+                dtype=theano.config.floatX)
+        return x * mask
+    return x
+
+
+def build_shared_zeros(shape, name):
+    """ Builds a theano shared variable filled with a zeros numpy array """
+    return shared(value=numpy.zeros(shape, dtype=theano.config.floatX),
+            name=name, borrow=True)
+
+
+class Linear(object):
+    """ Basic linear transformation layer (W.X + b) """
+    def __init__(self, rng, input, n_in, n_out, W=None, b=None):
+        if W is None:
+            W_values = numpy.asarray(rng.uniform(
+                low=-numpy.sqrt(6. / (n_in + n_out)),
+                high=numpy.sqrt(6. / (n_in + n_out)),
+                size=(n_in, n_out)), dtype=theano.config.floatX)
+            W_values *= 4  # This works for sigmoid activated networks!
+            W = theano.shared(value=W_values, name='W', borrow=True)
+        else:
+            print('W already set !!!')
+        if b is None:
+            b = build_shared_zeros((n_out,), 'b')
+        self.input = input
+        self.W = W
+        self.b = b
+        self.params = [self.W, self.b]
+        self.output = T.dot(self.input, self.W) + self.b
+
+    def __repr__(self):
+        return "Linear"
+
+
+# class PretrainedLinear(object):
+#     """ Basic linear transformation layer (W.X + b) """
[…]

### 2015-03-19 prni.py

@@ -1,0 +1,540 @@
+
+"""
+HCP: supervised network decomposition by articial neural networks
+
+Danilo Bzdok, 2015
+danilobzdok@gmail.com
+"""
+
+print __doc__
+
+import os
+import os.path as op
+import numpy as np
+from scipy.linalg import norm
+import nibabel as nib
+from sklearn.grid_search import RandomizedSearchCV
+from sklearn.base import BaseEstimator
+from nilearn.input_data import NiftiMasker
+import theano
+import theano.tensor as T
+print('Running THEANO on %s' % theano.config.device)
+from nilearn.image import concat_imgs
+import joblib
+
+import dnn
+
+WRITE_DIR = op.join(os.getcwd(), 'prni2015_2')
+if not op.exists(WRITE_DIR):
+    os.mkdir(WRITE_DIR)
+
+##############################################################################
+# load+preprocess data
+##############################################################################
+
+mask_img = nib.load('grey10_icbm_5mm_bin.nii.gz')
+# mask_img = nib.load('grey10_icbm.nii')
+nifti_masker = NiftiMasker(mask_img=mask_img, smoothing_fwhm=False,
+                           standardize=False)
+nifti_masker.fit()
+mask_nvox = nifti_masker.mask_img_.get_data().sum()
+
+print('Loading data...')
+X, labels, subs = joblib.load('/git/dl_nets/preload_HT')
+
+# X1 = joblib.load('/git/cohort/archi/preload_1st_HT')
+# labels1 = joblib.load('/git/cohort/archi/preload_1st_HT_labels')
+# subs1 = joblib.load('/git/cohort/archi/preload_1st_HT_subs')
+# X2, labels2, subs2 = joblib.load('/git/cohort/archi/preload_2nd_HT')
+# 
+# X = concat_imgs([X1, X2], accept_4d=True)
+# del X1; del X2
+# X = nifti_masker.transform(X)
+# labels = list(labels1) + list(labels2)
+# subs = list(subs1) + list(subs2)
+
+print('done :)')
+
+sub_ids = np.unique(subs)
+train_inds = np.in1d(subs, sub_ids[:450])
+test_inds = np.in1d(subs, sub_ids[450:])
+# sanity-check separation of subjects
+assert(not np.any(np.in1d(subs[train_inds], sub_ids[450:])))
+assert(not np.any(np.in1d(subs[test_inds], sub_ids[:450])))
[…]

### 2015-05-14 nips.py
interm

@@ -1,8 +1,16 @@
 """
 HCP: semi-supervised network decomposition by low-rank logistic regression
+"""
+"""
+preliminary conclusion:
+- the combined AE and LR loss yields best classification results
+- Regarding rest data, 20-comp PCA work better than uncompressed data
 
-Danilo Bzdok, 2015
-danilobzdok@gmail.com
+questions:
+- the AE cost is much bigger, is it ok to T.log() that?
+No because it renders the objective function non-convex.
+- how can we compute task-typical whole-brain images onces we have learned
+the supervised problem?
 """
 
 print __doc__
@@ -10,11 +18,13 @@ print __doc__
 import os
 import os.path as op
 import numpy as np
+import glob
 from scipy.linalg import norm
 import nibabel as nib
 from sklearn.grid_search import RandomizedSearchCV
 from sklearn.base import BaseEstimator
 from sklearn.preprocessing import StandardScaler
+from sklearn.metrics import precision_recall_fscore_support
 from nilearn.input_data import NiftiMasker
 import theano
 import theano.tensor as T
@@ -39,7 +49,8 @@ mask_nvox = nifti_masker.mask_img_.get_data().sum()
 
 print('Loading data...')
 
-X_task, labels, subs = joblib.load('/git/dl_nets/preload_HT_10mm_ero2')
+# ARCHI task
+X_task, labels, subs = joblib.load('preload_HT_10mm_ero2')
 
 labels = np.int32(labels)
 
@@ -50,24 +61,42 @@ X_task = X_task[new_inds]
 labels = labels[new_inds]
 subs = subs[new_inds]
 
-X_rest = nifti_masker.transform(
-    '/git/cohort/archi/preload_HR20persub_10mm_ero2.nii')
-    
+# rest
+# X_rest = nifti_masker.transform('preload_HR20persub_10mm_ero2.nii')
+X_rest = nifti_masker.transform('preload_HRpca20_10mm.nii')
+
 X_task = StandardScaler().fit_transform(X_task)
 X_rest = StandardScaler().fit_transform(X_rest)
 
+# HACK!
+X_rest = np.vstack((X_rest, X_rest))
 print('done :)')
 
+# ARCHI task
+AT_niis, AT_labels, AT_subs = joblib.load('preload_AT_10_ero2')
+AT_X = nifti_masker.transform(AT_niis)
+AT_X = StandardScaler().fit_transform(AT_X)
+
 ##############################################################################
-# compute results
+# define computation graph
 ##############################################################################
 
[…]

### 2015-05-15 paper_nips/paper.tex
news

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{bbm}
\usepackage{algorithm,algorithmic}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{bbm}
\usepackage[titletoc]{appendix}

\def\B#1{\bm{#1}}
%\def\B#1{\mathbf{#1}}
\def\trans{\mathsf{T}}

%\renewcommand{\labelitemi}{--}

\newtheorem{theorem}{Theorem} \newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}{Remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Semi-supervised low-rank logistic regression for
high-dimensional neuroimaging datasets}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\DeclareMathOperator{\proj}{proj}
\DeclareMathOperator{\soft}{soft}
\DeclareMathOperator{\prox}{prox}
\DeclareMathOperator{\Prox}{Prox}
\DeclareMathOperator{\im}{im}

% macros from michael's .tex
\DeclareMathOperator{\dist}{dist} % The distance.
\DeclareMathOperator{\argmin}{argmin}
\DeclareMathOperator{\argmax}{argmax}
\DeclareMathOperator{\Id}{Id}
\DeclareMathOperator{\abs}{abs}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}

%\nipsfinalcopy % Uncomment for camera-ready version
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\author{Danilo Bzdok, Michael Eickenberg, Bertrand Thirion, Ga\"el
  Varoquaux \\\\\textbf{\textit{email:} }firstname.lastname@inria.fr}

\begin{abstract}
Imaging neuroscience links human behavior to aspects of brain
biology in ever-increasing datasets.
%
Existing neuroimaging methods typically perform either discovery of
neurobiological structure or evaluate targeted hypotheses on mental tasks.
%
Modelling mental operations however hinges
on the pertinence of the assumed neurobiological structure.
%
We therefore propose to the solve the unsupervised dimensionality reduction
and supervised task classification in
an identical statistical learning problem.
%
We show that this approach yields more accurate and more interpretable
neural models of psychological tasks in a reference neuroimaging dataset.
%

\textbf{keywords}: dimensionality reduction; semi-supervised learning
bioinformatics; fMRI; systems neuroscience

\end{abstract}

\section{Introduction}
%
Methods used in neuroimaging research can be grouped by discovering
neurobiological structure or revealing the neural correlates associated
with mental tasks.
To discover coherent distributions of activation structure across time,
independent component analysis (ICA; \cite{beckmann2005}) is often used
to decompose the BOLD (blood-oxygen level-dependent) signals into the
important modes of variation.
The ensuing spatial activation patterns are believed to represent
brain networks of
functionally interaction brain regions.
Similarly, sparse PCA (SPCA; \cite{varoqu2011}) has been used to
separate brain activity signals into parsimonious network components.
Thus extracted brain networks have been shown to be
manifestations of electrophysiological oscillation structure \cite{hipp15}.
Their fundamental role is
attested by continued covariation during sleep and anesthesia.
%
Network discovery is typically performed by applying ICA or SPCA on
unlabeled "resting-state" data. These capture brain dynamics
during ongoing random thought without controlled environmental input.
The rationale is that the biggest fraction of BOLD signals do
not correlate with a particular behavior, stimulus, or experimental task.

\linebreak

On the other hand, to investigate
the neural correlates underlying mental tasks,
the general linear model (GLM; \cite{friston94}) is the dominant approach.
The contribution of
individual brain voxels is estimated
according to a design matrix of experimental tasks.
Alternatively, psychophysiological interactions (PPI; \cite{friston97}),
elucidate the functional interactions between voxels as a function
of experimental tasks.
Dynamic causal modeling (DCM; \cite{stephan04}), in turn, quantifies directed,
task-driven influences between regions
by treating the brain as a nonlinear dynamic system with unknown
neuronal states. As a last example, always more neuroimaging studies model
experimental tasks by training classification algorithms on brain signals
\cite{poldrack09decoding}.
All these methods are applied to labeled task data that capture brain dynamics
during stimulus-guided behavior.
Two important conclusions can be drawn.
First, the mentioned supervised neuroimaging analyses operate
without exception in a single-voxel space. This ignores the fact that the BOLD
signal contains coherent spatial activation patterns.
Second, existing neuroimaging analyses do not acknowledge the fact that the
task-induced changes of the BOLD signal amount to less than 5\%
of baseline activity \cite{fox07}. They do thus not exploit the high similarity
of BOLD dynamics of the human brain during experimental tasks and at rest.
Indeed, similar brain networks were observed when applying ICA on
large quantities of
rest and task data \cite{smith2009}.

\linebreak

Both aspects can be combined into a 
semi-supervised (i.e., use task and rest data)
low-rank (i.e., perform network decomposition)
approach for neuroimaging analyses.

Yet, special interest lies in the data-driven
discovery of neurobiological structure that explains differences
in cognitive conditions. Integrating unsupervised discovery and
supervised classification should preferentially identify
neurobiological structure that allows for the best predictive models. We
propose such an approach based on conjoint gradient-descent to optimize
linear decomposition and multi-task disambiguation.

Find the most discriminative linear combination of the main direction of
variation in the data

modes of variation in the brain and the neural correlates of mental
operations have mostly been studied separately so far
their correspondence remains largely unclear

we provide principled computational framework to link these previously
unconnected domains of study.

learn a manifold / low-dimensional coordinate system
motivates

There is much uncertainty about the most pertinent representation
of neural activation information

PCA, SPCA and ICA are produce compact representations of uncorrelated or
independent components

combine exploratory and confirmatory analysis steps in a single framework
allows for  overlapping response pattersn

adjust the weight assigned to each spatial pattern to maximize
discriminatory power

useful intermediate representation

integrate recently gained domain knowledge into statistical learning
problems.

Traditionally, dimensionality reduction depended on linear methods such as PCA
Autoencoder \cite{hinton06}:
one-layered learning model that computes distributed
representation from the inputs by improving reconstructions from them.
dimensionality reduction by learning the underlying manifolds
for efficient representation of the data
bottleneck
-> use AE als feature extractions
guide the estimator where to put probability mass in the inputs
AE is a common feature learning algorithm
if hidden layer linear and squared error loss, it behaves like a PCA \cite{baldi1989neural}
convex with unique global minimum
optimization objective is the same as for PCA

Most nonsupervised learning algorithms can be implemented as variants in the 
autoencoder framework. This includes decmoposition and clustering task

Linear autoencoders with tied weights and specific regularization terms
yield efficient algorithms for ICA / nonlinearity -> then closer to feed-forward neural network
\cite{le2011ica}

ICA, Sparse autoencoders, and sparse coding are mathematically equivalent
under mild conditions \cite{le2011ica}

AE can caputer global and local distribution

and SPCA (??).
-> an autoencoder can capture various different compression procedures

- reformulate the AE prbleme as a regularization boejctive

-> backpropagation: solve the unsupervised ata repsentation and supervised
problem in an identical process.
coompute the error derivates of the weights
-> gradually modified according to possible compression algorithms

without a-priori knowledge of the structure
affine transformation to compression information in the input
class of problems
architecture
statistical models

directions of greatest variance; represents each data point by its coordinates along each of these directions

LR:
these weights are then fine-tuned by the LR

Even if linear should scan a wide space of intersring model, since
theoretcially, a two-layered neural networks can learn almost any function.

learn to the optimal projection of the input data
into the main direction of variation and
then classify along these principal directions.
Vincent2010: it has been shown experimen- tally that
beginning by optimizing an unsupervised criterion, oblivious of the specific
classification problem, can actually greatly help in eventually achieving
superior performance for that classifica- tion problem.

among the models it will favor those that feature most neurobiological
pertinence

From the perspective of dictionary learning, the first layer can be
viewed as a set of basis functions whose linear combinations are learned
in the second layer \cite{olshausen96}.

project the data-generating process on the axes of main variation
autoencoder allows to reverse-engineer proerties of the data-generating
neural process \cite{olshausen96}.

learn complex properties of the brain's task response patterns

convert high-dimensional data to low-dimensional network codes

data should be more easily separable in the network spacew
low-dimensional, linear structure

equates with a shallow neural network without non-linearities

%
\section{Methods}
%
Using
Human Connectome Project (HCP) data (n=500), optimal low-rank projections and
logistic-regression models are identified in a same gradient descent. Brain
network decompositions are thus exposed that explain task-descriminative spatial
patterns.

Data was drawn from 498 unrelated, healthy HCP participants.
All provided informed consent to the Washington
University in St. Louis institutional review board. HCP tasks 
were selected that feature known suitability as localizers
and reliability
across participants \cite{barch2013}.
Mostly block-design, but also event-related, paradigms were
administered on 1) working memory/cognitive control processing, 2)
incentive processing, 3) visual and somatosensory-motor processing,
4) language processing (semantic and phonological processing),
5) social cognition, 6) relational processing, and 7) emotional
processing. All data were acquired on the same Siemens Skyra 3T scanner.
Whole-brain EPI acquisitions were acquired with a
32 channel head coil (TR=720ms, TE=33.1ms, flip angle=52°, BW=2290Hz/Px,
in-plane FOV=$280\times180$mm, 72 slices, 2.0mm isotropic voxels).
The “minimally preprocessed” pipeline \cite{glass13} includes
gradient unwarping, motion correction, fieldmap-based EPI distortion
correction, brain-boundary-based registration of EPI to structural
T1-weighted scan, non-linear (FNIRT) registration into MNI space,
and grand-mean intensity normalization. Activation maps were spatially
smoothed by a Gaussian kernel of 4mm (FWHM). A general linear model (GLM) was
implemented by FILM from the FSL suite with model regressors from convolution
with a “canonical” hemodynamic response function and from temporal derivatives.
HCP tasks were conceived to modulate activation
in a maximum of different brain regions and neural systems. Indeed, at
least 70\% of the participants showed consistent brain activity in
contrasts from the task battery, which certifies excellent
coverage \cite{barch2013}.
In sum, the HCP task dataset incorporated 8650 first-level activity maps
from 18 diverse paradigms administered to 498 participants.
All maps were downsampled to a common 36x43x36 space of
5mm isotropic voxels and gray-matter masked (at least 10\%).
All analyses were based on task maps of
13,657 voxels representing Z values in gray matter.
\linebreak

Unsupervised and supervised learning were combined into a low-rank logistic
regression problem. The 13,657 z values from each activity map were
subject to a first linear
projection into Z latent components (i.e., 1, 5, 10, .., 100).
These hidden brain networks
loadings were subsequently projected into the 18 class space for multinomial
logistic regression. The goal was to find the two weight matrices
(input-to-hidden: 13,657 x Z and hidden-to-output: Z x 18) and their
corresponding bias vectors. Non-linearities were not applied on the
transformation results. Weights and biases were initialized by Gaussian
random noise. Gradient descent updated these matrices and vectors in each
iteration (100 samples per batch, 250 epochs). Using the chain rule, the
partial
derivates for the update were computed for the transformation into the
latent space and the subsequent transformation into the class space. We choose
the RMSprop algorithm \cite{dauphin15} with an initial learning rate of
0.01. Early stopping ensured that the best weight matrices/vectors were
retained in each epoch cycle. This was evaluated by the prediction
accuracy on a validation set (10\% of the training data) at each iteration.
\linebreak

This approach was benchmarked against independent dimensionality reduction and
learning of a classification function. Data reduction was performed on one
half of the data by ICA and SPCA.
ICA unmixed the
BOLD signals into separate spatial components
by minimizing their mutual information \cite{hyvarinen2000}.
This iterative blind source separation was
realized by a parallel FASTICA implementation (200 maximum iterations,
per-iteration tolerance of 0.0001,
initialized by a random mixing matrix, preliminary whitening).
SPCA separated the BOLD signals into
network components with few regions, which scales well to 
large datasets \cite{varoqu2011}.
This regression-type optimization problem constrained by
$\ell_1$-penalty in an implementation
without orthogonality assumptions
(1000 maximum iterations, per-iteration tolerance of 1 * 10\textsuperscript{-8}, sparsity
alpha=1, ridge-shrinkage at 0.01, Lasso path computed with coordinate
descent). Each linear decomposition revealed the specified number of
latent network components in one half the HCP data.
\linebreak
The extracted hidden network components were subsequently used
to reduce the remain half of task maps to a considerably smaller number of component
loadings. 13,657 voxels were thus condensed into 1, 5, 10, .., 100
measures of network involvement using ridge regression
(regularization alpha parameter=0.001, using
Cholesky solver). Multinomial logistic regression was finally performed on the
brain network loadings to learn classifying the 18 cognitive tasks.
Importantly, in all approaches, the final classification models were tested
on an unseen test set (10\% of data).
\linebreak

linear decomposition + elastic net penalty -> has characteristics of a
sparse PCA

Learning of the identity function as trivial solution is avoided by
the both bottleneck and the sparsity-enforcing ElasticNet penalty.

o maximize intranetwork homoge- neity and between-network heterogeneity.

AE:
coordinate system for points on the manifold

FORMULA
X e R^d real-valued input
'reconstruction error' criterion
encoder: input -> hidden reprsentation
Its parameter set is θ = {W, b}, where W is a d′ × d weight matrix and b is an offset vector of dimensionality d′.
decoder: gθ′ (y) = s(W′y + b′), (3)
with appropriately sized parameters θ′ = {W′,b′}.
where b0 and b1 are biase vectors
average reconstruction error
pen. avoids learning trivial soutoin - identitiy

affine autoencoder with squared error loss
tied weights beteen W0 and W1, i.e., W1 = W0^T
X hat is a dterministic function of X
It can thus be said that training an autoencoder to minimize reconstruction error amounts to maximizing a lower bound on the mutual information between input X and learnt repre- sentation Y .

rest data is already compressed to a sparse PCA -> remote form of 
pretraining of the first layer
bottleneck -> unter-complete representing, i.e. lossy compression

as a pretraining strategy

are decomposed into d networks,
in a way that maximizes the homogeneity of
function within each network while maximizing the heterogeneity between them.

output is vector-valued -> generative properties?
optimal transformation matrix

forward propegation: input through architecture to generate output
reconstruction cost
backward propegation: generate the deltas for all units / model paramters
corresponds to the derivative of the cost function with respect to the parameters

100 components close to sparse over-complete representation because
Xrest contains only 20 components

LR:

FORMULA

mutiple output units

RMSprop:
batch size:

20 components: high bias/low variance
100 compoentens: low bias/high variance

Generating samples from the learned statistical model

minimizing the discrepancy between the orig- inal data and its reconstruction.
The required gradients are easily obtained by using the chain rule to
backpropagate error derivatives first through the decoder network
and then through the encoder network

theano/python

\section{Experimental Results}
The source code, which can reproduce the most important results and
figures, is available at \url{http://github.com/banilo/nips2015}.

all vectors are column vectors

affine encoder and decoder

Low-rank regression outperformed serial ICA/SparsePCA and logistic regression.

Modifications of the mode that di not improve the genralization performacne:
dropo-out, input corruption, addining non-linearities (sigmoid, tanh),

introducing a
nonlinearity (sigmoid, tanh) into the system did not improve predictive accuracy but
elastic did -> most useful decomposition has characteristics of a SPCA
and not PCA or ICA

outperforms plain vanilla LR
inject prior domain knowledge into the learning process

\section{Discussion and Conclusion}
%

-hypothesis space includes sparse PCA and PCA but not ICA since no
linearity

respect structure in the fMRI data

- if linearity, then would be closer to the notion of 1-hidden layer neural
network rather than low-rank logistic regression

We hope that these results stimulate the development of
even more powerful semi-supervised classification methods

improve computational tractability, prediction accuracy, and interpretability
neuroimaging datasets
does it produce testable predictions?
-> we can test predictive value of network-network architectures across
mental domains.

open window to study the correspondence between brain dynamics during
every-day mind-wandering and task-focused brain states.

classifier that operates in a (sparse) network space, rather than in a voxel space.
domain-specifc classification algorithms

repertoire of mental operations that the human brain can perform

large quantities of neuroimaging data
logistic regression in an autoencoder paradigm

automatically learn a mapping to and from a brain-network space

statistical structure^

PCA:
captures structure in the data that is well described by Gaussian clouds,
linear pair-wise correlations are most important form of statistical
dependence/orthongonal components

-> BOLD images can readily be reduced to linear combinations of
sparse spatial structures

scale different classificatiuon architectures to large neuroimaging data

task data might concentrate near a low-dimensional manifold of brain networks

neurobiolgocial fidelity

entities of the neuroimaging domain

%
\paragraph{Acknowledgment}
{\small
Data were provided the Human Connectome Project.
}

\small
\bibliographystyle{splncs03}
\bibliography{nips_refs}

\end{document}

### 2015-05-15 paper_nips/paper.tex
first intro finished

@@
 
-%\nipsfinalcopy % Uncomment for camera-ready version
+\nipsfinalcopy % Uncomment for camera-ready version
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Existing neuroimaging methods typically perform either discovery of
-neurobiological structure or evaluate targeted hypotheses on mental tasks.
+neurobiological structure or evaluation of explicit hypotheses on mental tasks.
 %
-Modelling mental operations however hinges
+Modelling mental tasks however hinges
 on the pertinence of the assumed neurobiological structure.
 %
-We therefore propose to the solve the unsupervised dimensionality reduction
+We therefore propose to solve the unsupervised dimensionality reduction
 and supervised task classification in
@@
 
-\textbf{keywords}: dimensionality reduction; semi-supervised learning
+\textbf{keywords}: dimensionality reduction; semi-supervised learning;
 bioinformatics; fMRI; systems neuroscience
@@
 brain networks of
-functionally interaction brain regions.
+functionally interacting brain regions.
 Similarly, sparse PCA (SPCA; \cite{varoqu2011}) has been used to
@@
 Thus extracted brain networks have been shown to be
-manifestations of electrophysiological oscillation structure \cite{hipp15}.
+manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
 Their fundamental role is
@@
 by treating the brain as a nonlinear dynamic system with unknown
-neuronal states. As a last example, always more neuroimaging studies model
+ni  neuronal states. As a last example, always more neuroimaging studies model
 experimental tasks by training classification algorithms on brain signals
@@
 of baseline activity \cite{fox07}. They do thus not exploit the high similarity
-of BOLD dynamics of the human brain during experimental tasks and at rest.
+of BOLD dynamics of the human brain at rest and during experimental tasks.
 Indeed, similar brain networks were observed when applying ICA on
@@
 
-Both aspects can be combined into a 
-semi-supervised (i.e., use task and rest data)
+Both aspects can be conjointly exploited in a 
+semi-supervised (i.e., use rest and task data)
 low-rank (i.e., perform network decomposition)
-approach for neuroimaging analyses.
-
-Yet, special interest lies in the data-driven
-discovery of neurobiological structure that explains differences
-in cognitive conditions. Integrating unsupervised discovery and
-supervised classification should preferentially identify
-neurobiological structure that allows for the best predictive models. We
-propose such an approach based on conjoint gradient-descent to optimize
-linear decomposition and multi-task disambiguation.
-
-Find the most discriminative linear combination of the main direction of
-variation in the data
-
-modes of variation in the brain and the neural correlates of mental
-operations have mostly been studied separately so far
-their correspondence remains largely unclear
-
-we provide principled computational framework to link these previously
-unconnected domains of study.
-
-learn a manifold / low-dimensional coordinate system
-motivates
-
-There is much uncertainty about the most pertinent representation
-of neural activation information
-
-PCA, SPCA and ICA are produce compact representations of uncorrelated or
-independent components
-
-combine exploratory and confirmatory analysis steps in a single framework
-allows for  overlapping response pattersn
-
-adjust the weight assigned to each spatial pattern to maximize
-discriminatory power
-
-useful intermediate representation
-
-integrate recently gained domain knowledge into statistical learning
-problems.
-
-Traditionally, dimensionality reduction depended on linear methods such as PCA
-Autoencoder \cite{hinton06}:
-one-layered learning model that computes distributed
-representation from the inputs by improving reconstructions from them.
-dimensionality reduction by learning the underlying manifolds
-for efficient representation of the data
-bottleneck
--> use AE als feature extractions
-guide the estimator where to put probability mass in the inputs
-AE is a common feature learning algorithm
-if hidden layer linear and squared error loss, it behaves like a PCA \cite{baldi1989neural}
-convex with unique global minimum
-optimization objective is the same as for PCA
-
-Most nonsupervised learning algorithms can be implemented as variants in the 
-autoencoder framework. This includes decmoposition and clustering task
-
-Linear autoencoders with tied weights and specific regularization terms
-yield efficient algorithms for ICA / nonlinearity -> then closer to feed-forward neural network
-\cite{le2011ica}
-
-ICA, Sparse autoencoders, and sparse coding are mathematically equivalent
-under mild conditions \cite{le2011ica}
-
-AE can caputer global and local distribution
-
-and SPCA (??).
--> an autoencoder can capture various different compression procedures
-
-- reformulate the AE prbleme as a regularization boejctive
-
--> backpropagation: solve the unsupervised ata repsentation and supervised
-problem in an identical process.
-coompute the error derivates of the weights
--> gradually modified according to possible compression algorithms
-
-without a-priori knowledge of the structure
-affine transformation to compression information in the input
-class of problems
-architecture
-statistical models
-
-directions of greatest variance; represents each data point by its coordinates along each of these directions
-
-LR:
-these weights are then fine-tuned by the LR
-
-Even if linear should scan a wide space of intersring model, since
-theoretcially, a two-layered neural networks can learn almost any function.
-
-learn to the optimal projection of the input data
-into the main direction of variation and
-then classify along these principal directions.
-Vincent2010: it has been shown experimen- tally that
-beginning by optimizing an unsupervised criterion, oblivious of the specific
-classification problem, can actually greatly help in eventually achieving
-superior performance for that classifica- tion problem.
-
-among the models it will favor those that feature most neurobiological
-pertinence
-
-From the perspective of dictionary learning, the first layer can be
-viewed as a set of basis functions whose linear combinations are learned
-in the second layer \cite{olshausen96}.
-
-project the data-generating process on the axes of main variation
-autoencoder allows to reverse-engineer proerties of the data-generating
+approach to neuroimaging analysis.
+%
+The integration of brain-network discovery in a 
+supervised classification goal should identify the
+neurobiological structure that allows for the best predictive models.
+%
+Autoencoders suggest themselves because they can emulate
+variants of most nonsupervised learning algorithms,
+including PCA, SPCA, and ICA.
+Autoencoders
+are one-layered learning models that find local and global
+representation from the inputs
+by improving reconstructions from them \cite{hinton06}.
+%
+They behave like a PCA
+in case of one linear hidden layer and a squared error loss
+\cite{baldi1989neural}.
+This architecture yields a convex optimization objective
+with unique global minimum.
+Autoencoders behave like a SPCA if shrinkage terms are added for the
+matrix weights in the optimization objective.
+In turn, they behave like an ICA in case of a nonlinear convex
+function of the first-layer activation and tied weights \cite{le2011ica}.
+These authors further demonstrated that ICA, sparse autoencoders, and 
+sparse coding are mathematically equivalent
+under mild conditions.
+In this way, autoencoders can flexibly project the neuroimaging data
+onto the axes of main variation and thus
+reverse-engineer properties of the data-generating
 neural process \cite{olshausen96}.
 
-learn complex properties of the brain's task response patterns
+\linebreak
 
-convert high-dimensional data to low-dimensional network codes
-
-data should be more easily separable in the network spacew
-low-dimensional, linear structure
-
-equates with a shallow neural network without non-linearities
+In the present investigation,
+an autoencoder will be fed by (unlabeled) rest data and
+integrated as bottleneck
+into a low-rank logistic regression fed by (labeled) task data.
+Using the chain rule in back-propagation, we can then
+solve the unsupervised data representation and a supervised classification
+in an identical statistical learning problem.
+%
+From the perspective of dictionary learning, the first layer can be
+viewed as a learned set of basis functions
+whose linear combinations are learned
+in the second layer \cite{olshausen96}.
+%
+Neurobiologically, this allows 
+delineating a low-dimensional manifold of brain network patterns and then 
+distinguishing mental tasks
+by their most discriminative linear combinations.
+%
+Theoretically, a big reduction of the model variance is expected by using the
+resting-state autoencoder as regularization mechanisms
+to put probability mass on the most neurobiological
+valid members of the model space.
+%
+The generalization performance should consequently be improved due to 
+much reduced Vapnik-Chervonenkis dimensions of the classification estimator.
+%
+Taken together,
+the important modes of variation in brain dynamics and
+the neural correlates subserving mental operations
+have mostly been studied in isolation.
+We provide a principled computational framework to link these previously
+unconnected domains of systems neuroscience.
 
@@
 
+coompute the error derivates of the weights
+-> gradually modified according to possible compression algorithms
+
 affine autoencoder with squared error loss
@@
 
+There is much uncertainty about the most pertinent representation
+of neural activation information
+
 %

### 2015-05-17 paper_nips/paper.tex
solver passage

@@
 \usepackage[titletoc]{appendix}
+\usepackage{wrapfig}
+\usepackage{afterpage}
+\usepackage{amssymb}
 
@@
 \title{Semi-supervised low-rank logistic regression for
-high-dimensional neuroimaging datasets}
+high-dimensional neuroimaging data}
 
@@
 
-\author{Danilo Bzdok, Michael Eickenberg, Bertrand Thirion, Ga\"el
-  Varoquaux \\\\\textbf{\textit{email:} }firstname.lastname@inria.fr}
+\author{Danilo Bzdok, Michael Eickenberg, Olivier, Grisel,
+  Bertrand Thirion,
+  Ga\"el Varoquaux \\\\\textbf{\textit{email:} }firstname.lastname@inria.fr}
 
@@
 functionally interacting brain regions.
-Similarly, sparse PCA (SPCA; \cite{varoqu2011}) has been used to
+Similarly, sparse principal component analysis (SPCA; \cite{varoqu2011})
+has been used to
 separate brain activity signals into parsimonious network components.
@@
 manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
-Their fundamental role is
+Their fundamental role in brain organization is
 attested by continued covariation during sleep and anesthesia.
@@
 during ongoing random thought without controlled environmental input.
-The rationale is that the biggest fraction of BOLD signals do
-not correlate with a particular behavior, stimulus, or experimental task.
+The biggest fraction of the BOLD signals are known
+not to correlate with a particular behavior, stimulus, or experimental task.
 
@@
 by treating the brain as a nonlinear dynamic system with unknown
-ni  neuronal states. As a last example, always more neuroimaging studies model
+neuronal states. As a last example, always more neuroimaging studies model
 experimental tasks by training classification algorithms on brain signals
@@
 without exception in a single-voxel space. This ignores the fact that the BOLD
-signal contains coherent spatial activation patterns.
+signal exhibits coherent spatial activation patterns.
 Second, existing neuroimaging analyses do not acknowledge the fact that the
@@
 of baseline activity \cite{fox07}. They do thus not exploit the high similarity
-of BOLD dynamics of the human brain at rest and during experimental tasks.
-Indeed, similar brain networks were observed when applying ICA on
-large quantities of
-rest and task data \cite{smith2009}.
+of BOLD dynamics in the human brain at rest and during experimental tasks.
+Indeed, very similar brain networks were observed when applying ICA
+separately on rest and task data \cite{smith2009}.
 
@@
 
-Both aspects can be conjointly exploited in a 
+Both biological properties can be conjointly exploited in a 
 semi-supervised (i.e., use rest and task data)
 low-rank (i.e., perform network decomposition)
-approach to neuroimaging analysis.
+approach.
 %
@@
 Autoencoders
-are one-layered learning models that find local and global
-representation from the inputs
-by improving reconstructions from them \cite{hinton06}.
+are one-layered learning models that condense the input data to
+local and global representations
+by improving reconstruction from them \cite{hinton06}.
 %
@@
 reverse-engineer properties of the data-generating
-neural process \cite{olshausen96}.
+neural processes \cite{olshausen96}.
 
@@
 %
-Theoretically, a big reduction of the model variance is expected by using the
-resting-state autoencoder as regularization mechanisms
-to put probability mass on the most neurobiological
+Theoretically, a big reduction of the model variance is expected by
+regularization using resting-state autoencoder
+to put probability mass on the most neurobiologically
 valid members of the model space.
@@
 unconnected domains of systems neuroscience.
+
+\begin{wrapfigure}{r}{0.40\textwidth}
+  \begin{center}
+    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
+  \end{center}
+  \caption {\textbf{Architecture}
+  }
+\end{wrapfigure}
 
@@
 sparse PCA
+\begin{wrapfigure}{r}{0.40\textwidth}
+  \begin{center}
+    \includegraphics[width=0.40\textwidth]{figures/figure3.png}
+  \end{center}
+  \caption {\textbf{Equilibrium between autoencoder and low-rank logistic regression}
+  One learned decomposition component (out of 20) between the only-autoencoder
+  (\textit{uppermost panel}) and only-logistic-regression
+  (\textit{lowermost panel}) scenarios.
+  The congruent structure (\textit{red, middle column})
+  in the
+  posterior cingulate cortex, posterior mid-cingulate cortex, and medial
+  prefrontal cortex that is hence important in decompositions of the rest and
+  task neuroimaging data. With increasing weight on the supervised learning,
+  non-congruent structure emerges in the early (\textit{blue}) and
+  lateral (\textit{red}) visual cortex
+  (\textit{right column}). Matrix weights were z-scored.
+  }
+\end{wrapfigure}
 
@@
 coordinate system for points on the manifold
+low-dimensional bottleneck layer
 
@@
 
-LR:
+whereas elasticnet on the model paramters is regulization
+
+HINTS:
+In fact, the constraint by a rest-data autoencoder qualifies as a
+\textit{hint} rather than regularization \cite{abu1994hints}.
+Its purpose is not to prevent overfitting but to introduce
+prior knowledge on known properties of the unknown target function $f$.
+Rather than relying only on input-output pairs in the learning process,
+we thus narrow our hypothesis set to the biologically most plausible solutions.
+That is, we reduce the search space in a way that
+is compatible with the expected representation of BOLD activity signals.
 
@@
 
+{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda\frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; W,b)) \hspace{0.2cm} + \hspace{0.2cm} (1-\lambda)\frac{1}{N_{X_{rest}}}\|X_{rest}-\hat X_{rest}(W,b)\|^2
+
 mutiple output units
 
-RMSprop:
-batch size:
+\linebreak
+
+As solver, we chose \textit{rmsprop},
+a mini-batch version of rprop \cite{dauphin2015rmsprop}.
+This procedure dictates an adaptive learning rate
+for each model parameter by
+scaled gradients from a running average.
+Gradient normalization by RMSprop
+is known to effectively exploit curvature information.
+We opted for a small batch size of $100$, given the high degree of
+redundancy in $X_{rest}$ and $X_{task}$.
+The matrix parameters were initalized by Gaussian random values multiplied
+by a gain of $0.004$. The bias parameters were initalized to be zero.
+The normalization factor is computed by
+
+\begin{eqnarray*}
+v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)(\nabla f)^2
+\end{eqnarray*}
+
+where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
+$0.9$ to deemphasize the magnitude of the gradient.
+The update step is
+given by
+
+\begin{eqnarray*}
+\theta^{(t+1)} = \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{v^{(t+1)} + \epsilon}}
+\end{eqnarray*}
+
+where $\alpha$ is the learning rate and $\epsilon$ a damping factor.
+The hyper-parameter $\alpha$ was set to $0.00001$ by prior manuel
+cross-validation and $\epsilon$ was set to $1^{-6}$.
+%
+Note that we have also experimented with other solvers
+(stochastic gradient descent, adadelta, and adagrad) but found that
+rmsprop converged faster and with higher generalization performance.
+
+\linebreak
 
@@
 theano/python
+The analyses were performed in Python.
+We used \textit{nilearn} to handle
+the high-dimensional neuroimaging data and
+\textit{theano} for automatic gradient
+differentiation of symbolic computation graphs.
+All Python scripts that generated the results are
+accessible online for reproducibility and reuse
+\url{http://github.com/banilo/nips2015}.
 
 \section{Experimental Results}
-The source code, which can reproduce the most important results and
-figures, is available at \url{http://github.com/banilo/nips2015}.
 
@@
 Low-rank regression outperformed serial ICA/SparsePCA and logistic regression.
+
+reduction of n gray-matter voxels to n components
 
@@
 %
+There is an increasing occurrence of high-dimensional problems in the
+neuroimaging domain. This calls for new statistical learning algorithms that
+behave well in large-cohort settings. Ideally, they should acknowledge
+and exploit existing widely-accepted neuroscientific knowledge.
+In the present work, we propose such an estimator that learns
+dimensionality reduction
+in a neurobiologically valid and interpretable fashion.
+
+\linebreak
 
@@
 of neural activation information
+put probability mas shwere we expeted neurobiological strucuture
+
+canonical set of brain networks
 
@@
 {\small
-Data were provided the Human Connectome Project.
+Data were provided the Human Connectome Project. The study was supported
+by the German National Academic Foundation (D.B.).
 }

### 2015-05-18 paper_nips/paper.tex
add missing files

@@
 
-\maketitle
-
-\author{Danilo Bzdok, Michael Eickenberg, Olivier, Grisel,
+\author{Danilo Bzdok, Michael Eickenberg, Olivier Grisel,
   Bertrand Thirion,
   Ga\"el Varoquaux \\\\\textbf{\textit{email:} }firstname.lastname@inria.fr}
+
+\maketitle
 
@@
 
-\linebreak
-
 On the other hand, to investigate
@@
 separately on rest and task data \cite{smith2009}.
-
-\linebreak
 
@@
 neural processes \cite{olshausen96}.
-
-\linebreak
 
@@
 latent space and the subsequent transformation into the class space. We choose
-the RMSprop algorithm \cite{dauphin15} with an initial learning rate of
+the RMSprop algorithm \cite{rmsprop} with an initial learning rate of
 0.01. Early stopping ensured that the best weight matrices/vectors were
@@
 FORMULA
-X e R^d real-valued input
+%X e R^d real-valued input
 'reconstruction error' criterion
 encoder: input -> hidden reprsentation
-Its parameter set is θ = {W, b}, where W is a d′ × d weight matrix and b is an offset vector of dimensionality d′.
-decoder: gθ′ (y) = s(W′y + b′), (3)
-with appropriately sized parameters θ′ = {W′,b′}.
+%Its parameter set is θ = {W, b}, where W is a d′ × d weight matrix and b is an offset vector of dimensionality d′.
+%decoder: gθ′ (y) = s(W′y + b′), (3)
+%with appropriately sized parameters θ′ = {W′,b′}.
 where b0 and b1 are biase vectors
@@
 coompute the error derivates of the weights
--> gradually modified according to possible compression algorithms
+gradually modified according to possible compression algorithms
 
 affine autoencoder with squared error loss
-tied weights beteen W0 and W1, i.e., W1 = W0^T
+%tied weights beteen W0 and W1, i.e., W1 = W0^T
 X hat is a dterministic function of X
@@
 
-rest data is already compressed to a sparse PCA -> remote form of 
+rest data is already compressed to a sparse PCA remote form of 
 pretraining of the first layer
@@
 
-HINTS:
+\paragraph{Hints.}
 In fact, the constraint by a rest-data autoencoder qualifies as a
@@
 
-{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda\frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; W,b)) \hspace{0.2cm} + \hspace{0.2cm} (1-\lambda)\frac{1}{N_{X_{rest}}}\|X_{rest}-\hat X_{rest}(W,b)\|^2
+\begin{eqnarray}
+  \begin{split}
+{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) &= \lambda\frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; W,b)) \hspace{0.2cm} \\
+&+ \hspace{0.2cm} (1-\lambda)\frac{1}{N_{X_{rest}}}\|X_{rest}-\hat X_{rest}(W,b)\|^2
+  \end{split}
+  \label{eq:loss_equ}
+\end{eqnarray}
 
 mutiple output units
-
-\linebreak
-
-As solver, we chose \textit{rmsprop},
-a mini-batch version of rprop \cite{dauphin2015rmsprop}.
+aCcording to equation \eqref{eq:loss_equ}
+As solver, we chose \textit{rmsprop} \cite{rmsprop},
+a mini-batch version of rprop.
 This procedure dictates an adaptive learning rate
@@
 by a gain of $0.004$. The bias parameters were initalized to be zero.
-The normalization factor is computed by
 
-\begin{eqnarray*}
-v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)(\nabla f)^2
-\end{eqnarray*}
+With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
+The normalization factor and the update rule for $\theta$
+are given by
+
+\begin{eqnarray}
+  \begin{split}
+v^{(t+1)} &= \rho v^{(t+1)} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
+%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
+\\
+\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{v^{(t+1)} + \epsilon}},
+  \end{split}
+\end{eqnarray}
 
@@
 $0.9$ to deemphasize the magnitude of the gradient.
-The update step is
-given by
-
-\begin{eqnarray*}
-\theta^{(t+1)} = \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{v^{(t+1)} + \epsilon}}
-\end{eqnarray*}
-
-where $\alpha$ is the learning rate and $\epsilon$ a damping factor.
+Further $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
 The hyper-parameter $\alpha$ was set to $0.00001$ by prior manuel
@@
 rmsprop converged faster and with higher generalization performance.
-
-\linebreak
 
@@
 
-theano/python
+theanopython
 The analyses were performed in Python.
 We used \textit{nilearn} to handle
-the high-dimensional neuroimaging data and
-\textit{theano} for automatic gradient
-differentiation of symbolic computation graphs.
+the high-dimensional neuroimaging data 
+\cite{abrah14}
+and
+\textit{theano} for automatic
+differentiation of symbolic computation graphs
+\cite{bastien2012theano, bergstra2010theano}.
 All Python scripts that generated the results are
@@
 
-\linebreak
-
 -hypothesis space includes sparse PCA and PCA but not ICA since no
@@
 
-statistical structure^
+statistical structure
 
@@
 
+FUTURE
+link the restings-tate component to the pattern sof cognitive proceses
+reduce information von resting-state data in neurological and psychiatric
+populations for discovery of neurobiological sub-groups as well as
+prediction of disease trajectories and drug responses
+
 %

### 2015-05-18 nips3mm.py

@@ -1,0 +1,1032 @@
+"""
+HCP: semi-supervised network decomposition by low-rank logistic regression
+"""
+"""
+preliminary conclusion:
+- the combined AE and LR loss yields best classification results
+- Regarding rest data, 20-comp PCA work better than uncompressed data
+questions:
+- the AE cost is much bigger, is it ok to T.log() that?
+No because it renders the objective function non-convex.
+- compare to nonlinearity + PCA only
+- how to genreate examples from 100 last model paramters during the
+gradient descent?
+"""
+
+print __doc__
+
+import os
+import os.path as op
+import numpy as np
+import glob
+from scipy.linalg import norm
+import nibabel as nib
+from sklearn.grid_search import RandomizedSearchCV
+from sklearn.base import BaseEstimator
+from sklearn.preprocessing import StandardScaler
+from nilearn.input_data import NiftiMasker
+from sklearn.metrics import precision_recall_fscore_support
+from sklearn.metrics import confusion_matrix
+import theano
+import theano.tensor as T
+from matplotlib import pylab as plt
+print('Running THEANO on %s' % theano.config.device)
+from nilearn.image import concat_imgs
+import joblib
+import time
+
+RES_NAME = 'nips3mm'
+WRITE_DIR = op.join(os.getcwd(), RES_NAME)
+if not op.exists(WRITE_DIR):
+    os.mkdir(WRITE_DIR)
+
+##############################################################################
+# load+preprocess data
+##############################################################################
+
+mask_img = 'grey10_icbm_3mm_bin.nii.gz'
+nifti_masker = NiftiMasker(mask_img=mask_img, smoothing_fwhm=False,
+                           standardize=False)
+nifti_masker.fit()
+mask_nvox = nifti_masker.mask_img_.get_data().sum()
+
+print('Loading data...')
+
+# ARCHI task
+X_task, labels = joblib.load('preload_HT_3mm')
+
+labels = np.int32(labels)
+
+# contrasts are IN ORDER -> shuffle!
+new_inds = np.arange(0, X_task.shape[0])
+np.random.shuffle(new_inds)
+X_task = X_task[new_inds]
+labels = labels[new_inds]
+# subs = subs[new_inds]
+
+# rest
+# X_rest = nifti_masker.transform('preload_HR20persub_10mm_ero2.nii')
+# X_rest = nifti_masker.transform('dump_rs_spca_s12_tmp')
[…]

### 2015-05-18 nips3mm_h38.py

@@ -1,0 +1,985 @@
+"""
+HCP: semi-supervised network decomposition by low-rank logistic regression
+"""
+"""
+preliminary conclusion:
+- the combined AE and LR loss yields best classification results
+- Regarding rest data, 20-comp PCA work better than uncompressed data
+questions:
+- the AE cost is much bigger, is it ok to T.log() that?
+No because it renders the objective function non-convex.
+"""
+
+print __doc__
+
+import os
+import os.path as op
+import numpy as np
+import glob
+from scipy.linalg import norm
+import nibabel as nib
+from sklearn.grid_search import RandomizedSearchCV
+from sklearn.base import BaseEstimator
+from sklearn.preprocessing import StandardScaler
+from nilearn.input_data import NiftiMasker
+from sklearn.metrics import precision_recall_fscore_support
+from sklearn.metrics import confusion_matrix
+import theano
+import theano.tensor as T
+print('Running THEANO on %s' % theano.config.device)
+from nilearn.image import concat_imgs
+import joblib
+import time
+
+RES_NAME = 'nips3mm_h38'
+WRITE_DIR = op.join(os.getcwd(), RES_NAME)
+if not op.exists(WRITE_DIR):
+    os.mkdir(WRITE_DIR)
+
+##############################################################################
+# load+preprocess data
+##############################################################################
+
+mask_img = 'grey10_icbm_3mm_bin.nii.gz'
+nifti_masker = NiftiMasker(mask_img=mask_img, smoothing_fwhm=False,
+                           standardize=False)
+nifti_masker.fit()
+mask_nvox = nifti_masker.mask_img_.get_data().sum()
+
+print('Loading data...')
+
+# ARCHI task
+X_task, labels = joblib.load('preload_HT38_3mm')
+
+labels = np.int32(labels)
+
+# contrasts are IN ORDER -> shuffle!
+new_inds = np.arange(0, X_task.shape[0])
+np.random.shuffle(new_inds)
+X_task = X_task[new_inds]
+labels = labels[new_inds]
+# subs = subs[new_inds]
+
+# rest
+# X_rest = nifti_masker.transform('preload_HR20persub_10mm_ero2.nii')
+# X_rest = nifti_masker.transform('dump_rs_spca_s12_tmp')
+rs_spca_data = joblib.load('dump_rs_spca_s12_tmp')
+rs_spca_niis = nib.Nifti1Image(rs_spca_data,
+                               nifti_masker.mask_img_.get_affine())
[…]

### 2015-05-18 nips3mm_vanilla.py

@@ -1,0 +1,905 @@
+"""
+HCP: semi-supervised network decomposition by low-rank logistic regression
+"""
+"""
+"""
+
+print __doc__
+
+import os
+import os.path as op
+import numpy as np
+import glob
+from scipy.linalg import norm
+import nibabel as nib
+from sklearn.grid_search import RandomizedSearchCV
+from sklearn.base import BaseEstimator
+from sklearn.preprocessing import StandardScaler
+from nilearn.input_data import NiftiMasker
+from sklearn.metrics import precision_recall_fscore_support
+from sklearn.metrics import confusion_matrix
+import theano
+import theano.tensor as T
+from matplotlib import pylab as plt
+print('Running THEANO on %s' % theano.config.device)
+from nilearn.image import concat_imgs
+import joblib
+import time
+
+RES_NAME = 'nips3mm_vanilla'
+WRITE_DIR = op.join(os.getcwd(), RES_NAME)
+if not op.exists(WRITE_DIR):
+    os.mkdir(WRITE_DIR)
+
+##############################################################################
+# load+preprocess data
+##############################################################################
+
+mask_img = 'grey10_icbm_3mm_bin.nii.gz'
+nifti_masker = NiftiMasker(mask_img=mask_img, smoothing_fwhm=False,
+                           standardize=False)
+nifti_masker.fit()
+mask_nvox = nifti_masker.mask_img_.get_data().sum()
+
+print('Loading data...')
+
+# ARCHI task
+X_task, labels = joblib.load('preload_HT_3mm')
+
+labels = np.int32(labels)
+
+# contrasts are IN ORDER -> shuffle!
+new_inds = np.arange(0, X_task.shape[0])
+np.random.shuffle(new_inds)
+X_task = X_task[new_inds]
+labels = labels[new_inds]
+# subs = subs[new_inds]
+
+X_task = StandardScaler().fit_transform(X_task)
+
+# ARCHI task
+AT_niis, AT_labels, AT_subs = joblib.load('preload_AT_3mm')
+AT_X = nifti_masker.transform(AT_niis)
+AT_X = StandardScaler().fit_transform(AT_X)
+print('done :)')
+
+##############################################################################
+# define computation graph
+##############################################################################
+
+class SSEncoder(BaseEstimator):
+    def __init__(self, gain1, learning_rate, max_epochs=100,
+                 l1=0.1, l2=0.1):
+        """
+        Parameters
+        ----------
+        lambda : float
[…]

### 2015-05-18 nips3mm_vanilla.py
LR vanilla working 3mm

@@ -101,35 +101,10 @@ class SSEncoder(BaseEstimator):
         
     def get_param_pool(self):
         cur_params = (
-            self.W0s, self.bW0s
+            self.V0s, self.bV0
         )
         return cur_params
         
-    def test_performance_in_other_dataset(self):
-        from sklearn.linear_model import LogisticRegression
-        from sklearn.cross_validation import StratifiedShuffleSplit
-
-        compr_matrix = self.W0s.get_value().T  # currently best compression
-        AT_X_compr = np.dot(compr_matrix, AT_X.T).T
-        clf = LogisticRegression(penalty='l1')
-        folder = StratifiedShuffleSplit(y=AT_labels, n_iter=5, test_size=0.2)
-
-        acc_list = []
-        prfs_list = []
-        for (train_inds, test_inds) in folder:
-            clf.fit(AT_X_compr[train_inds, :], AT_labels[train_inds])
-            pred_y = clf.predict(AT_X_compr[test_inds, :])
-
-            acc = (pred_y == AT_labels[test_inds]).mean()
-            prfs_list.append(precision_recall_fscore_support(
-                             AT_labels[test_inds], pred_y))
-
-            acc_list.append(acc)
-
-        compr_mean_acc = np.mean(acc_list)
-        prfs = np.asarray(prfs_list).mean(axis=0)
-        return compr_mean_acc, prfs
-
     def fit(self, X_task, y):
         DEBUG_FLAG = True
 
@@ -139,6 +114,7 @@ class SSEncoder(BaseEstimator):
         self.input_taskdata = T.matrix(dtype='float32', name='input_taskdata')
         self.input_restdata = T.matrix(dtype='float32', name='input_restdata')
         self.params_from_last_iters = []
+        n_input = X_task.shape[1]
 
         index = T.iscalar(name='index')
         
@@ -172,7 +148,6 @@ class SSEncoder(BaseEstimator):
             self.dbg_lr_cost_ = list()
             self.dbg_ae_nonimprovesteps = list()
             self.dbg_acc_other_ds_ = list()
-            self.dbg_combined_cost_ = list()
             self.dbg_prfs_ = list()
             self.dbg_prfs_other_ds_ = list()
 
@@ -218,6 +193,7 @@ class SSEncoder(BaseEstimator):
         # optimization loop
         start_time = time.time()
         lr_last_cost = np.inf
+        ae_cur_cost = np.inf
         no_improve_steps = 0
[…]

### 2015-05-19 nips3mm_vanilla.py
before plot dict

@@ -350,51 +350,46 @@ V1_mat = estimator.V1s.get_value().T
 np.save(cur_path + 'V1comps', V1_mat)
 dump_comps(nifti_masker, fname, comps, threshold=0.25)
 
-# equally scaled plots
-# import re
-# pkgs = glob.glob(RES_NAME + '/*dbg_epochs*.npy')
-# dbg_epochs_ = np.load(pkgs[0])
-# n_comps = [100]
-# pkgs = glob.glob(RES_NAME + '/*dbg_acc_train*.npy')
-# for n_comp in n_comps:
-#     plt.figure()
-#     for p in pkgs:
-#         lambda_param = np.float(re.search('lambda=(.{4})', p).group(1))
-#         # n_hidden = int(re.search('comp=(?P<comp>.{1,2,3})_', p).group('comp'))
-#         n_hidden = int(re.search('comp=(.{1,3})_', p).group(1))
-#         if n_comp != n_hidden:
-#             continue
-#         
-#         dbg_acc_train_ = np.load(p)
-#         
-#         cur_label = 'n_comp=%i' % n_hidden
-#         cur_label += '/'
-#         cur_label += 'lambda=%.2f' % lambda_param
-#         cur_label += '/'
-#         if not '_AE' in p:
-#             cur_label += 'LR only!'
-#         elif 'subRS' in p:
-#             cur_label += 'RSnormal'
-#         elif 'spca20RS' in p:
-#             cur_label += 'RSspca20'
-#         elif 'pca20RS' in p:
-#             cur_label += 'RSpca20'
-#         cur_label += '/'
-#         cur_label += 'separate decomp.' if 'decomp_separate' in p else 'joint decomp.'
-#         cur_label += '' if '_AE' in p else '/LR only!'
-#         plt.plot(
-#             dbg_epochs_,
-#             dbg_acc_train_,
-#             label=cur_label)
-#     plt.title('Low-rank LR+AE L1=0.1 L2=0.1 res=3mm combined-loss')
-#     plt.legend(loc='lower right', fontsize=9)
-#     plt.yticks(np.linspace(0., 1., 11))
-#     plt.ylabel('training accuracy')
-#     plt.xlabel('epochs')
-#     plt.ylim(0., 1.05)
-#     plt.grid(True)
-#     plt.show()
[…]

### 2015-05-23 paper_nips/paper.tex
methods standing

@@
 %
-Using
-Human Connectome Project (HCP) data (n=500), optimal low-rank projections and
-logistic-regression models are identified in a same gradient descent. Brain
+\paragraph{Data.}
+Unsupervised projection models into a lower-dimensional space and
+supervised logistic-regression models are learned in concert
+in a same parameter gradient descent.
+Brain
 network decompositions are thus exposed that explain task-descriminative spatial
-patterns.
+patterns of neural activity.
 
-Data was drawn from 498 unrelated, healthy HCP participants.
-All provided informed consent to the Washington
-University in St. Louis institutional review board. HCP tasks 
-were selected that feature known suitability as localizers
-and reliability
-across participants \cite{barch2013}.
-Mostly block-design, but also event-related, paradigms were
-administered on 1) working memory/cognitive control processing, 2)
+As the currently biggest openly-accessible reference dataset,
+we chose the Human Connectome Project (HCP) for the present analyses
+\cite{barch2013}.
+Neuroimaging task data with labels of ongoing cognitive processes
+were drawn from 500 unrelated,
+healthy HCP participants.
+18 HCP tasks 
+were selected that are known to elicit reliable neural activity
+across participants.
+The task paradigms include
+1) working memory/cognitive control processing, 2)
 incentive processing, 3) visual and somatosensory-motor processing,
@@
 in-plane FOV=$280\times180$mm, 72 slices, 2.0mm isotropic voxels).
-The “minimally preprocessed” pipeline \cite{glass13} includes
+The “minimally preprocessed” pipeline includes
 gradient unwarping, motion correction, fieldmap-based EPI distortion
 correction, brain-boundary-based registration of EPI to structural
-T1-weighted scan, non-linear (FNIRT) registration into MNI space,
+T1-weighted scans, non-linear (FNIRT) registration into MNI space,
 and grand-mean intensity normalization. Activation maps were spatially
-smoothed by a Gaussian kernel of 4mm (FWHM). A general linear model (GLM) was
+smoothed by a Gaussian kernel of 4mm (FWHM). A GLM was
 implemented by FILM from the FSL suite with model regressors from convolution
@@
 coverage \cite{barch2013}.
-In sum, the HCP task dataset incorporated 8650 first-level activity maps
-from 18 diverse paradigms administered to 498 participants.
-All maps were downsampled to a common 36x43x36 space of
-5mm isotropic voxels and gray-matter masked (at least 10\%).
-All analyses were based on task maps of
-13,657 voxels representing Z values in gray matter.
-\linebreak
+In sum, the HCP task data incorporated 8650 first-level activity maps
+from 18 diverse paradigms administered to 498 participants (2 removed
+due to incomplete data).
+All maps were resampled to a common 60x72x60 space of
+3mm isotropic voxels and gray-matter masked (at least 10\% tissue
+probability).
+All supservised analyses were based on labeled HCP task maps of
+79,941 voxels representing Z values in gray matter.
 
-Unsupervised and supervised learning were combined into a low-rank logistic
-regression problem. The 13,657 z values from each activity map were
-subject to a first linear
-projection into Z latent components (i.e., 1, 5, 10, .., 100).
-These hidden brain networks
-loadings were subsequently projected into the 18 class space for multinomial
-logistic regression. The goal was to find the two weight matrices
-(input-to-hidden: 13,657 x Z and hidden-to-output: Z x 18) and their
-corresponding bias vectors. Non-linearities were not applied on the
-transformation results. Weights and biases were initialized by Gaussian
-random noise. Gradient descent updated these matrices and vectors in each
-iteration (100 samples per batch, 250 epochs). Using the chain rule, the
-partial
-derivates for the update were computed for the transformation into the
-latent space and the subsequent transformation into the class space. We choose
-the RMSprop algorithm \cite{rmsprop} with an initial learning rate of
-0.01. Early stopping ensured that the best weight matrices/vectors were
-retained in each epoch cycle. This was evaluated by the prediction
-accuracy on a validation set (10\% of the training data) at each iteration.
-\linebreak
+These labeled data were complemented by unlabeled activity maps
+from HCP acquisitions of unconstrained resting-state activity.
+These reflect brain activity in the absence of controlled thought.
+In line with the goal of the present study, acquisition of the data was
+specifically aimed at the study of task-rest correspondence.
+From each included participant, we included two
+time-series for left and right phase encoding
+with 1,200 maps of multiband, gradient-echo planar imaging acquired
+during a period of 15min (TR=720 ms, TE=33.1 ms, flip angle=52°,
+FOV=$280\times180$mm, and 2.0mm isotropic voxels). Besides run duration,
+the task acquisitions were identical to the resting-state fMRI acquisitions
+for maximal compatibility between task and rest data.
+We here drew on ``minimally preprocessed'' rest data
+from 200 randomly selected healthy, unrelated participants.
+PCA was applied to each set of 1,200 rest maps for
+denoising by keeping only the 20 main modes of
+variation.
+In sum, the HCP rest data concatenated
+8000 unlabeled, noise-cleaned rest maps with
+40 brain images from each of 200 randomly selected participants.
 
-This approach was benchmarked against independent dimensionality reduction and
-learning of a classification function. Data reduction was performed on one
-half of the data by ICA and SPCA.
-ICA unmixed the
-BOLD signals into separate spatial components
-by minimizing their mutual information \cite{hyvarinen2000}.
-This iterative blind source separation was
-realized by a parallel FASTICA implementation (200 maximum iterations,
-per-iteration tolerance of 0.0001,
-initialized by a random mixing matrix, preliminary whitening).
-SPCA separated the BOLD signals into
-network components with few regions, which scales well to 
-large datasets \cite{varoqu2011}.
-This regression-type optimization problem constrained by
-$\ell_1$-penalty in an implementation
-without orthogonality assumptions
-(1000 maximum iterations, per-iteration tolerance of 1 * 10\textsuperscript{-8}, sparsity
-alpha=1, ridge-shrinkage at 0.01, Lasso path computed with coordinate
-descent). Each linear decomposition revealed the specified number of
-latent network components in one half the HCP data.
-\linebreak
-The extracted hidden network components were subsequently used
-to reduce the remain half of task maps to a considerably smaller number of component
-loadings. 13,657 voxels were thus condensed into 1, 5, 10, .., 100
-measures of network involvement using ridge regression
-(regularization alpha parameter=0.001, using
-Cholesky solver). Multinomial logistic regression was finally performed on the
-brain network loadings to learn classifying the 18 cognitive tasks.
-Importantly, in all approaches, the final classification models were tested
-on an unseen test set (10\% of data).
-\linebreak
+We were further interested in the utility of the optimal low-rank projection
+in one task dataset for dimensionality reduction in another task dataset.
+To this end, the HCP-derived network decompositions were used as preliminary
+step in the classifcation problem of another large-cohort dataset.
+The ARCHI dataset \cite{pinel07} provides activity maps from
+diverse experimental tasks, including auditory and visual perception, motor action,
+reading, language comprehension and mental calculation.
+81 right-handed healthy participants
+(3 not included in present analyses due to incomplete data)
+without psychiatric or
+neurological history participated in four fMRI sessions acquired under
+different experimental paradigms.
+The functional maps were warped into
+the MNI space and resampled to isotropic 3mm resolution.
+Whole-brain EPI data were acquired with the same Siemens Trio with a 32
+channel head coil (TR=2400ms, TE=30ms, flip angle=60°, in-plane
+FOV=$19.2\times19.2$cm, 40 slices, 3.0mm isotropic voxels).
+Standard preprocessing was performed with Nipype \cite{gorgo11}, including
+slice timing, motion correction, alignment, and spatial normalization.
+Activation maps were spatially smoothed by
+a Gaussian kernel of 5mm (FWHM).
+Analogous to HCP data, the second task dataset incorporated 1404
+labeled, grey-matter masked, and z-scored activity maps
+from 18 diverse tasks acquired in 78 participants.
 
-linear decomposition + elastic net penalty -> has characteristics of a
-sparse PCA
+\paragraph{Unsupervised layer.}
+The labeled and unlabeled data were fed into a linear statistical model.
+The model is composed of an autoencoder and a low-rank logistic regression.
+On the one hand, the affine autoencoder takes the input 
+$x$ and projects it into coordinates of a latent representation $z$
+by
+
+\begin{eqnarray}
+  \begin{split}
+  z &= W_0 x + b_{W_0} \\
+  x' &= W_1 z + b_{W_1}
+  \end{split}
+  \label{eq:autoenc}
+\end{eqnarray}
+
+where $x \in \mathbb{R}^{d}$ denotes the vector of $d=79,941$ voxel from each
+rest map, $z \in \mathbb{R}^{n}$ is the $n$ hidden dimensions (i.e.,
+distributed macroscopical activity patterns), and 
+$x' \in \mathbb{R}^{d}$ is the reconstruction vector of the original activity map
+from the hidden variables. Further, ${W_0}$ denotes the weight matrix that
+transforms
+from input space into the hidden space (i.e., encoder),
+${W_1}$ is the weight matirx for back-projection from the hidden space to the
+output space (i.e., decoder), ${b_{W_0}}$ and ${b_{W_1}}$ are bias vectors.
+Note that ${W_0}$ and ${W_1}$ are tied
+$W_0$ = $W_1^T$.
+The optimal model parameters ${W_0}, {b_{W_0}}, {b_{W_1}}$ are found by
+minimizing the reconstruction difference by squared error
+
+\begin{eqnarray}
+  {\mathcal{L_{AE}}}(x, x') = || \x - \x' ||^2
+\end{eqnarray}
+
+This reconstruction error criterion equates with
+maximizing a lower bound on the mutual information between
+input and the learned representation.
+Non-linearities were not applied on the
+transformation results.
+
+\paragraph{Supervised layer.}
+On the other hand, such lossy compression by a low-dimensional bottleneck
+is also imposed by the first layer of the low-rank
+logistic regression.
+
+\begin{eqnarray}
+  \begin{split}
+    P(Y=i|x, V_0,V_1,b_{V_0}, b_{V_1}) &= softmax_i(V_1 (V_0 x + b_{V_0}) + b_{V_1}) \\
+    &= \frac {e^{V_1_i (V_0_i x + b_{V_0}_i) + b_{V_1}_i}} {\sum_j e^{V_1_j (V_0_j x + b_{V_0}_j) + b_{V_1}_j}}            
+  \end{split}
+  \label{eq:lr}
+\end{eqnarray}
+
+\begin{eqnarray}
+  \begin{split}
+    {\mathcal{L_{LR}}}(x, y) = \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1})
+\end{split}
+\label{eq:lr_loss}
+\end{eqnarray}
+
 \begin{wrapfigure}{r}{0.40\textwidth}
@@
 
-Learning of the identity function as trivial solution is avoided by
-the both bottleneck and the sparsity-enforcing ElasticNet penalty.
-
-o maximize intranetwork homoge- neity and between-network heterogeneity.
-
-AE:
-coordinate system for points on the manifold
-low-dimensional bottleneck layer
-
-FORMULA
-%X e R^d real-valued input
-'reconstruction error' criterion
-encoder: input -> hidden reprsentation
-%Its parameter set is θ = {W, b}, where W is a d′ × d weight matrix and b is an offset vector of dimensionality d′.
-%decoder: gθ′ (y) = s(W′y + b′), (3)
-%with appropriately sized parameters θ′ = {W′,b′}.
-where b0 and b1 are biase vectors
-average reconstruction error
-pen. avoids learning trivial soutoin - identitiy
-
-coompute the error derivates of the weights
-gradually modified according to possible compression algorithms
-
-affine autoencoder with squared error loss
-%tied weights beteen W0 and W1, i.e., W1 = W0^T
-X hat is a dterministic function of X
-It can thus be said that training an autoencoder to minimize reconstruction error amounts to maximizing a lower bound on the mutual information between input X and learnt repre- sentation Y .
-
-rest data is already compressed to a sparse PCA remote form of 
-pretraining of the first layer
-bottleneck -> unter-complete representing, i.e. lossy compression
-
-as a pretraining strategy
-
-are decomposed into d networks,
-in a way that maximizes the homogeneity of
-function within each network while maximizing the heterogeneity between them.
-
-output is vector-valued -> generative properties?
-optimal transformation matrix
-
-forward propegation: input through architecture to generate output
-reconstruction cost
-backward propegation: generate the deltas for all units / model paramters
-corresponds to the derivative of the cost function with respect to the parameters
-
-100 components close to sparse over-complete representation because
-Xrest contains only 20 components
-
-whereas elasticnet on the model paramters is regulization
-
-\paragraph{Hints.}
-In fact, the constraint by a rest-data autoencoder qualifies as a
-\textit{hint} rather than regularization \cite{abu1994hints}.
-Its purpose is not to prevent overfitting but to introduce
-prior knowledge on known properties of the unknown target function $f$.
-Rather than relying only on input-output pairs in the learning process,
-we thus narrow our hypothesis set to the biologically most plausible solutions.
-That is, we reduce the search space in a way that
-is compatible with the expected representation of BOLD activity signals.
-
-FORMULA
+\paragraph{Layer combination.}
+Importantly, the optimization problem of the linear autoencoder
+and the low-rank logistic regression
+are linked one two levels. First, their transformation matrices mapping from
+input to the latent space are identical.
+    $V_0 = W_0$
+We thus search for a compression of the 79,941 voxel values into $n$ latent
+components that represent an optimal encoding for both
+rest and task activity data.
+Second, the objectives of the autoencoder and the low-rank
+logistic regression are combined into a same loss function.
 
@@
   \begin{split}
-{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) &= \lambda\frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; W,b)) \hspace{0.2cm} \\
-&+ \hspace{0.2cm} (1-\lambda)\frac{1}{N_{X_{rest}}}\|X_{rest}-\hat X_{rest}(W,b)\|^2
+{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda\frac{1}{N_{X_{task}}} {\mathcal{L_{LR}}}
++ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \ell_1(\theta) + \ell_2(\theta)
+  \label{eq:loss_equ}
 \end{split}
-  \label{eq:loss_equ}
 \end{eqnarray}
 
-mutiple output units
-aCcording to equation \eqref{eq:loss_equ}
+In so doing, we search for the combined set of model parameters
+$\theta=\{V_0,V_1,b_{V_0}, b_{V_1}, b_{V_0}, b_{V_1}\}$
+with respect to the (unsupervised) reconstruction error and the
+(supervised) task classification. For parameter shrinkage, we
+impose a combination of \ell_1 and \ell_1 penalties (i.e., elasticnet)
+on the parameters \theta.
+
+\paragraph{Optimization.}
+The statistical learning problem was approximated in the neuroimaging data
+by computing an updating the parameter derivates of the 
+semi-supervised low-rank logistic regression.
+The required gradients are easily obtained by using the chain rule to
+backpropagate error derivatives first through the decoder network
+and then through the encoder network 
 As solver, we chose \textit{rmsprop} \cite{rmsprop},
@@
 by a gain of $0.004$. The bias parameters were initalized to be zero.
-
 With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
@@
 
-20 components: high bias/low variance
-100 compoentens: low bias/high variance
+\paragraph{Hints.}
+In fact, the constraint by a rest-data autoencoder qualifies as a
+\textit{hint} rather than regularization \cite{abu1994hints}.
+Its purpose is not to prevent overfitting but to introduce
+prior knowledge on known properties of the unknown target function $f$.
+Rather than relying only on input-output pairs in the learning process,
+we thus narrow our hypothesis set to the biologically most plausible solutions.
+That is, we reduce the search space in a way that
+is compatible with the expected representation of BOLD activity signals.
 
-Generating samples from the learned statistical model
-
-minimizing the discrepancy between the orig- inal data and its reconstruction.
-The required gradients are easily obtained by using the chain rule to
-backpropagate error derivatives first through the decoder network
-and then through the encoder network
-
-theanopython
+\paragraph{Implementation.}
 The analyses were performed in Python.
@@
 
+This was evaluated by the prediction
+accuracy on a validation set (20\% of the training data) at each iteration.
+
 all vectors are column vectors
@@
 reduction of n gray-matter voxels to n components
+
+20 components: high bias/low variance
+100 compoentens: low bias/high variance
+
+Generating samples from the learned statistical model
 
@@
 inject prior domain knowledge into the learning process
+
+class weights:
+the right model should have high probability in only some places
 

### 2015-05-23 nips3mm.py

@@ -447,11 +447,11 @@ def dump_comps(masker, compressor, components, threshold=2):
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
 
-n_comps = []
+n_comps = [100]
 # n_comps = [40, 30, 20, 10, 5]
 for n_comp in n_comps:
     # for lambda_param in [0]:
-    for lambda_param in [0.25, 0.50, 0, 0.75, 1]:
+    for lambda_param in [0.1]:
         l1 = 0.1
         l2 = 0.1
         my_title = r'Low-rank LR + AE (combined loss, shared decomp): n_comp=%i L1=%.1f L2=%.1f lambda=%.2f res=3mm spca20RS' % (
@@ -463,30 +463,13 @@ for n_comp in n_comps:
             gain1=0.004,  # empirically determined by CV
             learning_rate = np.float32(0.00001),  # empirically determined by CV,
             max_epochs=500, l1=l1, l2=l2, lambda_param=lambda_param)
-
+        
         estimator.fit(X_rest, X_task, labels)
-
-        # my_title = r'Low-rank LR: n_comp=%i L1=%.1f L2=%.1f res=10mm pca20RS' % (
-        #     n_comp, l1, l2
-        # )
-        # FONTS = 12
-        # plt.figure(facecolor='white', figsize=(8, 6))
-        # plt.plot(np.log(estimator.dbg_ae_cost_), label='cost autoencoder')
-        # plt.plot(estimator.dbg_lr_cost_, label='cost logistic')
-        # plt.plot(estimator.dbg_acc_train_, label='score training set')
-        # plt.plot(estimator.dbg_acc_val_, label='score validation set')
-        # plt.plot(estimator.dbg_acc_other_ds_, label='other-datset acc')
-        # plt.legend(loc='best', fontsize=12)
-        # plt.xlabel('epoch', fontsize=FONTS)
-        # plt.ylabel('misc', fontsize=FONTS)
-        # plt.yticks(np.arange(12), np.arange(12))
-        # plt.grid(True)
-        # plt.title(my_title)
-        # plt.show()
 
[…]

### 2015-05-23 nips3mm_h38.py

@@ -738,7 +738,7 @@ for n_comp in n_comps:  # AE
     plt.show()
     plt.savefig(op.join(WRITE_DIR, 'loss_combined_%icomps.png' % n_comp))
 
-pkgs = glob.glob(RES_NAME + '/*lambda=0.5*dbg_prfs_.npy')
+pkgs = glob.glob(RES_NAME + '/*lambda=0.25*dbg_prfs_.npy')
 for n_comp in n_comps:
     plt.figure()
     for p in pkgs:
@@ -761,7 +761,7 @@ for n_comp in n_comps:
             cur_label += 'RSpca20'
         cur_label += '/'
         cur_label += 'separate decomp.' if 'decomp_separate' in p else 'joint decomp.'
-        for i in np.arange(18):
+        for i in np.arange(38):
             plt.plot(
                 dbg_epochs_,
                 np.array(dbg_prfs_)[:, 0, i],
@@ -777,7 +777,7 @@ for n_comp in n_comps:
     plt.savefig(op.join(WRITE_DIR, 'prec_inds_%icomps.png' % n_comp))
 
 # in-dataset recall at lambda=0.5
-pkgs = glob.glob(RES_NAME + '/*lambda=0.5*dbg_prfs_.npy')
+pkgs = glob.glob(RES_NAME + '/*lambda=0.25*dbg_prfs_.npy')
 for n_comp in n_comps:
     plt.figure()
     for p in pkgs:
@@ -801,7 +801,7 @@ for n_comp in n_comps:
             cur_label += 'RSpca20'
         cur_label += '/'
         cur_label += 'separate decomp.' if 'decomp_separate' in p else 'joint decomp.'
-        for i in np.arange(18):
+        for i in np.arange(38):
             plt.plot(
                 dbg_epochs_,
                 np.array(dbg_prfs_)[:, 1, i],
@@ -817,7 +817,7 @@ for n_comp in n_comps:
     plt.savefig(op.join(WRITE_DIR, 'rec_inds_%icomps.png' % n_comp))
 
 # in-dataset f1 at lambda=0.5
-pkgs = glob.glob(RES_NAME + '/*lambda=0.5*dbg_prfs_.npy')
+pkgs = glob.glob(RES_NAME + '/*lambda=0.25*dbg_prfs_.npy')
 for n_comp in n_comps:
     plt.figure()
     for p in pkgs:
[…]

### 2015-05-23 nips3mm_recovery.py

@@ -2,6 +2,8 @@
 HCP: semi-supervised network decomposition by low-rank logistic regression
 """
 """
+preliminary conclusion:
+- abc
 """
 
 print __doc__
@@ -26,45 +28,11 @@ from nilearn.image import concat_imgs
 import joblib
 import time
 
-RES_NAME = 'nips3mm_vanilla'
-WRITE_DIR = op.join(os.getcwd(), RES_NAME)
-if not op.exists(WRITE_DIR):
-    os.mkdir(WRITE_DIR)
+LR_AE_DIR = 'nips3mm'
+LR_DIR = 'nips3mm_vanilla'
 
 ##############################################################################
-# load+preprocess data
-##############################################################################
-
-mask_img = 'grey10_icbm_3mm_bin.nii.gz'
-nifti_masker = NiftiMasker(mask_img=mask_img, smoothing_fwhm=False,
-                           standardize=False)
-nifti_masker.fit()
-mask_nvox = nifti_masker.mask_img_.get_data().sum()
-
-print('Loading data...')
-
-# ARCHI task
-X_task, labels = joblib.load('preload_HT_3mm')
-
-labels = np.int32(labels)
-
-# contrasts are IN ORDER -> shuffle!
-new_inds = np.arange(0, X_task.shape[0])
-np.random.shuffle(new_inds)
-X_task = X_task[new_inds]
-labels = labels[new_inds]
-# subs = subs[new_inds]
-
-X_task = StandardScaler().fit_transform(X_task)
-
-# ARCHI task
-AT_niis, AT_labels, AT_subs = joblib.load('preload_AT_3mm')
-AT_X = nifti_masker.transform(AT_niis)
-AT_X = StandardScaler().fit_transform(AT_X)
-print('done :)')
-
-##############################################################################
-# define computation graph
+# abc
 ##############################################################################
 
 class SSEncoder(BaseEstimator):
@@ -267,125 +235,117 @@ class SSEncoder(BaseEstimator):
         else:
             return acc
 
-
 ##############################################################################
-# plot figures
+# abc
[…]

### 2015-05-24 paper_nips/paper.tex
starting results

@@
 %
-Existing neuroimaging methods typically perform either discovery of
-neurobiological structure or evaluation of explicit hypotheses on mental tasks.
+Existing neuroimaging methods typically perform either discovery of unknown
+neural structure or testing of neural structure associated with mental tasks.
 %
-Modelling mental tasks however hinges
+However, modelling the neural correlates of mental tasks hinges
 on the pertinence of the assumed neurobiological structure.
@@
 We show that this approach yields more accurate and more interpretable
-neural models of psychological tasks in a reference neuroimaging dataset.
+neural models of psychological tasks in a reference dataset.
 %
@@
 Methods used in neuroimaging research can be grouped by discovering
-neurobiological structure or revealing the neural correlates associated
+neurobiological structure or assessing the neural correlates associated
 with mental tasks.
-To discover coherent distributions of activation structure across time,
-independent component analysis (ICA; \cite{beckmann2005}) is often used
+To discover spatial distributions of activation structure across time,
+independent component analysis (ICA) \cite{beckmann2005} is often used
 to decompose the BOLD (blood-oxygen level-dependent) signals into the
-important modes of variation.
+main modes of variation.
 The ensuing spatial activation patterns are believed to represent
 brain networks of
-functionally interacting brain regions.
-Similarly, sparse principal component analysis (SPCA; \cite{varoqu2011})
+functionally interacting brain regions \cite{smith2009}.
+Similarly, sparse principal component analysis (SPCA) \cite{varoqu2011}
 has been used to
-separate brain activity signals into parsimonious network components.
+separate BOLD signals into parsimonious network components.
 Thus extracted brain networks have been shown to be
 manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
-Their fundamental role in brain organization is
+Their fundamental organizational role is further
 attested by continued covariation during sleep and anesthesia.
@@
 unlabeled "resting-state" data. These capture brain dynamics
-during ongoing random thought without controlled environmental input.
-The biggest fraction of the BOLD signals are known
-not to correlate with a particular behavior, stimulus, or experimental task.
+during ongoing random thought without controlled environmental stimulation.
+A large proportion of the BOLD signal is known
+not to correlate with a particular behavior, stimulus, or experimental task
+\cite{fox07}.
 
@@
 the neural correlates underlying mental tasks,
-the general linear model (GLM; \cite{friston94}) is the dominant approach.
+the general linear model (GLM) is the dominant approach \cite{friston94}.
 The contribution of
@@
 according to a design matrix of experimental tasks.
-Alternatively, psychophysiological interactions (PPI; \cite{friston97}),
+Alternatively, psychophysiological interactions (PPI)
 elucidate the functional interactions between voxels as a function
-of experimental tasks.
-Dynamic causal modeling (DCM; \cite{stephan04}), in turn, quantifies directed,
+of experimental tasks \cite{friston97}.
+Dynamic causal modeling (DCM), in turn, quantifies directed,
 task-driven influences between regions
 by treating the brain as a nonlinear dynamic system with unknown
-neuronal states. As a last example, always more neuroimaging studies model
+neuronal states \cite{stephan04}.
+As a last example, always more neuroimaging studies model
 experimental tasks by training classification algorithms on brain signals
@@
 First, the mentioned supervised neuroimaging analyses operate
-without exception in a single-voxel space. This ignores the fact that the BOLD
-signal exhibits coherent spatial activation patterns.
+without exception in a voxel space. This ignores the fact that the BOLD
+signal exhibits distributed spatial activation patterns.
 Second, existing neuroimaging analyses do not acknowledge the fact that the
-task-induced changes of the BOLD signal amount to less than 5\%
+task-induced changes of the BOLD signal typically amount to less than 5\%
 of baseline activity \cite{fox07}. They do thus not exploit the high similarity
-of BOLD dynamics in the human brain at rest and during experimental tasks.
-Indeed, very similar brain networks were observed when applying ICA
-separately on rest and task data \cite{smith2009}.
+of BOLD dynamics in the human brain at rest and during experimental tasks,
+as observed using ICA \cite{smith2009}.
 
-Both biological properties can be conjointly exploited in a 
+Both these biological properties can be conjointly exploited in a 
 semi-supervised (i.e., use rest and task data)
@@
 %
-The integration of brain-network discovery in a 
-supervised classification goal should identify the
-neurobiological structure that allows for the best predictive models.
+The integration of brain-network discovery into a 
+supervised classification framework should identify the most relevant
+neurobiological structure for the prediction problem at hand.
 %
 Autoencoders suggest themselves because they can emulate
-variants of most nonsupervised learning algorithms,
-including PCA, SPCA, and ICA.
+variants of most unsupervised learning algorithms,
+including PCA, SPCA, and ICA \cite{hinton06}.
 Autoencoders
@@
 local and global representations
-by improving reconstruction from them \cite{hinton06}.
+by improving reconstruction from them.
 %
@@
 in case of one linear hidden layer and a squared error loss
+%
+\begin{wrapfigure}{r}{0.40\textwidth}
+  \centering
+  %\captionsetup{labelformat=nonumber}
+    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
+  \caption {\textbf{Architecture}
+  }
+\end{wrapfigure}
+%
 \cite{baldi1989neural}.
@@
 with unique global minimum.
-Autoencoders behave like a SPCA if shrinkage terms are added for the
-matrix weights in the optimization objective.
-In turn, they behave like an ICA in case of a nonlinear convex
+Autoencoders behave like a SPCA if shrinkage terms are added to the
+model weights in the optimization objective.
+Moreover, they behave characteristics of an ICA in case of a nonlinear convex
 function of the first-layer activation and tied weights \cite{le2011ica}.
@@
 under mild conditions.
-In this way, autoencoders can flexibly project the neuroimaging data
-onto the axes of main variation and thus
+Hence, autoencoders can flexibly project the neuroimaging data
+onto the main axes of variation and thus
 reverse-engineer properties of the data-generating
 neural processes \cite{olshausen96}.
-
 In the present investigation,
-an autoencoder will be fed by (unlabeled) rest data and
+a linear autoencoder will be fed by (unlabeled) rest data and
 integrated as bottleneck
 into a low-rank logistic regression fed by (labeled) task data.
-Using the chain rule in back-propagation, we can then
-solve the unsupervised data representation and a supervised classification
+Using the chain rule for grdient descent, we can then
+solve the unsupervised data representation and the supervised classification
 in an identical statistical learning problem.
-%
-From the perspective of dictionary learning, the first layer can be
-viewed as a learned set of basis functions
+From the perspective of dictionary learning, the first layer represents
+the discovered set of basis functions
 whose linear combinations are learned
-in the second layer \cite{olshausen96}.
+by the second layer \cite{olshausen96}.
 %
@@
 %
-Theoretically, a big reduction of the model variance is expected by
-regularization using resting-state autoencoder
-to put probability mass on the most neurobiologically
-valid members of the model space.
+Theoretically, a reduced model variance should be entailed by
+resting-state autoencoders that
+put probability mass on the most neurobiologically
+valid members in the model space.
 %
 The generalization performance should consequently be improved due to 
-much reduced Vapnik-Chervonenkis dimensions of the classification estimator.
+reduced Vapnik-Chervonenkis dimensions of the trained estimator.
 %
@@
 
-\begin{wrapfigure}{r}{0.40\textwidth}
-  \begin{center}
-    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
-  \end{center}
-  \caption {\textbf{Architecture}
-  }
-\end{wrapfigure}
-
-%
 \section{Methods}
@@
 \paragraph{Data.}
-Unsupervised projection models into a lower-dimensional space and
-supervised logistic-regression models are learned in concert
-in a same parameter gradient descent.
-Brain
-network decompositions are thus exposed that explain task-descriminative spatial
-patterns of neural activity.
-
 As the currently biggest openly-accessible reference dataset,
-we chose the Human Connectome Project (HCP) for the present analyses
+we chose the Human Connectome Project (HCP) resources
 \cite{barch2013}.
@@
 in-plane FOV=$280\times180$mm, 72 slices, 2.0mm isotropic voxels).
-The “minimally preprocessed” pipeline includes
+The ``minimally preprocessed'' pipeline includes
 gradient unwarping, motion correction, fieldmap-based EPI distortion
 correction, brain-boundary-based registration of EPI to structural
-T1-weighted scans, non-linear (FNIRT) registration into MNI space,
+T1-weighted scans, nonlinear (FNIRT) registration into MNI space,
 and grand-mean intensity normalization. Activation maps were spatially
@@
 probability).
-All supservised analyses were based on labeled HCP task maps of
-79,941 voxels representing Z values in gray matter.
+The supservised analyses were based on labeled HCP task maps with
+79,941 voxels of interest representing z-values in gray matter.
 
@@
 These reflect brain activity in the absence of controlled thought.
-In line with the goal of the present study, acquisition of the data was
+In line with the goal of the present study, acquisition of these data was
 specifically aimed at the study of task-rest correspondence.
-From each included participant, we included two
+From each participant, we included two
 time-series for left and right phase encoding
@@
 To this end, the HCP-derived network decompositions were used as preliminary
-step in the classifcation problem of another large-cohort dataset.
+step in the classifcation problem of another large cohort.
 The ARCHI dataset \cite{pinel07} provides activity maps from
@@
 
+The labeled and unlabeled data were fed into a linear statistical model
+composed of autoencoder and low-rank logistic regression.
+
 \paragraph{Unsupervised layer.}
-The labeled and unlabeled data were fed into a linear statistical model.
-The model is composed of an autoencoder and a low-rank logistic regression.
-On the one hand, the affine autoencoder takes the input 
+The affine autoencoder takes the input 
 $x$ and projects it into coordinates of a latent representation $z$
@@
 
-where $x \in \mathbb{R}^{d}$ denotes the vector of $d=79,941$ voxel from each
+where $x \in \mathbb{R}^{d}$ denotes the vector of $d=79,941$
+voxel values from each
 rest map, $z \in \mathbb{R}^{n}$ is the $n$ hidden dimensions (i.e.,
-distributed macroscopical activity patterns), and 
+distributed neural activity patterns), and 
 $x' \in \mathbb{R}^{d}$ is the reconstruction vector of the original activity map
@@
 transforms
-from input space into the hidden space (i.e., encoder),
+from input space into the hidden space (encoder),
 ${W_1}$ is the weight matirx for back-projection from the hidden space to the
-output space (i.e., decoder), ${b_{W_0}}$ and ${b_{W_1}}$ are bias vectors.
-Note that ${W_0}$ and ${W_1}$ are tied
+output space (decoder), ${b_{W_0}}$ and ${b_{W_1}}$ are bias vectors.
+Note that ${W_0}$ and ${W_1}$ are tied,
 $W_0$ = $W_1^T$.
 The optimal model parameters ${W_0}, {b_{W_0}}, {b_{W_1}}$ are found by
-minimizing the reconstruction difference by squared error
+minimizing the reconstruction difference according to squared error
 
 \begin{eqnarray}
-  {\mathcal{L_{AE}}}(x, x') = || \x - \x' ||^2
+  {\mathcal{L_{AE}}}(x, x') = || x - x' ||^2
 \end{eqnarray}
@@
 input and the learned representation.
-Non-linearities were not applied on the
-transformation results.
+Nonlinearities were not applied on the
+activations in the first layer.
 
 \paragraph{Supervised layer.}
-On the other hand, such lossy compression by a low-dimensional bottleneck
+Lossy compression by a low-dimensional bottleneck
 is also imposed by the first layer of the low-rank
-logistic regression.
+multinomial logistic regression.
+It gives the probability of an input $x$ to belong
+to a class $i$
 
@@
 
+where the matrix $V_0 \in \mathbb{R}^{dxn}$ transforms the input$x \in \mathbb{R}^{d}$
+into $n$ latent components
+and the matrix $V_1 \in \mathbb{R}^{nxc}$ projects the latent components
+onto hyperplanes that reflect $c$ class probabilities.
+$b_{V_0}$ and $b_{V_1}$ are corresponding
+bias vectors.
+The loss function is given by
+
 \begin{eqnarray}
   \begin{split}
-    {\mathcal{L_{LR}}}(x, y) = \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1})
+    {\mathcal{L_{LR}}}(x, y) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1})
 \end{split}
@@
 \end{eqnarray}
-
-\begin{wrapfigure}{r}{0.40\textwidth}
-  \begin{center}
-    \includegraphics[width=0.40\textwidth]{figures/figure3.png}
-  \end{center}
-  \caption {\textbf{Equilibrium between autoencoder and low-rank logistic regression}
-  One learned decomposition component (out of 20) between the only-autoencoder
-  (\textit{uppermost panel}) and only-logistic-regression
-  (\textit{lowermost panel}) scenarios.
-  The congruent structure (\textit{red, middle column})
-  in the
-  posterior cingulate cortex, posterior mid-cingulate cortex, and medial
-  prefrontal cortex that is hence important in decompositions of the rest and
-  task neuroimaging data. With increasing weight on the supervised learning,
-  non-congruent structure emerges in the early (\textit{blue}) and
-  lateral (\textit{red}) visual cortex
-  (\textit{right column}). Matrix weights were z-scored.
-  }
-\end{wrapfigure}
 
@@
 are linked one two levels. First, their transformation matrices mapping from
-input to the latent space are identical.
-    $V_0 = W_0$
+input to the latent space are identical
+\begin{eqnarray}
+  V_0 = W_0
+\end{eqnarray}
 We thus search for a compression of the 79,941 voxel values into $n$ latent
-components that represent an optimal encoding for both
+components that represent an optimal latent code for both
 rest and task activity data.
 Second, the objectives of the autoencoder and the low-rank
-logistic regression are combined into a same loss function.
+logistic regression are combined into a common loss function.
 
@@
   \begin{split}
-{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda\frac{1}{N_{X_{task}}} {\mathcal{L_{LR}}}
+{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
 + (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \ell_1(\theta) + \ell_2(\theta)
@@
 
-In so doing, we search for the combined set of model parameters
-$\theta=\{V_0,V_1,b_{V_0}, b_{V_1}, b_{V_0}, b_{V_1}\}$
+In so doing, we search for the combined model parameters
+$\theta=\{V_0,V_1,b_{V_0}, b_{V_1}, b_{W_0}, b_{W_1}\}$
 with respect to the (unsupervised) reconstruction error and the
-(supervised) task classification. For parameter shrinkage, we
-impose a combination of \ell_1 and \ell_1 penalties (i.e., elasticnet)
-on the parameters \theta.
+(supervised) task classification.
+${\mathcal{L_{AE}}}$ is devided by ${N_{X_{rest}}}$ to equilibrate both
+loss terms to the same order of magnitude.
+For parameter shrinkage, we
+impose a combination of $\ell_1$ and $\ell_2$ penalties (i.e., elasticnet)
+on the parameters $\theta$.
 
 \paragraph{Optimization.}
-The statistical learning problem was approximated in the neuroimaging data
-by computing an updating the parameter derivates of the 
+The common objective was approximated in the neuroimaging data
+by updating parameters by derivates of the 
 semi-supervised low-rank logistic regression.
@@
 backpropagate error derivatives first through the decoder network
-and then through the encoder network 
+and then through the encoder network.
 As solver, we chose \textit{rmsprop} \cite{rmsprop},
@@
 The matrix parameters were initalized by Gaussian random values multiplied
-by a gain of $0.004$. The bias parameters were initalized to be zero.
+by a gain of $0.004$. The bias parameters were initalized to zero.
 With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
@@
 The hyper-parameter $\alpha$ was set to $0.00001$ by prior manuel
-cross-validation and $\epsilon$ was set to $1^{-6}$.
+cross-validation and $\epsilon$ was set to $10^{-6}$.
 %
@@
 prior knowledge on known properties of the unknown target function $f$.
-Rather than relying only on input-output pairs in the learning process,
+Rather than only relying on input-output pairs in the learning process,
 we thus narrow our hypothesis set to the biologically most plausible solutions.
@@
 and
-\textit{theano} for automatic
+\textit{theano} for automatic, numerically stable
 differentiation of symbolic computation graphs
@@
 accessible online for reproducibility and reuse
-\url{http://github.com/banilo/nips2015}.
+\url{http://github.com/anonymous/anonymous}.
+% url{http://github.com/banilo/nips2015}.
+
+\bigskip
+
+\begin{figure}
+\begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/figure_weights.jpg}
+\end{centering}
+\caption{\textbf{Comparing the interpretability of classification weights}
+The voxel predictors corresponding to 5 (of 18) psychological tasks (\textit{rows})
+from the HCP dataset \cite{barch2013}.
+\textit{Left column:} ordinary logistic regression (based on same
+implementation but without bottleneck layer or autoencoder),
+\textit{middle column:} semi-supervised low-rank logistic regression
+(latent components $n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
+\textit{right column:} voxel-wise average of whole-brain activity
+across the first-level activity maps from 498 participants
+(tresholded at 0.25).
+All values z-scored.
+}
+\end{figure}
+
+\begin{figure}
+\begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/support_bars.png}
+\end{centering}
+\caption{\textbf{Relation between support recovery and out-of-dataset performance}
+\textit{Upper row:} the fit-for-purpose of the classification models
+is quantified by linear correlation between the
+z-scored average task image and the z-scored weights from
+ordinary (\texit{red}) and present (\textit{blue colors}) logistic
+regression ($\lambda=0.25, 0.50, 0.75, 1.00$).
+\textit{Lower row:} ordinary logistic regression is applied to the
+ARCHI dataset after projection into the latent dimensions learned from
+the HCP task data.
+}
+\end{figure}
+
+\bigskip
 
@@
 
-This was evaluated by the prediction
-accuracy on a validation set (20\% of the training data) at each iteration.
-
-all vectors are column vectors
-
-affine encoder and decoder
-
-Low-rank regression outperformed serial ICA/SparsePCA and logistic regression.
-
-reduction of n gray-matter voxels to n components
+\paragraph{Classification performance}
+\begin{itemize}
+  \item{This was evaluated by the prediction
+accuracy on a validation set (20\% of the training data) at each iteration.}
+\end{itemize}
 
@@
 
-Generating samples from the learned statistical model
+\paragraph{Support recovery}
+\begin{itemize}
+  \item{fit for purpose; the right model should have
+  high probability in the right places}
+  \item{comparing to ordinary logistic regression, low-rank logreg yielded class weights 
+  that were much more similar to feature of the respective training examples,
+  for any choice of $\lambda$}
+  \item{the higher $\lambda$, the better support recovery}
+  \item{conversely, the higher lambda (i.e., bigger emphasis on the
+  resting-state autoencoder), the more useful the low-dimensional transformations
+  for classification in an independent dataset (ARCHI) as measured by f1 score
+  for higher (i.e., 50 and 100) but not lower (i.e., 20) numbers
+  of latent components}
+\end{itemize}
 
-Modifications of the mode that di not improve the genralization performacne:
-dropo-out, input corruption, addining non-linearities (sigmoid, tanh),
+\paragraph{Anecdotal observations}
+\begin{itemize}
+  \item{Modifications of the mode that did not improve the genralization performacne:
+  dropo-out, input corruption (p=0.1/0.3/0.5), relus}
+  \item{no penalty or only L1 or L2 penalty generally yielded worse
+  generalization performance}
+  \item{low L1/L2 values (0.1) yielded better results
+  than higher values (0.3, 0.5)}
+  \item{``pretraining'' of the bottleneck by settings weights to
+  preliminary PCA, SPCA or ICA weights, neither for autoencoder pretraining }
+  \item{present approach better than first first PCA, SPCA or ICA and
+  subsequently Logreg on loadings}
+  \item{introducing a
+  nonlinearity (sigmoid, tanh, softplus) into the system
+  did not improve predictive accuracy but
+  elastic did -> most useful decomposition has characteristics of a SPCA
+  and not PCA or ICA}
+  \item{introduced an addition layer with many units (50/100) did not
+  improve generalization}
+\end{itemize}
 
-introducing a
-nonlinearity (sigmoid, tanh) into the system did not improve predictive accuracy but
-elastic did -> most useful decomposition has characteristics of a SPCA
-and not PCA or ICA
-
-outperforms plain vanilla LR
-inject prior domain knowledge into the learning process
-
-class weights:
-the right model should have high probability in only some places
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.40\textwidth]{figures/figure3.png}
+  \end{center}
+  \caption {\textbf{
+  Equilibrium between autoencoder and low-rank logistic regression}
+  One learned decomposition component (out of 20) between the only-autoencoder
+  (\textit{uppermost panel}) and only-logistic-regression
+  (\textit{lowermost panel}) scenarios.
+  The congruent structure (\textit{red, middle column})
+  in the
+  posterior cingulate cortex, posterior mid-cingulate cortex, and medial
+  prefrontal cortex appears to be important in decompositions of the rest and
+  task neuroimaging data. With increasing weight on the supervised learning,
+  non-congruent structure emerges in the early (\textit{blue}) and
+  lateral (\textit{red}) visual cortex
+  (\textit{right column}). Matrix weights were z-scored.
+  }
+\end{figure}
 
@@
 There is an increasing occurrence of high-dimensional problems in the
-neuroimaging domain. This calls for new statistical learning algorithms that
-behave well in large-cohort settings. Ideally, they should acknowledge
+neuroimaging domain. Yet, there is also much uncertainty about the
+most pertinent representation
+of neural activation information.
+This calls for new statistical learning algorithms that
+behave well in large cohorts. Ideally, they should acknowledge
 and exploit existing widely-accepted neuroscientific knowledge.
@@
 in a neurobiologically valid and interpretable fashion.
+%
+Unsupervised projection models into a lower-dimensional space and
+supervised logistic-regression models are learned in concert
+in a same parameter gradient descent.
+Brain
+network decompositions are thus exposed that explain task-descriminative spatial
+patterns of neural activity.
+
+classifier that operates in a (sparse) network space, rather than in a voxel space.
+domain-specifc classification algorithms
 
@@
 
-respect structure in the fMRI data
+-respect structure in the fMRI data
 
-- if linearity, then would be closer to the notion of 1-hidden layer neural
-network rather than low-rank logistic regression
-
-We hope that these results stimulate the development of
+- We hope that these results stimulate the development of
 even more powerful semi-supervised classification methods
 
-improve computational tractability, prediction accuracy, and interpretability
+learn optimal parameters
+
+- improve computational tractability, prediction accuracy, and interpretability
 neuroimaging datasets
-does it produce testable predictions?
+- does it produce testable predictions for targeted experimental investigation?
 -> we can test predictive value of network-network architectures across
@@
 every-day mind-wandering and task-focused brain states.
-
-classifier that operates in a (sparse) network space, rather than in a voxel space.
-domain-specifc classification algorithms
 
@@
 
-There is much uncertainty about the most pertinent representation
-of neural activation information
 put probability mas shwere we expeted neurobiological strucuture
@@
 canonical set of brain networks
+
+new baseline
 

### 2015-05-24 nips3mm.py

@@ -1058,10 +1058,10 @@ for p in pkgs:
     plt.savefig(op.join(WRITE_DIR, new_fname))
 
 # out-of-dataset f1 score summary plots
-for n_comp in [5, 20, 50, 100]:
+for n_comp in [20, 50, 100]:
     f1_mean_per_lambda = list()
     f1_std_per_lambda = list()
-    lambs = [0.1, 0.25, 0.5, 0.75, 1.0]
+    lambs = [0.25, 0.5, 0.75, 1.0]
     for target_lambda in lambs:
         pkgs = glob.glob(RES_NAME + '/*n_comp=%i*lambda=%.2f*dbg_prfs_other_ds_.npy' %
             (n_comp, target_lambda))
@@ -1077,9 +1077,10 @@ for n_comp in [5, 20, 50, 100]:
     f1_std_per_lambda = np.array(f1_std_per_lambda)
 
     plt.figure()
-    ind = np.arange(5)
+    ind = np.arange(4)
     width = 1.
-    colors = [(7., 116., 242.),
+    colors = [#(7., 116., 242.), #(7., 176., 242.)
+        #(7., 136., 217.), (7., 40., 164.), (1., 4., 64.)]
         (7., 176., 242.), (7., 136., 217.), (7., 40., 164.), (1., 4., 64.)]
     my_colors = [(x/256, y/256, z/256) for x, y, z in colors]
     plt.bar(ind, f1_mean_per_lambda, yerr=f1_std_per_lambda,
@@ -1090,7 +1091,8 @@ for n_comp in [5, 20, 50, 100]:
     tick_strs = [u'low-rank $\lambda=%.2f$' % val for val in lambs]
     plt.xticks(ind + width / 2., tick_strs, rotation=320)
     plt.ylim(.5, 1.0)
-    plt.yticks(np.linspace(0.5, 1., 6), np.linspace(0.5, 1., 6))
+    plt.grid(True)
+    plt.yticks(np.linspace(0.5, 1., 11), np.linspace(0.5, 1., 11))
     plt.tight_layout()
     out_path2 = op.join(WRITE_DIR, 'f1_bars_comp=%i.png' % n_comp)
     plt.savefig(out_path2)

### 2015-05-24 nips3mm_recovery.py

@@ -280,10 +280,10 @@ lr_supp_z = zscore(lr_supp, axis=1)
 
 # get LR/AE weights
 WRITE_DIR = 'nips3mm_recovery'
-lambs = [0.1, 0.25, 0.5, 0.75, 1]
+lambs = [0.25, 0.5, 0.75, 1]
 import re
 from scipy.stats import pearsonr
-for n_comp in [5, 20, 50, 100]:
+for n_comp in [20, 50, 100]:
     corr_means_lr = np.zeros((len(lambs), 18))
     corr_means_lr_ae = np.zeros((len(lambs), 18))
     for ilamb, lamb in enumerate(lambs):
@@ -315,6 +315,7 @@ for n_comp in [5, 20, 50, 100]:
                 print('r/lr: %.4f' % r2)
                 corr_means_lr[ilamb, i] = r2
 
+    # boxplot
     plt.figure()
     corrs = np.vstack((corr_means_lr[0, :], corr_means_lr_ae))
     plt.boxplot(corrs.T)
@@ -330,10 +331,11 @@ for n_comp in [5, 20, 50, 100]:
     out_path = op.join(WRITE_DIR, 'supp_recov_comp=%i.png' % n_comp)
     plt.savefig(out_path)
 
+    # barplot
     plt.figure()
-    ind = np.arange(6)
+    ind = np.arange(5)
     width = 1.
-    colors = [(242., 62., 22.), (7., 196., 255.),
+    colors = [(242., 62., 22.), #(7., 196., 255.),
         (7., 176., 242.), (7., 136., 217.), (7., 40., 164.), (1., 4., 64.)]
     my_colors = [(x/256, y/256, z/256) for x, y, z in colors]
     plt.bar(ind, np.mean(corrs, axis=1), yerr=np.std(corrs, axis=1),

### 2015-05-25 paper_nips/paper.tex
first draft

@@
 model weights in the optimization objective.
-Moreover, they behave characteristics of an ICA in case of a nonlinear convex
-function of the first-layer activation and tied weights \cite{le2011ica}.
+Moreover, they have the characteristics of an ICA in case of tied weights
+and feeding the first-layer activation into any nonlinear convex
+function \cite{le2011ica}.
 These authors further demonstrated that ICA, sparse autoencoders, and 
@@
 neural processes \cite{olshausen96}.
+
 In the present investigation,
@@
 into a low-rank logistic regression fed by (labeled) task data.
-Using the chain rule for grdient descent, we can then
+Using the chain rule for gradient descent, we can then
 solve the unsupervised data representation and the supervised classification
@@
 %
-Theoretically, a reduced model variance should be entailed by
+Theoretically, a reduction in model variance should be achieved by
 resting-state autoencoders that
 put probability mass on the most neurobiologically
-valid members in the model space.
+valid models in the hypothesis set.
 %
@@
 To this end, the HCP-derived network decompositions were used as preliminary
-step in the classifcation problem of another large cohort.
+step in the classifcation problem of another large sample.
 The ARCHI dataset \cite{pinel07} provides activity maps from
@@
 
-\paragraph{Unsupervised layer.}
+\paragraph{Linear autoencoder.}
 The affine autoencoder takes the input 
@@
 from input space into the hidden space (encoder),
-${W_1}$ is the weight matirx for back-projection from the hidden space to the
-output space (decoder), ${b_{W_0}}$ and ${b_{W_1}}$ are bias vectors.
+${W_1}$ is the weight matrix for back-projection from the hidden variables to the
+output space (decoder).
+${b_{W_0}}$ and ${b_{W_1}}$ are bias vectors.
 Note that ${W_0}$ and ${W_1}$ are tied,
@@
 The optimal model parameters ${W_0}, {b_{W_0}}, {b_{W_1}}$ are found by
-minimizing the reconstruction difference according to squared error
+minimizing the imperfection of reconstruction according to squared error
 
@@
 
-\paragraph{Supervised layer.}
+\paragraph{Reduced-rank logistic regression.}
 Lossy compression by a low-dimensional bottleneck
@@
 It gives the probability of an input $x$ to belong
-to a class $i$
+to a class $i \in \mathbb{N}$
 
@@
 
-where the matrix $V_0 \in \mathbb{R}^{dxn}$ transforms the input$x \in \mathbb{R}^{d}$
+where the matrix $V_0 \in \mathbb{R}^{dxn}$ transforms the input $x \in \mathbb{R}^{d}$
 into $n$ latent components
@@
 Second, the objectives of the autoencoder and the low-rank
-logistic regression are combined into a common loss function.
+logistic regression are interpolated in the common loss function
 
@@
   \begin{split}
-{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
+{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = (\lambda) {\mathcal{L_{LR}}}
 + (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \ell_1(\theta) + \ell_2(\theta)
@@
 For parameter shrinkage, we
-impose a combination of $\ell_1$ and $\ell_2$ penalties (i.e., elasticnet)
-on the parameters $\theta$.
+impose a combination of $\ell_1$ and $\ell_2$ penalties
+$\forall p \in \theta$ (i.e., elasticnet).
 
@@
 The required gradients are easily obtained by using the chain rule to
-backpropagate error derivatives first through the decoder network
-and then through the encoder network.
+backpropagate error derivatives through the linear network.
 As solver, we chose \textit{rmsprop} \cite{rmsprop},
@@
 scaled gradients from a running average.
-Gradient normalization by RMSprop
+Gradient normalization by rmsprop
 is known to effectively exploit curvature information.
@@
 $0.9$ to deemphasize the magnitude of the gradient.
-Further $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
-The hyper-parameter $\alpha$ was set to $0.00001$ by prior manuel
+Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
+The hyper-parameter $\alpha$ was set to $0.00001$ by prior manual
 cross-validation and $\epsilon$ was set to $10^{-6}$.
@@
 In fact, the constraint by a rest-data autoencoder qualifies as a
-\textit{hint} rather than regularization \cite{abu1994hints}.
+\textit{hint}
+rather than regularization in a strict sense \cite{abu1994hints}.
 Its purpose is not to prevent overfitting but to introduce
@@
 \begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/accuracies.png}
+\end{centering}
+\caption{\textbf{Generalization performance}
+Depicts in-sample (apply current model on training set, \textit{left column}),
+out-of-sample (current model on test set, \textit{middle column}),
+and out-of-dataset (compress other dataset using the
+learned decomposition for classification, \textit{right column}) accuracy.
+Curves for different choices of $\lambda$ as well as for ordinary
+logistic regression (\textit{yellowish})
+across iterations over the HCP task data (epochs).
+}
+\end{figure}
+
+\begin{figure}
+\begin{centering}
 \includegraphics[width=1.00\textwidth]{figures/figure_weights.jpg}
@@
 across the first-level activity maps from 498 participants
-(tresholded at 0.25).
+(thresholded at 0.25).
 All values z-scored.
@@
 z-scored average task image and the z-scored weights from
-ordinary (\texit{red}) and present (\textit{blue colors}) logistic
+ordinary (\textit{red}) and present (\textit{blue colors}) logistic
 regression ($\lambda=0.25, 0.50, 0.75, 1.00$).
@@
 
-\bigskip
-
 \section{Experimental Results}
+\paragraph{General}
+\begin{itemize}
+\item{20 components: high bias/low variance
+100 compoentens: low bias/high variance}
+\end{itemize}
 
@@
   \item{This was evaluated by the prediction
-accuracy on a validation set (20\% of the training data) at each iteration.}
+accuracy on a validation set (20\% of the training data) at each iteration
+over the ARCHI task data}
+  \item{the less lambda, the faster the convergence, with small differences
+  though; smaller differences with larger number of components}
+  \item{both 20 and 100 hidden components achieve 97\% out-of-sample
+  accuracy}
+  \item{ordinary logistic regression converges faster but with stagnation at
+  92\% out-of-sample accuracy}
+  \item{classification at chance 6\% in case of lambda==0 because 
+  only the unsupervised layer of the estimator is optimized}
+  \item{important: the thus only resting-state-trained decomposition achieves
+  the same usefulness in the ARCHI classification than the
+  decomposition derived in semi-supervised fashion}
 \end{itemize}
-
-20 components: high bias/low variance
-100 compoentens: low bias/high variance
 
@@
 \begin{itemize}
-  \item{Modifications of the mode that did not improve the genralization performacne:
-  dropo-out, input corruption (p=0.1/0.3/0.5), relus}
+  \item{Modifications of the modek that did not improve the generalization performance:
+  drop-out, input corruption (p=0.1/0.3/0.5), RELUs}
   \item{no penalty or only L1 or L2 penalty generally yielded worse
   generalization performance}
-  \item{low L1/L2 values (0.1) yielded better results
+  \item{low L1/L2 penalties (0.1) yielded better results
   than higher values (0.3, 0.5)}
   \item{``pretraining'' of the bottleneck by settings weights to
-  preliminary PCA, SPCA or ICA weights, neither for autoencoder pretraining }
-  \item{present approach better than first first PCA, SPCA or ICA and
-  subsequently Logreg on loadings}
+  preliminary PCA, SPCA or ICA weights did not give better accuracies,
+  neither for autoencoder pretraining }
+  \item{present approach better than first PCA, SPCA or ICA and
+  subsequently Logreg on loadings, even if each on the whole data}
   \item{introducing a
@@
 There is an increasing occurrence of high-dimensional problems in the
-neuroimaging domain. Yet, there is also much uncertainty about the
+neuroimaging domain. Yet, there is also considerable uncertainty about the
 most pertinent representation
@@
 This calls for new statistical learning algorithms that
-behave well in large cohorts. Ideally, they should acknowledge
-and exploit existing widely-accepted neuroscientific knowledge.
-In the present work, we propose such an estimator that learns
+behave well in large samples. Ideally, they should acknowledge
+and exploit widely-accepted neuroscientific knowledge.
+We here propose such an estimator that learns
 dimensionality reduction
-in a neurobiologically valid and interpretable fashion.
-%
-Unsupervised projection models into a lower-dimensional space and
-supervised logistic-regression models are learned in concert
-in a same parameter gradient descent.
-Brain
-network decompositions are thus exposed that explain task-descriminative spatial
-patterns of neural activity.
+in a neurobiological and interpretable fashion.
 
-classifier that operates in a (sparse) network space, rather than in a voxel space.
-domain-specifc classification algorithms
+Using the flexibility of autoencoders, we
+learn the optimal decomposition from high-dimensional
+voxel brain space into the most important activation patterns for a
+statistical learning question of interest.
+The higher generalization accuracy and support recovery, comparing to
+state-of-the-art classification, hold potential
+for adoption in various every-day analyses in the neuroimaging field.
+Besides increased performance, these models are more interpretable by
+automatically learning a mapping to and from a brain-network space.
+This domain-specifc classification algorithms
+encourages departure from the artificial and statistically
+less attractive voxel space.
+From a neurobiological perspective,
+we demonstrate that neural activity underlying defined mental operations
+can readily be explained by linear combinations of the main activation
+patterns. That is,
+neural activity data as measured by fMRI probably concentrate near
+a low-dimensional manifold of brain networks.
+These fundamental building blocks of brain organization might
+facilitate the quest for the cognitive primitives of
+human thought.
+Moreover, data-intense neuroimaging studies in the future can be made tractable
+by reduction to the neurobiological essence.
+This can be particulary useful in multi-center clinical studies
+on the prediction of disease trajectories and drug treatment responses
+in psychiatric and neurological populations.
+We hope that these first steps stimulate development towards
+powerful semi-supervised classification methods.
 
--hypothesis space includes sparse PCA and PCA but not ICA since no
-linearity
-
--respect structure in the fMRI data
-
-- We hope that these results stimulate the development of
-even more powerful semi-supervised classification methods
-
-learn optimal parameters
-
-- improve computational tractability, prediction accuracy, and interpretability
-neuroimaging datasets
-- does it produce testable predictions for targeted experimental investigation?
--> we can test predictive value of network-network architectures across
-mental domains.
-
-open window to study the correspondence between brain dynamics during
-every-day mind-wandering and task-focused brain states.
-
-repertoire of mental operations that the human brain can perform
-
-large quantities of neuroimaging data
-logistic regression in an autoencoder paradigm
-
-automatically learn a mapping to and from a brain-network space
-
-statistical structure
-
-PCA:
-captures structure in the data that is well described by Gaussian clouds,
-linear pair-wise correlations are most important form of statistical
-dependence/orthongonal components
-
--> BOLD images can readily be reduced to linear combinations of
-sparse spatial structures
-
-scale different classificatiuon architectures to large neuroimaging data
-
-task data might concentrate near a low-dimensional manifold of brain networks
-
-neurobiolgocial fidelity
-
-entities of the neuroimaging domain
-
-put probability mas shwere we expeted neurobiological strucuture
-
-canonical set of brain networks
-
-new baseline
-
-FUTURE
-link the restings-tate component to the pattern sof cognitive proceses
-reduce information von resting-state data in neurological and psychiatric
-populations for discovery of neurobiological sub-groups as well as
-prediction of disease trajectories and drug responses
-
-%
 \paragraph{Acknowledgment}
@@
 Data were provided the Human Connectome Project. The study was supported
-by the German National Academic Foundation (D.B.).
-}
+by the German National Academic Foundation (D.B.)}
 

### 2015-05-26 nips3mm.py
before 1st BT correctoin

@@ -447,7 +447,7 @@ def dump_comps(masker, compressor, components, threshold=2):
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
 
-n_comps = [100]
+n_comps = [20]
 # n_comps = [40, 30, 20, 10, 5]
 for n_comp in n_comps:
     # for lambda_param in [0]:
@@ -462,7 +462,7 @@ for n_comp in n_comps:
             n_hidden=n_comp,
             gain1=0.004,  # empirically determined by CV
             learning_rate = np.float32(0.00001),  # empirically determined by CV,
-            max_epochs=500, l1=l1, l2=l2, lambda_param=lambda_param)
+            max_epochs=2500, l1=l1, l2=l2, lambda_param=lambda_param)
         
         estimator.fit(X_rest, X_task, labels)
 

### 2015-05-26 paper_nips/paper.tex
bertrands 1st comments on paper

@@
 on the pertinence of the assumed neurobiological structure.
+%XXX: this last statement is not cristal clear.
+%
+% XXX: Big gap between these two sentences. It is unclear what you call 'supervised' or 'unsupervised'.
 %
@@
 First, the mentioned supervised neuroimaging analyses operate
-without exception in a voxel space. This ignores the fact that the BOLD
+without exception in a voxel space. 
+%XXX:not exactly: the DCM is based on ROIs typically. PPI also ?
+
+This ignores the fact that the BOLD
 signal exhibits distributed spatial activation patterns.
@@
 task-induced changes of the BOLD signal typically amount to less than 5\%
-of baseline activity \cite{fox07}. They do thus not exploit the high similarity
+of baseline activity \cite{fox07}. 
+%
+% XXX I think that this is wrong. Moreover, it is not completely  clear how crucial this is. I think that there are better arguments to introduce the joint analysis of rest and task.
+% For example, the existence of abundant RS data is an important practical reson to consider them. 'Better sample the manifold of brain states'
+% 
+They do thus not exploit the high similarity
 of BOLD dynamics in the human brain at rest and during experimental tasks,
@@
 
-Both these biological properties can be conjointly exploited in a 
-semi-supervised (i.e., use rest and task data)
+Both these biological properties can be conjointly exploited in a mixed
+ (i.e., use rest and task data)
 low-rank (i.e., perform network decomposition)
@@
 The integration of brain-network discovery into a 
-supervised classification framework should identify the most relevant
+supervised classification formally yields a semi-supervised learning framework that should identify the most relevant
 neurobiological structure for the prediction problem at hand.
@@
     \includegraphics[width=0.40\textwidth]{figures/figure1.png}
-  \caption {\textbf{Architecture}
+  \caption {\textbf{Architecture. XXX develop the cpation. Main msg ?}
   }
@@
 and feeding the first-layer activation into any nonlinear convex
+%% ICA contrasts are rarely convex ?!?
 function \cite{le2011ica}.
@@
 under mild conditions.
+%
 Hence, autoencoders can flexibly project the neuroimaging data
 onto the main axes of variation and thus
-reverse-engineer properties of the data-generating
+reverse-engineer properties of the underlying
 neural processes \cite{olshausen96}.
@@
 into a low-rank logistic regression fed by (labeled) task data.
-Using the chain rule for gradient descent, we can then
-solve the unsupervised data representation and the supervised classification
-in an identical statistical learning problem.
+% Using the chain rule for gradient descent, 
+We can then solve the unsupervised data representation and the
+supervised classification in an identical statistical learning
+problem.
+%
 From the perspective of dictionary learning, the first layer represents
@@
 reduced Vapnik-Chervonenkis dimensions of the trained estimator.
+% I would not make direct claims on *VC* dimension, since you don't consider it explictly
 %
@@
 and grand-mean intensity normalization. Activation maps were spatially
-smoothed by a Gaussian kernel of 4mm (FWHM). A GLM was
+smoothed with a Gaussian kernel of 4mm (FWHM). A GLM was
 implemented by FILM from the FSL suite with model regressors from convolution
@@
 coverage \cite{barch2013}.
+%
+% No, this is not what is called 'coverage'
+%
 In sum, the HCP task data incorporated 8650 first-level activity maps
@@
 $x' \in \mathbb{R}^{d}$ is the reconstruction vector of the original activity map
-from the hidden variables. Further, ${W_0}$ denotes the weight matrix that
+from the hidden variables. 
+% XXX use bold letters for vectors and matrices
+Further, ${W_0}$ denotes the weight matrix that
 transforms
@@
     P(Y=i|x, V_0,V_1,b_{V_0}, b_{V_1}) &= softmax_i(V_1 (V_0 x + b_{V_0}) + b_{V_1}) \\
-    &= \frac {e^{V_1_i (V_0_i x + b_{V_0}_i) + b_{V_1}_i}} {\sum_j e^{V_1_j (V_0_j x + b_{V_0}_j) + b_{V_1}_j}}            
+    &= \frac {e^{V_{1i} (V_{0i} x + b_{V_{0i}}) + b_{V_{1i}}}} {\sum_j e^{V_{1j} (V_{0j} x + b_{V_{0j}}) + b_{V_{1j}}}}            
   \end{split}
@@
 \paragraph{Optimization.}
-The common objective was approximated in the neuroimaging data
-by updating parameters by derivates of the 
+The common objective was approximated 
+% XXX why 'approximated' ?
+in the neuroimaging data
+by updating parameters by derivates 
+% XXX derivates ???
+of the 
 semi-supervised low-rank logistic regression.
@@
 Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
-The hyper-parameter $\alpha$ was set to $0.00001$ by prior manual
-cross-validation and $\epsilon$ was set to $10^{-6}$.
+The hyper-parameter $\alpha$ was set to $0.00001$ by prior studies and $\epsilon$ was set to $10^{-6}$.
 %
@@
 \bigskip
-
+% XXX the accuracy figure is not very nice. export is in pdf.
+% XXX change the titles. remove the resolution information from the title.
+% XXX larger legends. they're not readable currently
 \begin{figure}
@@
 across iterations over the HCP task data (epochs).
+XXX: Message ??
 }
@@
 (thresholded at 0.25).
-All values z-scored.
+All values are z-scored.
+% XXX: Why is the z-map sparse ?!? 
 }
@@
 \begin{figure}
+% XXX: figure not readable
 \begin{centering}
@@
   (\textit{right column}). Matrix weights were z-scored.
+% XXX To me there is no clear message here. You can always interpolate between arbitrary maps... What does this bring to you ? 
  }
@@
 There is an increasing occurrence of high-dimensional problems in the
-neuroimaging domain. Yet, there is also considerable uncertainty about the
+neuroimaging domain. 
+%
+% XXX: meaningless and useless for your work
+%
+Yet, there is also considerable uncertainty about the
 most pertinent representation
@@
 in a neurobiological and interpretable fashion.
+% You really should emphasize the problem of accumulating data and knowledge which is much more interesting. Cite Yannick.
 

### 2015-05-27 nips3mm.py
added figures for nips

@@ -138,7 +138,8 @@ class SSEncoder(BaseEstimator):
         compr_matrix = self.W0s.get_value().T  # currently best compression
         AT_X_compr = np.dot(compr_matrix, AT_X.T).T
         clf = LogisticRegression(penalty='l1')
-        folder = StratifiedShuffleSplit(y=AT_labels, n_iter=5, test_size=0.2)
+        folder = StratifiedShuffleSplit(y=AT_labels, n_iter=5, test_size=0.2,
+                                        random_state=42)
 
         acc_list = []
         prfs_list = []
@@ -451,7 +452,7 @@ n_comps = [20]
 # n_comps = [40, 30, 20, 10, 5]
 for n_comp in n_comps:
     # for lambda_param in [0]:
-    for lambda_param in [0.1]:
+    for lambda_param in [0.50]:
         l1 = 0.1
         l2 = 0.1
         my_title = r'Low-rank LR + AE (combined loss, shared decomp): n_comp=%i L1=%.1f L2=%.1f lambda=%.2f res=3mm spca20RS' % (
@@ -462,7 +463,7 @@ for n_comp in n_comps:
             n_hidden=n_comp,
             gain1=0.004,  # empirically determined by CV
             learning_rate = np.float32(0.00001),  # empirically determined by CV,
-            max_epochs=2500, l1=l1, l2=l2, lambda_param=lambda_param)
+            max_epochs=500, l1=l1, l2=l2, lambda_param=lambda_param)
         
         estimator.fit(X_rest, X_task, labels)
 

### 2015-05-27 nips3mm.py
new weight figure with 4fwhm

@@ -426,9 +426,10 @@ class SSEncoder(BaseEstimator):
 # plot figures
 ##############################################################################
 
-def dump_comps(masker, compressor, components, threshold=2):
+def dump_comps(masker, compressor, components, threshold=2, fwhm=None):
     from scipy.stats import zscore
     from nilearn.plotting import plot_stat_map
+    from nilearn.image import smooth_img
 
     if isinstance(compressor, basestring):
         comp_name = compressor
@@ -447,6 +448,14 @@ def dump_comps(masker, compressor, components, threshold=2):
         plot_stat_map(gz_path, bg_img='colin.nii', threshold=threshold,
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
+                      
+        # optional: do smoothing
+        if fwhm is not None:
+            nii_z_fwhm = smooth_img(nii_z, fwhm=fwhm)
+            plot_stat_map(nii_z_fwhm, bg_img='colin.nii', threshold=threshold,
+                          cut_coords=(0, -2, 0), draw_cross=False,
+                          output_file=path_mask +
+                          ('zmap_%imm.png' % fwhm))
 
 n_comps = [20]
 # n_comps = [40, 30, 20, 10, 5]
@@ -995,12 +1004,14 @@ for n_comp in n_comps:
                 (target_lambda, n_comp)))
 
 # print network components (1st layer)
+from nilearn.image import smooth_img
 n_comp = 20
+lmbd = 0.25
 pkgs = glob.glob(RES_NAME + '/*W0comps.npy')
 for p in pkgs:
     lambda_param = np.float(re.search('lambda=(.{4})', p).group(1))
     n_hidden = int(re.search('comp=(.{1,3})_', p).group(1))
-    if n_comp != n_hidden:
+    if n_comp != n_hidden or lambda_param != lmbd:
         continue
         
     new_fname = 'comps_n=%i_lambda=%.2f_th0.0' % (n_hidden, lambda_param)
@@ -1022,14 +1033,17 @@ for p in pkgs:
     comps = np.dot(np.load(q), np.load(p))
         
     new_fname = 'comps_n=%i_lambda=%.2f_th0.0' % (n_hidden, lambda_param)
-    dump_comps(nifti_masker, new_fname, comps, threshold=0.0)
[…]

### 2015-05-27 nips3mm_h38.py

@@ -450,11 +450,11 @@ def dump_comps(masker, compressor, components, threshold=2):
                       output_file=path_mask + 'zmap.png')
 
 from matplotlib import pylab as plt
-n_comps = [20]
+n_comps = [100]
 # n_comps = [40, 30, 20, 10, 5]
 for n_comp in n_comps:
     # for lambda_param in [0]:
-    for lambda_param in [0.5, 0, 0.75, 1]:
+    for lambda_param in [0.25]:
         l1 = 0.1
         l2 = 0.1
         my_title = r'Low-rank LR + AE (combined loss, shared decomp): n_comp=%i L1=%.1f L2=%.1f lambda=%.2f res=3mm spca20RS' % (

### 2015-05-27 nips3mm_vanilla.py

@@ -272,9 +272,10 @@ class SSEncoder(BaseEstimator):
 # plot figures
 ##############################################################################
 
-def dump_comps(masker, compressor, components, threshold=2):
+def dump_comps(masker, compressor, components, threshold=2, fwhm=None):
     from scipy.stats import zscore
     from nilearn.plotting import plot_stat_map
+    from nilearn.image import smooth_img
 
     if isinstance(compressor, basestring):
         comp_name = compressor
@@ -293,7 +294,14 @@ def dump_comps(masker, compressor, components, threshold=2):
         plot_stat_map(gz_path, bg_img='colin.nii', threshold=threshold,
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
-
+                      
+        # optional: do smoothing
+        if fwhm is not None:
+            nii_z_fwhm = smooth_img(nii_z, fwhm=fwhm)
+            plot_stat_map(nii_z_fwhm, bg_img='colin.nii', threshold=threshold,
+                          cut_coords=(0, -2, 0), draw_cross=False,
+                          output_file=path_mask +
+                          ('zmap_%imm.png' % fwhm))
 
 l1 = 0.1
 l2 = 0.1
@@ -388,4 +396,4 @@ for k, v in d.iteritems():
 pkgs = ['nips3mm_vanilla/V0comps.npy']
 comps = np.load(pkgs[0])
 new_fname = 'comps_th0.0'
-dump_comps(nifti_masker, new_fname, comps, threshold=0.0)
+dump_comps(nifti_masker, new_fname, comps, threshold=0.0, fwhm=4)

### 2015-05-28 paper_nips/paper.tex
updated discussion structure

@@
 %
-However, modelling the neural correlates of mental tasks hinges
-on the pertinence of the assumed neurobiological structure.
-%XXX: this last statement is not cristal clear.
+However, testing hypotheses on the neural correlates of mental tasks
+depends on the representation of the observations.
 %
-% XXX: Big gap between these two sentences. It is unclear what you call 'supervised' or 'unsupervised'.
-%
-We therefore propose to solve the unsupervised dimensionality reduction
-and supervised task classification in
+We therefore propose to optimize
+representation and task classification in
 an identical statistical learning problem.
@@
 %
-Methods used in neuroimaging research can be grouped by discovering
+Methods for neuroimaging research can be grouped by discovering
 neurobiological structure or assessing the neural correlates associated
 with mental tasks.
-To discover spatial distributions of activation structure across time,
-independent component analysis (ICA) \cite{beckmann2005} is often used
-to decompose the BOLD (blood-oxygen level-dependent) signals into the
-main modes of variation.
-The ensuing spatial activation patterns are believed to represent
+To discover, on the one hand, spatial distributions of neural activity
+structure across time,
+independent component analysis (ICA) \cite{beckmann2005} is often used.
+It decomposes the BOLD (blood-oxygen level-dependent) signals into the
+primary modes of variation.
+The ensuing spatial activity patterns are believed to represent
 brain networks of
-functionally interacting brain regions \cite{smith2009}.
+functionally interacting regions \cite{smith2009}.
 Similarly, sparse principal component analysis (SPCA) \cite{varoqu2011}
@@
 separate BOLD signals into parsimonious network components.
-Thus extracted brain networks have been shown to be
+Thus extracted brain networks are probably
 manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
 Their fundamental organizational role is further
-attested by continued covariation during sleep and anesthesia.
+attested by continued covariation during sleep and anesthesia \cite{fox07}.
 %
-Network discovery is typically performed by applying ICA or SPCA on
-unlabeled "resting-state" data. These capture brain dynamics
+Network discovery by applying ICA or SPCA is typically performed on
+task-unrelated (i.e., \textit{unlabeled}) ``resting-state'' data.
+These capture brain dynamics
 during ongoing random thought without controlled environmental stimulation.
-A large proportion of the BOLD signal is known
+In fact, a large proportion of the BOLD signal is known
 not to correlate with a particular behavior, stimulus, or experimental task
@@
 
-On the other hand, to investigate
+To test, on the other hand,
 the neural correlates underlying mental tasks,
@@
 Alternatively, psychophysiological interactions (PPI)
-elucidate the functional interactions between voxels as a function
-of experimental tasks \cite{friston97}.
-Dynamic causal modeling (DCM), in turn, quantifies directed,
-task-driven influences between regions
-by treating the brain as a nonlinear dynamic system with unknown
-neuronal states \cite{stephan04}.
+elucidate the influence of one brain region on another conditioned
+by experimental tasks \cite{friston97}.
 As a last example, always more neuroimaging studies model
@@
 \cite{poldrack09decoding}.
-All these methods are applied to labeled task data that capture brain dynamics
+All these methods are applied to task-associated (i.e., \textit{labeled})
+data that capture brain dynamics
 during stimulus-guided behavior.
 Two important conclusions can be drawn.
-First, the mentioned supervised neuroimaging analyses operate
-without exception in a voxel space. 
-%XXX:not exactly: the DCM is based on ROIs typically. PPI also ?
-
+First, the mentioned supervised neuroimaging analyses typically yield
+results in a voxel space.
 This ignores the fact that the BOLD
-signal exhibits distributed spatial activation patterns.
-Second, existing neuroimaging analyses do not acknowledge the fact that the
-task-induced changes of the BOLD signal typically amount to less than 5\%
-of baseline activity \cite{fox07}. 
+signal exhibits spatially distributed activity patterns.
 %
-% XXX I think that this is wrong. Moreover, it is not completely  clear how crucial this is. I think that there are better arguments to introduce the joint analysis of rest and task.
-% For example, the existence of abundant RS data is an important practical reson to consider them. 'Better sample the manifold of brain states'
-% 
-They do thus not exploit the high similarity
-of BOLD dynamics in the human brain at rest and during experimental tasks,
+Second, existing supervised neuroimaging analyses
+do not acknowledge the abundance
+of easily acquired resting-state data \cite{biswaldiscovery}.
+These may allow better sampling the manifold of brain states
+due to the high task-rest similarities of neural activity patterns,
 as observed using ICA \cite{smith2009}.
@@
 Both these biological properties can be conjointly exploited in a mixed
- (i.e., use rest and task data)
+ (i.e., use task and rest data)
 low-rank (i.e., perform network decomposition)
@@
 %
-The integration of brain-network discovery into a 
-supervised classification formally yields a semi-supervised learning framework that should identify the most relevant
-neurobiological structure for the prediction problem at hand.
+The integration of brain-network discovery into
+supervised classification formally yields a semi-supervised learning
+framework. The most relevant
+neurobiological structure should hence be identified
+for the prediction problem at hand.
 %
@@
 Moreover, they have the characteristics of an ICA in case of tied weights
-and feeding the first-layer activation into any nonlinear convex
-%% ICA contrasts are rarely convex ?!?
-function \cite{le2011ica}.
+and adding a nonlinear convex transformation \cite{le2011ica}.
 These authors further demonstrated that ICA, sparse autoencoders, and 
@@
 into a low-rank logistic regression fed by (labeled) task data.
-% Using the chain rule for gradient descent, 
 We can then solve the unsupervised data representation and the
@@
 valid models in the hypothesis set.
-%
-The generalization performance should consequently be improved due to 
-reduced Vapnik-Chervonenkis dimensions of the trained estimator.
-% I would not make direct claims on *VC* dimension, since you don't consider it explictly
 %
@@
 T1-weighted scans, nonlinear (FNIRT) registration into MNI space,
-and grand-mean intensity normalization. Activation maps were spatially
+and grand-mean intensity normalization. Activity maps were spatially
 smoothed with a Gaussian kernel of 4mm (FWHM). A GLM was
@@
 with a “canonical” hemodynamic response function and from temporal derivatives.
-HCP tasks were conceived to modulate activation
+HCP tasks were conceived to modulate activity
 in a maximum of different brain regions and neural systems. Indeed, at
@@
 slice timing, motion correction, alignment, and spatial normalization.
-Activation maps were spatially smoothed by
+activity maps were spatially smoothed by
 a Gaussian kernel of 5mm (FWHM).
@@
 The affine autoencoder takes the input 
-$x$ and projects it into coordinates of a latent representation $z$
+$\mathbf{x}$ and projects it into coordinates of a
+latent representation $\mathbf{z}$
 by
@@
   \begin{split}
-  z &= W_0 x + b_{W_0} \\
-  x' &= W_1 z + b_{W_1}
+    \mathbf{z} &= \mathbf{W_0} \mathbf{x} + \mathbf{b_{W_0}} \\
+    \mathbf{x'} &= \mathbf{W_1} \mathbf{z} + \mathbf{b_{W_1}}
   \end{split}
@@
 
-where $x \in \mathbb{R}^{d}$ denotes the vector of $d=79,941$
+where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79,941$
 voxel values from each
-rest map, $z \in \mathbb{R}^{n}$ is the $n$ hidden dimensions (i.e.,
+rest map,
+$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$ hidden dimensions (i.e.,
 distributed neural activity patterns), and 
-$x' \in \mathbb{R}^{d}$ is the reconstruction vector of the original activity map
+$\mathbf{x' \in \mathbb{R}^{d}}$
+is the reconstruction vector of the original activity map
 from the hidden variables. 
 % XXX use bold letters for vectors and matrices
-Further, ${W_0}$ denotes the weight matrix that
+Further, $\mathbf{W_0}$ denotes the weight matrix that
 transforms
 from input space into the hidden space (encoder),
-${W_1}$ is the weight matrix for back-projection from the hidden variables to the
+$\mathbf{W_1}$ is the weight matrix for back-projection
+from the hidden variables to the
 output space (decoder).
-${b_{W_0}}$ and ${b_{W_1}}$ are bias vectors.
-Note that ${W_0}$ and ${W_1}$ are tied,
-$W_0$ = $W_1^T$.
-The optimal model parameters ${W_0}, {b_{W_0}}, {b_{W_1}}$ are found by
+$\mathbf{b_{W_0}}$ and $\mathbf{b_{W_1}}$ are bias vectors.
+Note that $\mathbf{W_0}$ and $\mathbf{W_1}$ are tied,
+$\mathbf{W_0} = \mathbf{W_1^T}$.
+The optimal model parameters $\mathbf{W_0, b_{W_0}, b_{W_1}}$ are found by
 minimizing the imperfection of reconstruction according to squared error
@@
 \begin{eqnarray}
-  {\mathcal{L_{AE}}}(x, x') = || x - x' ||^2
+  {\mathcal{L_{AE}}}(\mathbf{x}, \mathbf{x'}) = || \mathbf{x} - \mathbf{x'} ||^2
 \end{eqnarray}
@@
 multinomial logistic regression.
-It gives the probability of an input $x$ to belong
+It gives the probability of an input $\mathbf{x}$ to belong
 to a class $i \in \mathbb{N}$
@@
   \begin{split}
-    P(Y=i|x, V_0,V_1,b_{V_0}, b_{V_1}) &= softmax_i(V_1 (V_0 x + b_{V_0}) + b_{V_1}) \\
-    &= \frac {e^{V_{1i} (V_{0i} x + b_{V_{0i}}) + b_{V_{1i}}}} {\sum_j e^{V_{1j} (V_{0j} x + b_{V_{0j}}) + b_{V_{1j}}}}            
+    P(Y=i|\mathbf{x, V_0,V_1,b_{V_0}, b_{V_1}}) &= softmax_i(\mathbf{V_1 (V_0 x + b_{V_0}) + b_{V_1}}) \\
+    &= \frac {e^{\mathbf{V_{1i} (V_{0i} x + b_{V_{0i}}) + b_{V_{1i}}}}} {\sum_j e^{\mathbf{V_{1j} (V_{0j} x + b_{V_{0j}}) + b_{V_{1j}}}}}            
   \end{split}
@@
 
-where the matrix $V_0 \in \mathbb{R}^{dxn}$ transforms the input $x \in \mathbb{R}^{d}$
+where the matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
+transforms the input $\mathbf{x \in \mathbb{R}^{d}}$
 into $n$ latent components
-and the matrix $V_1 \in \mathbb{R}^{nxc}$ projects the latent components
+and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxc}}$
+projects the latent components
 onto hyperplanes that reflect $c$ class probabilities.
-$b_{V_0}$ and $b_{V_1}$ are corresponding
+$\mathbf{b_{V_0}}$ and $\mathbf{b_{V_1}}$ are corresponding
 bias vectors.
@@
   \begin{split}
-    {\mathcal{L_{LR}}}(x, y) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1})
+    {\mathcal{L_{LR}}}(\mathbf{x, y}) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|\mathbf{x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1}})
 \end{split}
@@
 \begin{eqnarray}
-  V_0 = W_0
+  \mathbf{V_0} = \mathbf{W_0}
 \end{eqnarray}
@@
 In so doing, we search for the combined model parameters
-$\theta=\{V_0,V_1,b_{V_0}, b_{V_1}, b_{W_0}, b_{W_1}\}$
+$\theta=\{\mathbf{V_0,V_1,b_{V_0}, b_{V_1}, b_{W_0}, b_{W_1}}\}$
 with respect to the (unsupervised) reconstruction error and the
@@
   \begin{split}
-v^{(t+1)} &= \rho v^{(t+1)} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
+    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
 %v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
 \\
-\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{v^{(t+1)} + \epsilon}},
+\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
   \end{split}
@@
 \bigskip
-% XXX the accuracy figure is not very nice. export is in pdf.
-% XXX change the titles. remove the resolution information from the title.
-% XXX larger legends. they're not readable currently
 \begin{figure}
@@
 \begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/support_bars.png}
+\includegraphics[width=1.00\textwidth]{figures/semisup.png}
 \end{centering}
@@
 the HCP task data.
+
+\textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning
+}
 }
@@
 \end{itemize}
+
+the features of the classes show that we captured activaiton properties,
+not particpant or paradigm properties
 
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.40\textwidth]{figures/figure3.png}
-  \end{center}
-  \caption {\textbf{
-  Equilibrium between autoencoder and low-rank logistic regression}
-  One learned decomposition component (out of 20) between the only-autoencoder
-  (\textit{uppermost panel}) and only-logistic-regression
-  (\textit{lowermost panel}) scenarios.
-  The congruent structure (\textit{red, middle column})
-  in the
-  posterior cingulate cortex, posterior mid-cingulate cortex, and medial
-  prefrontal cortex appears to be important in decompositions of the rest and
-  task neuroimaging data. With increasing weight on the supervised learning,
-  non-congruent structure emerges in the early (\textit{blue}) and
-  lateral (\textit{red}) visual cortex
-  (\textit{right column}). Matrix weights were z-scored.
-% XXX To me there is no clear message here. You can always interpolate between arbitrary maps... What does this bring to you ? 
- }
-\end{figure}
-
 \section{Discussion and Conclusion}
-%
-There is an increasing occurrence of high-dimensional problems in the
-neuroimaging domain. 
-%
-% XXX: meaningless and useless for your work
-%
-Yet, there is also considerable uncertainty about the
-most pertinent representation
-of neural activation information.
-This calls for new statistical learning algorithms that
-behave well in large samples. Ideally, they should acknowledge
-and exploit widely-accepted neuroscientific knowledge.
-We here propose such an estimator that learns
-dimensionality reduction
-in a neurobiological and interpretable fashion.
-% You really should emphasize the problem of accumulating data and knowledge which is much more interesting. Cite Yannick.
-
+% Conceptual recap+gain of current paper
 Using the flexibility of autoencoders, we
 learn the optimal decomposition from high-dimensional
-voxel brain space into the most important activation patterns for a
-statistical learning question of interest.
+voxel brain space into the most important activity patterns for a
+supervised learning question.
 The higher generalization accuracy and support recovery, comparing to
-state-of-the-art classification, hold potential
-for adoption in various every-day analyses in the neuroimaging field.
+ordinary logistic regression, hold potential
+for adoption in various neuroimaging analyses.
 Besides increased performance, these models are more interpretable by
@@
 less attractive voxel space.
-From a neurobiological perspective,
-we demonstrate that neural activity underlying defined mental operations
-can readily be explained by linear combinations of the main activation
+Neurobiologically,
+neural activity underlying defined mental operations
+can be explained by linear combinations of the main activity
 patterns. That is,
-neural activity data as measured by fMRI probably concentrate near
+fMRI data probably concentrate near
 a low-dimensional manifold of brain networks.
@@
 human thought.
-Moreover, data-intense neuroimaging studies in the future can be made tractable
-by reduction to the neurobiological essence.
-This can be particulary useful in multi-center clinical studies
-on the prediction of disease trajectories and drug treatment responses
-in psychiatric and neurological populations.
 We hope that these first steps stimulate development towards
-powerful semi-supervised classification methods.
+powerful semi-supervised classification methods in systems neuroscience.
+
+% grand perspective
+In the future, automatic reduction of brain images to
+their neurobiological essence
+may facilitate data-intense neuroimaging studies.
+Initiatives for data collection are rapidly increasing
+in neuroscience \cite{poldrack2014data}.
+These promise integration
+of accumulating neuroscientific knowledge from neuroimaging databases.
+Tractability by
+condensed feature representations can avoid the ill-posed problem of
+learning the full distribution of activity patterns.
+%
+This is relevant to the 
+multi-class challenges spanning the human cognitive space
+\cite{schwartz2013mapping}
+and
+combination of neuroimaging with
+high-throughput genetic analyses, such as GWAS
+\cite{need2010gwas}.
+%
+The biggest socioeconomic potential may
+lie in across-hospital clinical studies that
+predict disease trajectories and drug responses
+in psychiatric/neurological populations
+\cite{frackowiak2015future, gustav2011cost}.
 
@@
 Data were provided the Human Connectome Project. The study was supported
-by the German National Academic Foundation (D.B.)}
+by the German National Academic Foundation (D.B.).
+}
 

### 2015-05-28 nips3mm_scarcity.py

@@ -8,6 +8,9 @@ preliminary conclusion:
 questions:
 - the AE cost is much bigger, is it ok to T.log() that?
 No because it renders the objective function non-convex.
+- compare to nonlinearity + PCA only
+- how to genreate examples from 100 last model paramters during the
+gradient descent?
 """
 
 print __doc__
@@ -18,7 +21,6 @@ import numpy as np
 import glob
 from scipy.linalg import norm
 import nibabel as nib
-from matplotlib import pylab as plt
 from sklearn.grid_search import RandomizedSearchCV
 from sklearn.base import BaseEstimator
 from sklearn.preprocessing import StandardScaler
@@ -27,15 +29,20 @@ from sklearn.metrics import precision_recall_fscore_support
 from sklearn.metrics import confusion_matrix
 import theano
 import theano.tensor as T
+from matplotlib import pylab as plt
 print('Running THEANO on %s' % theano.config.device)
 from nilearn.image import concat_imgs
 import joblib
 import time
 
-RES_NAME = 'nips3mm_h38'
+RES_NAME = 'nips3mm_scarcity'
 WRITE_DIR = op.join(os.getcwd(), RES_NAME)
 if not op.exists(WRITE_DIR):
     os.mkdir(WRITE_DIR)
+
+MAX_SAMPLES = 2000
+MAX_REST_SAMPLES = 1000
+MAX_TASK_SAMPLES = 100
 
 ##############################################################################
 # load+preprocess data
@@ -50,7 +57,7 @@ mask_nvox = nifti_masker.mask_img_.get_data().sum()
 print('Loading data...')
 
 # ARCHI task
-X_task, labels = joblib.load('preload_HT38_3mm')
+X_task, labels = joblib.load('preload_HT_3mm')
 
 labels = np.int32(labels)
 
@@ -72,15 +79,6 @@ del rs_spca_niis
 del rs_spca_data
 
 X_task = StandardScaler().fit_transform(X_task)
-
-# HACK!
-multiplier = np.int(np.ceil(X_task.shape[0] /
-                    (X_rest.shape[0] * 1.)))  # how much do we need to expand?
-X_rest = np.repeat(X_rest, multiplier, axis=0)
-new_inds2 = np.arange(0, X_rest.shape[0])
-np.random.shuffle(new_inds2)
-X_rest = X_rest[new_inds2]  # shuffle the neighboring elements (due to repeat)
-X_rest = X_rest[:X_task.shape[0]]  # shrink to size of task dataset
[…]

### 2015-05-29 paper_nips/paper.tex
new figures

@@
 including PCA, SPCA, and ICA \cite{hinton06}.
+\begin{wrapfigure}{r}{0.40\textwidth}
+  \centering
+  %\captionsetup{labelformat=nonumber}
+    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
+  \vspace{-0.7cm}
+  \caption {\textbf{Architecture.}
+  Linear autoencoders find an optimal compression of 79,941 brain voxels
+  into $n$ unknown activity patterns by improving reconstruction from them.
+  The decomposition matrix equates with the bottleneck of
+  a factored logistic regression.
+  Supervised multi-class learning on task data ($X_{task}$)
+  is thus guided by
+  unsupervised decomposition of rest data ($X_{rest}$).
+  }
+\end{wrapfigure}
+%
 Autoencoders
@@
 in case of one linear hidden layer and a squared error loss
-%
-\begin{wrapfigure}{r}{0.40\textwidth}
-  \centering
-  %\captionsetup{labelformat=nonumber}
-    \includegraphics[width=0.40\textwidth]{figures/figure1.png}
-  \caption {\textbf{Architecture. XXX develop the cpation. Main msg ?}
-  }
-\end{wrapfigure}
-%
 \cite{baldi1989neural}.
@@
 \caption{\textbf{Comparing the interpretability of classification weights}
-The voxel predictors corresponding to 5 (of 18) psychological tasks (\textit{rows})
+The voxel predictors corresponding to 5 exemplary
+(of 18 total) psychological tasks (\textit{rows})
 from the HCP dataset \cite{barch2013}.
-\textit{Left column:} ordinary logistic regression (based on same
-implementation but without bottleneck layer or autoencoder),
+\textit{Left column:} ordinary logistic regression (same
+implementation but without bottleneck or autoencoder),
 \textit{middle column:} semi-supervised low-rank logistic regression
 (latent components $n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
-\textit{right column:} voxel-wise average of whole-brain activity
-across the first-level activity maps from 498 participants
-(thresholded at 0.25).
-All values are z-scored.
-% XXX: Why is the z-map sparse ?!? 
+\textit{right column:} voxel-wise average of whole-brain
+activity maps from each task (thresholded at 0.25).
+Low-rank logistic regression
+a) puts higher absolute weights on relevant structure
+b) lower ones on irrelevant structure,
+and
+c) yields BOLD-typical local contiguity.
+All values are z-scored and smoothed by a Gaussian (fwhm=4).
 }
@@
 
+out of sample performance: extract features of general relevance
+beyond the supervised problem at hand
+
+support recovery:
+Low-rank logistic regression
+a) puts higher absolute weights on relevant structure
+b) lower ones on irrelevant structure,
+and
+c) yields BOLD-typical local contiguity.
+
 \paragraph{Classification performance}
@@
 
-\paragraph{Support recovery}
+\paragraph{Support recovery -> feature identification}
 \begin{itemize}
@@
   generalization performance}
-  \item{low L1/L2 penalties (0.1) yielded better results
-  than higher values (0.3, 0.5)}
   \item{``pretraining'' of the bottleneck by settings weights to
@@
 a low-dimensional manifold of brain networks.
-These fundamental building blocks of brain organization might
+Extracting fundamental building blocks of brain organization might
 facilitate the quest for the cognitive primitives of
@@
 their neurobiological essence
-may facilitate data-intense neuroimaging studies.
+may leverage data-intense neuroimaging studies.
 Initiatives for data collection are rapidly increasing
 in neuroscience \cite{poldrack2014data}.
-These promise integration
+These promise structured integration
 of accumulating neuroscientific knowledge from neuroimaging databases.
@@
 %
-This is relevant to the 
+This is not only relevant to the 
 multi-class challenges spanning the human cognitive space
 \cite{schwartz2013mapping}
+but also the
+multi-modal combination with
+high-resolution 3D models of brain anatomy \cite{amunts2013bigbrain}
 and
-combination of neuroimaging with
-high-throughput genetic analyses, such as GWAS
-\cite{need2010gwas}.
+high-throughput genome analyses \cite{need2010gwas}.
 %
@@
 {\small
-Data were provided the Human Connectome Project. The study was supported
+Data were provided by the Human Connectome Project. The study was supported
 by the German National Academic Foundation (D.B.).

### 2015-05-29 nips3mm.py

@@ -426,10 +426,12 @@ class SSEncoder(BaseEstimator):
 # plot figures
 ##############################################################################
 
-def dump_comps(masker, compressor, components, threshold=2, fwhm=None):
+def dump_comps(masker, compressor, components, threshold=2, fwhm=None,
+               perc=None):
     from scipy.stats import zscore
     from nilearn.plotting import plot_stat_map
     from nilearn.image import smooth_img
+    from scipy.stats import scoreatpercentile
 
     if isinstance(compressor, basestring):
         comp_name = compressor
@@ -441,18 +443,29 @@ def dump_comps(masker, compressor, components, threshold=2, fwhm=None):
                                                      n_comp, i_c + 1))
         nii_raw = masker.inverse_transform(comp)
         nii_raw.to_filename(path_mask + '.nii.gz')
+        
+        comp_z = zscore(comp)
+        
+        if perc is not None:
+            cur_thresh = scoreatpercentile(np.abs(comp_z), per=perc)
+            path_mask += '_perc%i' % perc
+            print('Applying percentile %.2f (threshold: %.2f)' % (perc, cur_thresh))
+        else:
+            cur_thresh = threshold
+            path_mask += '_thr%.2f' % cur_thresh
+            print('Applying threshold: %.2f' % cur_thresh)
 
-        nii_z = masker.inverse_transform(zscore(comp))
+        nii_z = masker.inverse_transform(comp_z)
         gz_path = path_mask + '_zmap.nii.gz'
         nii_z.to_filename(gz_path)
-        plot_stat_map(gz_path, bg_img='colin.nii', threshold=threshold,
+        plot_stat_map(gz_path, bg_img='colin.nii', threshold=cur_thresh,
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
                       
         # optional: do smoothing
         if fwhm is not None:
             nii_z_fwhm = smooth_img(nii_z, fwhm=fwhm)
-            plot_stat_map(nii_z_fwhm, bg_img='colin.nii', threshold=threshold,
+            plot_stat_map(nii_z_fwhm, bg_img='colin.nii', threshold=cur_thresh,
                           cut_coords=(0, -2, 0), draw_cross=False,
                           output_file=path_mask +
[…]

### 2015-05-29 nips3mm_h38.py

@@ -450,7 +450,7 @@ def dump_comps(masker, compressor, components, threshold=2):
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
 
-n_comps = [20]
+n_comps = [100]
 # n_comps = [40, 30, 20, 10, 5]
 for n_comp in n_comps:
     # for lambda_param in [0]:
@@ -517,7 +517,7 @@ STOP_CALCULATION
 import re
 pkgs = glob.glob(RES_NAME + '/*dbg_epochs_*.npy')
 dbg_epochs_ = np.load(pkgs[0])
-n_comps = [100]
+n_comps = [20]
 pkgs = glob.glob(RES_NAME + '/*dbg_acc_train_*.npy')
 for n_comp in n_comps:
     plt.figure()
@@ -817,7 +817,7 @@ for n_comp in n_comps:
     plt.savefig(op.join(WRITE_DIR, 'rec_inds_%icomps.png' % n_comp))
 
 # in-dataset f1 at lambda=0.5
-pkgs = glob.glob(RES_NAME + '/*lambda=0.25*dbg_prfs_.npy')
+pkgs = glob.glob(RES_NAME + '/*lambda=1.00*dbg_prfs_.npy')
 for n_comp in n_comps:
     plt.figure()
     p_ep = glob.glob(RES_NAME + '/*comp=%i*dbg_epochs_*.npy' % n_comp)
@@ -844,7 +844,7 @@ for n_comp in n_comps:
         cur_label += 'separate decomp.' if 'decomp_separate' in p else 'joint decomp.'
         for i in np.arange(38):
             plt.plot(
-                dbg_epochs_,
+                dbg_epochs_[:len(np.array(dbg_prfs_)[:, 2, i])],
                 np.array(dbg_prfs_)[:, 2, i],
                 label='task %i' % (i + 1))
     plt.title('Classification performance for 38 tasks')
@@ -855,7 +855,7 @@ for n_comp in n_comps:
     plt.xlabel('epochs')
     plt.grid(True)
     plt.show()
-    plt.savefig(op.join(WRITE_DIR, 'f1_inds_%icomps.png' % n_comp))
[…]

### 2015-05-29 nips3mm_scarcity.py

@@ -41,8 +41,8 @@ if not op.exists(WRITE_DIR):
     os.mkdir(WRITE_DIR)
 
 MAX_SAMPLES = 2000
-MAX_REST_SAMPLES = 1000
-MAX_TASK_SAMPLES = 100
+MAX_REST_SAMPLES = 1
+MAX_TASK_SAMPLES = 1000
 
 ##############################################################################
 # load+preprocess data
@@ -465,7 +465,8 @@ n_comps = [20]
 # n_comps = [40, 30, 20, 10, 5]
 for n_comp in n_comps:
     # for lambda_param in [0]:
-    for lambda_param in [0.5]:
+    #for lambda_param in [0.5]:
+    for lambda_param in [1.0]:
         l1 = 0.1
         l2 = 0.1
         my_title = r'Low-rank LR + AE (combined loss, shared decomp): n_comp=%i L1=%.1f L2=%.1f lambda=%.2f res=3mm spca20RS max%i r%i t%i' % (
@@ -1142,6 +1143,97 @@ for (ntask, nrest), col in zip([[1000, 1500], [1000, 1000], [1000, 500],
         label=cur_label, color=col, linewidth=2.5)
 
 plt.title('Semi-supervision in data-rich and data-scarce scenarios')
+plt.yticks(np.linspace(0., 1., 11))
+plt.ylabel(k)
+plt.xlabel('epochs')
+plt.grid(True)
+plt.ylim(0., 1.05)
+plt.axhline(1. / 18, xmin=0, xmax=1, color='#696969', linewidth=2.5,
+            label='chance', linestyle='--')
+plt.show()
+plt.legend(loc='lower right', fontsize=11,
+           ncol=1, shadow=True, title="task-maps : rest-maps", fancybox=True)
+plt.savefig(op.join(WRITE_DIR,
+            k.replace(' ', '_') + '_%icomps_scarcities.png' % n_comp))
+
+# pkgs = glob.glob(RES_NAME + '/*dbg_epochs*.npy')
+# dbg_epochs_ = np.load(pkgs[0])
+# 
+# k, v = ('out-of-sample performance', '/*_n_comp=%i*r%i_t%i*dbg_acc_val_*.npy')
+# n_comp = 20
+# path_main = 'nips3mm_scarcity'
+
+# bars
+plt.close('all')
+cols = ('#EFC94C', '#E27A3F', '#DF4949',
[…]

### 2015-05-30 paper_nips/paper.tex
removed empty line in caption causing my compilation to break

@@
 \usepackage{amssymb}
+\usepackage{booktabs}
 
@@
   \begin{split}
-{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = (\lambda) {\mathcal{L_{LR}}}
+{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
 + (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \ell_1(\theta) + \ell_2(\theta)
@@
 \begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/figure_weights.jpg}
+\includegraphics[width=1.00\textwidth]{figures/figure_weights_perc75.png}
 \end{centering}
-\caption{\textbf{Comparing the interpretability of classification weights}
+\caption{\textbf{Interpretability of classification weights}
 The voxel predictors corresponding to 5 exemplary
@@
 \textit{middle column:} semi-supervised low-rank logistic regression
-(latent components $n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
+($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
 \textit{right column:} voxel-wise average of whole-brain
-activity maps from each task (thresholded at 0.25).
+activity maps from each task.
 Low-rank logistic regression
@@
 c) yields BOLD-typical local contiguity.
-All values are z-scored and smoothed by a Gaussian (fwhm=4).
+All values are z-scored and thresholded at the 75th percentile.
 }
@@
 the HCP task data.
-
 \textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning
@@
 \section{Experimental Results}
-\paragraph{General}
+\paragraph{Serial versus parallel structure discovery and classification.}
+We first tested for an advantage of combined unsupervised decomposition
+and supervised classification learning.
+We benchmarked against
+performing data reduction on the (unlabeled)
+first half of the HCP data by ICA and SPCA ($n=5, 20, 50, 100$ components)
+and learning classification models in the (labeled) second half
+by ordinary logistic regression.
+%
+ICA performed iterative blind source separation
+by a parallel FASTICA implementation (200 maximum iterations,
+per-iteration tolerance of 0.0001,
+initialized by random mixing matrix, whitening).
+SPCA separated the BOLD signals into
+network components with few regions by
+a regression-type optimization problem constrained by
+$\ell_1$-penalty
+(no orthogonality assumptions, 1000 maximum iterations,
+per-iteration tolerance of 1 * 10\textsuperscript{-8}, sparsity
+alpha=1).
+%
+The second half of the data was projected onto the
+latent components discovered in first data half.
+The ensuing component loadings were submitted to ordinary
+logistic regression
+(one hidden layer, $\ell_1=0.1$, $\ell_2=0.1$, 500 maximum iterations).
+%
+This 2-step approaches were compared against semi-supervised low-rank
+logistic regression
+(SS-LR LogReg, two hidden layers, $\ell_1=0.1$, $\ell_2=0.1$,
+500 maximum iterations, $\lambda=0.75$).
+Importantly, all trained classification models were tested
+on a large, unseen test set (20\% of data) in the presented analyses.
+%
+Across choices for $n$, SS-LR LogReg
+achieved more than 96\% out-of-sample accuracy, whereas
+supervised learning based on ICA and SPCA loadings
+ranged from 37,53\% to 87,28\%
+and
+32,19\% to 83,97\%, respectively (Table \ref{table_one}).
+%
+These explorations attest to the advantage of directly searching for
+classification-relevant structure in the fMRI data,
+rather than solving supervised and unsupervised problems independently.
+This effect was particularly pronounced when assuming few hidden dimensions.
+
+\begin{table}[h]
+ \centering
+ \resizebox{0.4\textwidth}{!}{%
+ \begin{tabular}{lccc}
+ \hline
+ \textit{n}               & ICA + LogReg & SPCA + LogReg & SS-LR LogReg \\ \hline
+ \multicolumn{1}{l|}{5}   & 37,53 \%     & 32,19 \%      & 96,50 \%     \\
+ \multicolumn{1}{l|}{20}  & 80,98 \%     & 78,15 \%      & 97,33 \%     \\
+ \multicolumn{1}{l|}{50}  & 84,19 \%     & 83,97 \%      & 97,69 \%     \\
+ \multicolumn{1}{l|}{100} & 87,28 \%     & 82,19 \%      & 97,80 \%     \\ \hline
+ \end{tabular}
+ }
+ \vspace{-0.2cm}
+ \caption{Serial versus parallel dimensionality reduction and classification}
+ \label{table_one}
+ \end{table}
+
+\paragraph{Model performance.}
 \begin{itemize}
@@
 \end{itemize}
-
-the features of the classes show that we captured activaiton properties,
-not particpant or paradigm properties
-
-out of sample performance: extract features of general relevance
-beyond the supervised problem at hand
-
-support recovery:
-Low-rank logistic regression
-a) puts higher absolute weights on relevant structure
-b) lower ones on irrelevant structure,
-and
-c) yields BOLD-typical local contiguity.
-
-\paragraph{Classification performance}
 \begin{itemize}
@@
   decomposition derived in semi-supervised fashion}
+
+out of sample performance: extract features of general relevance
+beyond the supervised problem at hand
 \end{itemize}
 
-\paragraph{Support recovery -> feature identification}
+\begin{table}[h]
+  \centering
+  \resizebox{1.00\textwidth}{!}{%
+\begin{tabular}{l|cccc|llllcccc|llll|}
+  \hline
+                            & \multicolumn{4}{c|}{\textit{n = 5}}                                                                                                                & \multicolumn{4}{c|}{\textit{n = 20}}                                                                                                                       & \multicolumn{4}{c}{\textit{n = 100}}                                                       \\ \hline
+  \textit{}                 & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.50$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1.00$} & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.50$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1.00$}   & $\lambda=0.25$   & $\lambda=0.50$   & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1.00$} \\ \hline
+  Out-of-sample\\accuracy   & \textit{88,90\%}                   & 95,13\%                            & \textbf{96,49\%}                   & 95,72\%                             & 97,44\%                              & \textbf{97,80\%}                     & 97,33\%                              & \textit{97,32\%}                      & 97,21\%          & \textit{97,03\%} & \textbf{97,80\%} & 97,38\%                            \\
+  Precision (mean)          & \textit{86,98\%}                   & 94,86\%                            & \textbf{96,28\%}                   & 95,36\%                             & \textbf{97,38\%}                     & 97,06\%                              & \textit{96,98\%}                     & 97,02\%                               & 96,90\%          & \textit{96,47\%} & \textbf{97,50\%} & 96,91\%                            \\
+  Recall (mean)             & \textit{88,27\%}                   & 95,19\%                            & \textbf{96,57\%}                   & 95,70\%                             & \textbf{97,51\%}                     & 97,50\%                              & \textit{97,35\%}                     & 97,36\%                               & 97,20\%          & \textit{97,15\%} & \textbf{97,88\%} & 97,37\%                            \\
+  F1 score (mean)           & \textit{86,64\%}                   & 94,89\%                            & \textbf{96,38\%}                   & 95,43\%                             & \textbf{97,42\%}                     & 97,22\%                              & \textit{97,11\%}                     & 97,13\%                               & 97,00\%          & \textit{96,71\%} & \textbf{97,65\%} & 97,07\%                            \\ \hline
+  Out-of-dataset \\accuracy & 60,83\%                            & \textit{54,30\%}                   & 60,69\%                            & \textbf{62,92\%}                    & 79,72\%                              & \textbf{81,94\%} & 79,72\%           & \textit{79,44\%}                     & \textbf{82,08\%} & 81,66\%            & 81,25\%          & \textit{75,83\%}                   \\ \hline
+  \end{tabular}
+
+}
+\end{table}
+
+\paragraph{Feature identification.}
 \begin{itemize}
@@
   of latent components}
+
+support recovery:
+Low-rank logistic regression
+a) puts higher absolute weights on relevant structure
+b) lower ones on irrelevant structure,
+and
+c) yields BOLD-typical local contiguity.
+
+the features of the classes show that we captured activaiton properties,
+not particpant or paradigm properties
 \end{itemize}
 
-\paragraph{Anecdotal observations}
-\begin{itemize}
-  \item{Modifications of the modek that did not improve the generalization performance:
-  drop-out, input corruption (p=0.1/0.3/0.5), RELUs}
-  \item{no penalty or only L1 or L2 penalty generally yielded worse
-  generalization performance}
-  \item{``pretraining'' of the bottleneck by settings weights to
-  preliminary PCA, SPCA or ICA weights did not give better accuracies,
-  neither for autoencoder pretraining }
-  \item{present approach better than first PCA, SPCA or ICA and
-  subsequently Logreg on loadings, even if each on the whole data}
-  \item{introducing a
-  nonlinearity (sigmoid, tanh, softplus) into the system
-  did not improve predictive accuracy but
-  elastic did -> most useful decomposition has characteristics of a SPCA
-  and not PCA or ICA}
-  \item{introduced an addition layer with many units (50/100) did not
-  improve generalization}
-\end{itemize}
+\paragraph{Miscellaneous observations.}
+For the sake of completeness,
+we informally report modifications of the model
+that did not improve the generalization performance.
+%
+Introducing stochasticity into model learning by 
+a) input corruption of $\mathbf{X_{task}}$ or
+b) drop-out at the bottleneck $\mathbf{W_0}$ using binomial
+distributions ($p=0.1, 0.3, 0.5$) deteriorated model performance
+in all instances. 
+%
+Adding c) rectified linear units (ReLU) to $\mathbf{W_0}$ or
+other commonly used nonlinearities (d) sigmoid, e) softplus,
+f) hyperbolic tangent) all led to smaller classification accuracies.
+%
+Further, g) ``pretraining'' of the bottleneck $\mathbf{W_0}$
+(i.e., non-random initialization) by 
+either corresponding SPCA or ICA loadings did not exhibit improved accuracies,
+neither did h) autoencoder pretraining.
+%
+Moreover,
+introducing an additional i) overcomplete layer (100 units)
+after the bottleneck was not advantageous.
+%
+Finally, imposing imposing either j) only $\ell_1$-penalty or 
+k) only $\ell_2$-penalty was disadvantageous in all tested cases,
+which favored elasticnet regularization chosen in the above analyses.
 

### 2015-05-30 paper_nips/paper.tex
Added explicit description of method to abstract

@@
 \usepackage{booktabs}
+\usepackage{ulem}
 
@@
 
+\newcommand{\suggestadd}[1]{{\color{blue} #1}}
+\newcommand{\suggestremove}[1]{{\color{red} \sout{#1}}}
+
 \nipsfinalcopy % Uncomment for camera-ready version
@@
 %
+% XXX Michael: Added explicit outline of the model, description can be 
+% XXX improved
 We therefore propose to optimize
 representation and task classification in
-an identical statistical learning problem.
+\suggestremove{an identical} \suggestadd{a unified} statistical learning 
+problem\suggestadd{, by introducing a multinomial logistic regression with
+joint rank constraint, which we regularize by forcing representational
+capacity in the form of an autoencoder}.
 %

### 2015-05-30 paper_nips/paper.tex
suggest to remove bioinformatics tag
suggest BOLD signal variation instead of BOLD signal

@@
 
+% XXX Michael: Suggesting removal of bioinformatics tag,
+% XXX see http://en.wikipedia.org/wiki/Bioinformatics
 \textbf{keywords}: dimensionality reduction; semi-supervised learning;
-bioinformatics; fMRI; systems neuroscience
+\suggestremove{bioinformatics;} fMRI; systems neuroscience
 
@@
 during ongoing random thought without controlled environmental stimulation.
-In fact, a large proportion of the BOLD signal is known
+% XXX Michael: Added variation/variance, because otherwise that is trivial:
+In fact, a large proportion of the BOLD signal \suggestadd{variation} is 
+known
 not to correlate with a particular behavior, stimulus, or experimental task

### 2015-05-30 paper_nips/paper.tex
rephrase

@@
 by experimental tasks \cite{friston97}.
-As a last example, always more neuroimaging studies model
+% XXX Michael more and more or ever more:
+As a last example, \suggestremove{always} \suggestadd{more and} more 
+neuroimaging studies model
 experimental tasks by training classification algorithms on brain signals

### 2015-05-30 paper_nips/paper.tex
possible rephrasing
rest of intro

@@
 This ignores the fact that the BOLD
-signal exhibits spatially distributed activity patterns.
+signal exhibits spatially distributed \suggestremove{activity} patterns
+\suggestadd{of coherent activity}.
 %
 Second, existing supervised neuroimaging analyses
-do not acknowledge the abundance
+\suggestremove{do not acknowledge} \suggestadd{cannot exploit} the abundance
 of easily acquired resting-state data \cite{biswaldiscovery}.
-These may allow better sampling the manifold of brain states
+These may allow better \suggestremove{sampling} \suggestadd{discovery of} 
+the manifold of brain states
 due to the high task-rest similarities of neural activity patterns,
@@
 
+% XXX Michael: This parenthesis structure should probably be unravelled into
+% XXX a sentence.
 Both these biological properties can be conjointly exploited in a mixed
- (i.e., use task and rest data)
-low-rank (i.e., perform network decomposition)
+ (i.e., using task and rest data)
+low-rank (i.e., performing network decomposition)
 approach.
@@
 supervised classification formally yields a semi-supervised learning
+% XXX Michael: No, it doesn't necessarily. You can do that with labeled data
 framework. The most relevant
@@
 Autoencoders
-are one-layered learning models that condense the input data to
-local and global representations
-by improving reconstruction from them.
+are \suggestremove{one-layered} learning models 
+that condense the input data to
+\suggestremove{local and global} \suggestadd{(typically lower-dimensional)} 
+representations
+by \suggestremove{improving} \suggestadd{optimizing}
+reconstruction from them.
 %
-They behave like a PCA
+% XXX Michael: In fact, they behave like a non-centered PCA, 
+% XXX which amounts to just doing an SVD
+% XXX which btw is NOT a convex optimization problem, although easy to 
+% XXX calculate and with a global optimum.
+% XXX Autoencoders are also not one-layered by default. Let's make sure that
+% XXX we don't trip up by getting the technicalities wrong and getting 
+% XXX reviewers annoyed at us because of it.
+They behave like a \suggestadd{(truncated, non-centered)} PCA
 in case of one linear hidden layer and a squared error loss
 \cite{baldi1989neural}.
-This architecture yields a convex optimization objective
+This architecture yields a \suggestadd{non-}convex optimization objective
 with unique global minimum.
@@
 and adding a nonlinear convex transformation \cite{le2011ica}.
+% XXX What convex transformation? Why is that relevant? Convexity is only
+% XXX relevant for the loss
 These authors further demonstrated that ICA, sparse autoencoders, and 
@@
 Hence, autoencoders can flexibly project the neuroimaging data
-onto the main axes of variation and thus
+onto the main axes of variation and thus % XXX Michael <- that's PCA
 reverse-engineer properties of the underlying
 neural processes \cite{olshausen96}.
+% XXX Michael ^ I don't get that last part nor the citation.
 
 In the present investigation,
-a linear autoencoder will be fed by (unlabeled) rest data and
-integrated as bottleneck
-into a low-rank logistic regression fed by (labeled) task data.
+a linear autoencoder will be \suggestremove{fed by} \suggestadd{fit to} 
+(unlabeled) rest data and
+integrated as \suggestadd{a rank-reducing} bottleneck
+into a \suggestremove{low-rank} \suggestadd{multinomial} 
+logistic regression \suggestremove{fed by} \suggestadd{fit to} 
+(labeled) task data.
 We can then solve the unsupervised data representation and the
@@
 From the perspective of dictionary learning, the first layer represents
-the discovered set of basis functions
+% XXX Michael: This is the crucial difference: The first layer is an analysis
+% XXX operation: it extracts information in some way. It may be extracting
+% XXX activations for basis functions, in which case it would be the pseudo-
+% XXX inverse of those basis functions. But unless they are orthogonal, these
+% XXX weights definitely don't represent basis functions themselves. It is 
+% XXX the second layer of the autoencoder that represents basis functions. If
+% XXX weights are tied, they are one and the same thing, with interesting
+% XXX implications. But in general, only the second layer represents a basis.
+\suggestadd{projectors to} the discovered set of basis functions
 whose linear combinations are learned
 by the second layer \cite{olshausen96}.
+% XXX Michael: ^ This needs to be changed around: This first layer obtains
+% XXX activations, the second layer weights are basis functions which are
+% XXX multiplied to these activations
 %
 Neurobiologically, this allows 
-delineating a low-dimensional manifold of brain network patterns and then 
+delineating a low-dimensional \suggestremove{manifold}
+\suggestadd{vector space} of brain network patterns and then 
 distinguishing mental tasks
@@
 put probability mass on the most neurobiologically
+% XXX Michael: ^ MEH, we are not really in a probabilistic setting ...
 valid models in the hypothesis set.
@@
 have mostly been studied in isolation.
+% XXX Michael: Taken together, the important modes of variation in brain dynamics and the neural correlates subserving mental operations have mostly been studied in isolation?
+Are they taken together, or isolated?
+:)
 We provide a principled computational framework to link these previously

### 2015-05-30 paper_nips/paper.tex
field of view units

@@
 32 channel head coil (TR=720ms, TE=33.1ms, flip angle=52°, BW=2290Hz/Px,
-in-plane FOV=$280\times180$mm, 72 slices, 2.0mm isotropic voxels).
+in-plane FOV=$280\textrm{mm}\times180\textrm{mm}$, 72 slices, 2.0mm 
+isotropic voxels).
 The ``minimally preprocessed'' pipeline includes
@@
 during a period of 15min (TR=720 ms, TE=33.1 ms, flip angle=52°,
-FOV=$280\times180$mm, and 2.0mm isotropic voxels). Besides run duration,
+FOV=$280\textrm{mm}\times180\textrm{mm}$, and 2.0mm isotropic voxels). 
+Besides run duration,
 the task acquisitions were identical to the resting-state fMRI acquisitions
@@
 channel head coil (TR=2400ms, TE=30ms, flip angle=60°, in-plane
-FOV=$19.2\times19.2$cm, 40 slices, 3.0mm isotropic voxels).
+FOV=$192\textrm{mm}\times192\textrm{mm}$, 40 slices, 3.0mm isotropic voxels).
 Standard preprocessing was performed with Nipype \cite{gorgo11}, including

### 2015-05-30 paper_nips/paper.tex
removed heavy subscripts from biases by naming some of them c

@@
 slice timing, motion correction, alignment, and spatial normalization.
-activity maps were spatially smoothed by
+Activity maps were spatially smoothed by
 a Gaussian kernel of 5mm (FWHM).
@@
   \begin{split}
-    \mathbf{z} &= \mathbf{W_0} \mathbf{x} + \mathbf{b_{W_0}} \\
-    \mathbf{x'} &= \mathbf{W_1} \mathbf{z} + \mathbf{b_{W_1}}
+    \mathbf{z} &= \mathbf{W_0} \mathbf{x} + \mathbf{b_0} \\
+    \mathbf{x'} &= \mathbf{W_1} \mathbf{z} + \mathbf{b_1}
   \end{split}
@@
 
-where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79,941$
+where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79{,}941$
 voxel values from each
 rest map,
-$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$ hidden dimensions (i.e.,
-distributed neural activity patterns), and 
+$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$\suggestremove{ hidden dimensions 
+(i.e.,
+distributed neural activity patterns)}
+\suggestadd{-dimensional hidden state}, and 
 $\mathbf{x' \in \mathbb{R}^{d}}$
@@
 output space (decoder).
-$\mathbf{b_{W_0}}$ and $\mathbf{b_{W_1}}$ are bias vectors.
+$\mathbf{b_0}$ and $\mathbf{b_1}$ are bias vectors.
 Note that $\mathbf{W_0}$ and $\mathbf{W_1}$ are tied,
 $\mathbf{W_0} = \mathbf{W_1^T}$.
-The optimal model parameters $\mathbf{W_0, b_{W_0}, b_{W_1}}$ are found by
+The optimal model parameters $\mathbf{W_0, b_0, b_1}$ are found by
 minimizing the imperfection of reconstruction according to squared error
@@
   \begin{split}
-    P(Y=i|\mathbf{x, V_0,V_1,b_{V_0}, b_{V_1}}) &= softmax_i(\mathbf{V_1 (V_0 x + b_{V_0}) + b_{V_1}}) \\
+    P(Y=i|\mathbf{x, V_0,V_1,c_0, c_1}) &= softmax_i(\mathbf{V_1 (V_0 x + c_0) + c_1}) \\
     &= \frac {e^{\mathbf{V_{1i} (V_{0i} x + b_{V_{0i}}) + b_{V_{1i}}}}} {\sum_j e^{\mathbf{V_{1j} (V_{0j} x + b_{V_{0j}}) + b_{V_{1j}}}}}            
@@
 onto hyperplanes that reflect $c$ class probabilities.
-$\mathbf{b_{V_0}}$ and $\mathbf{b_{V_1}}$ are corresponding
+$\mathbf{c_0}$ and $\mathbf{c_1}$ are corresponding
 bias vectors.
@@
   \begin{split}
-    {\mathcal{L_{LR}}}(\mathbf{x, y}) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|\mathbf{x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1}})
+    {\mathcal{L_{LR}}}(\mathbf{x, y}) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|\mathbf{x^{(i)}; V_1 (V_0 x + c_0) + c_1})
 \end{split}
@@
 In so doing, we search for the combined model parameters
-$\theta=\{\mathbf{V_0,V_1,b_{V_0}, b_{V_1}, b_{W_0}, b_{W_1}}\}$
+$\theta=\{\mathbf{V_0,V_1,c_0, c_1, b_0, b_1}\}$
 with respect to the (unsupervised) reconstruction error and the

### 2015-05-30 paper_nips/paper.tex
rearranged some math

@@
 The optimal model parameters $\mathbf{W_0, b_0, b_1}$ are found by
-minimizing the imperfection of reconstruction according to squared error
+minimizing the expected squared reconstruction error
 
-\begin{eqnarray}
-  {\mathcal{L_{AE}}}(\mathbf{x}, \mathbf{x'}) = || \mathbf{x} - \mathbf{x'} ||^2
-\end{eqnarray}
-
-This reconstruction error criterion equates with
-maximizing a lower bound on the mutual information between
-input and the learned representation.
-Nonlinearities were not applied on the
+\begin{equation*}
+  \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
+  \mathbb E\left[\| \mathbf{x} - f_{AE}(\mathbf{x}) \|^2\right],
+\end{equation*}
+with \(f_{AE}(\mathbf x) = %
+\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}\)
+the autoencoder.
+% XXX Michael: I don't know to what extent the following makes sense or is
+% XXX correct, so I am commenting it
+%% This reconstruction error criterion equates with
+%% maximizing a lower bound on the mutual information between
+%% input and the learned representation.
+Nonlinearities were not applied to the
 activations in the first layer.

### 2015-05-30 paper_nips/paper.tex
Add a comment on analysis and synthesis and tied weights

@@
 $\mathbf{b_0}$ and $\mathbf{b_1}$ are bias vectors.
-Note that $\mathbf{W_0}$ and $\mathbf{W_1}$ are tied,
-$\mathbf{W_0} = \mathbf{W_1^T}$.
 The optimal model parameters $\mathbf{W_0, b_0, b_1}$ are found by
@@
 %% input and the learned representation.
+\suggestadd{
+Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
+$\mathbf{W_0} = \mathbf{W_1^T}$. In particular, this means that the weights
+learned are forced to take a two-fold function: That of \textit{analysis} 
+and that of \textit{synthesis}. Whereas the first layer \textit{analyzes}
+the data in order to obtain the cleanest latent representation, the second
+layer represents building blocks from which to \textit{synthesize} the data
+using the latent activations. Tying these functionalities makes the analysis
+side interpretable and pulls all non-zero singular values towards 1. 
 Nonlinearities were not applied to the
 activations in the first layer.
-
+}
 \paragraph{Reduced-rank logistic regression.}

### 2015-05-30 paper_nips/paper.tex
minor correction

@@
 
-\begin{equation*}
+\begin{equation}
   \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
   \mathbb E\left[\| \mathbf{x} - f_{AE}(\mathbf{x}) \|^2\right],
-\end{equation*}
+\end{equation}
 with \(f_{AE}(\mathbf x) = %
@@
 $\mathbf{W_0} = \mathbf{W_1^T}$. In particular, this means that the weights
-learned are forced to take a two-fold function: That of \textit{analysis} 
-and that of \textit{synthesis}. Whereas the first layer \textit{analyzes}
+learned are forced to take a two-fold function: That of signal 
+\textit{analysis} and that of signal \textit{synthesis}.
+Whereas the first layer \textit{analyzes}
 the data in order to obtain the cleanest latent representation, the second

### 2015-05-30 paper_nips/paper.tex
update

@@
 \DeclareMathOperator{\proj}{proj}
-\DeclareMathOperator{\soft}{soft}
+\DeclareMathOperator{\softmax}{softmax}
 \DeclareMathOperator{\prox}{prox}
@@
 It gives the probability of an input $\mathbf{x}$ to belong
-to a class $i \in \mathbb{N}$
+to a class $i \in \{1, \dots, l\}$
 
-\begin{eqnarray}
+\begin{equation}
   \begin{split}
-    P(Y=i|\mathbf{x, V_0,V_1,c_0, c_1}) &= softmax_i(\mathbf{V_1 (V_0 x + c_0) + c_1}) \\
-    &= \frac {e^{\mathbf{V_{1i} (V_{0i} x + b_{V_{0i}}) + b_{V_{1i}}}}} {\sum_j e^{\mathbf{V_{1j} (V_{0j} x + b_{V_{0j}}) + b_{V_{1j}}}}}            
+    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_{RL}(\mathbf x))
   \end{split}
   \label{eq:lr}
-\end{eqnarray}
-
-where the matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
+\end{equation}
+where \(f_{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
+multinomial logits and \(\softmax_i(x) = \exp(x_i)/\sum_j\exp(x_j)\).
+The matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
 transforms the input $\mathbf{x \in \mathbb{R}^{d}}$
 into $n$ latent components
-and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxc}}$
+and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxl}}$
 projects the latent components
-onto hyperplanes that reflect $c$ class probabilities.
+onto hyperplanes that reflect $l$ label probabilities.
 $\mathbf{c_0}$ and $\mathbf{c_1}$ are corresponding
@@
 
-\begin{eqnarray}
-  \begin{split}
-    {\mathcal{L_{LR}}}(\mathbf{x, y}) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|\mathbf{x^{(i)}; V_1 (V_0 x + c_0) + c_1})
-\end{split}
+\begin{equation}
+    \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
+\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1})
 \label{eq:lr_loss}
-\end{eqnarray}
+\end{equation}
 
@@
 and the low-rank logistic regression
-are linked one two levels. First, their transformation matrices mapping from
-input to the latent space are identical
+are linked on two levels. First, their transformation matrices mapping from
+input to the latent space are tied
 \begin{eqnarray}
-  \mathbf{V_0} = \mathbf{W_0}
+  \mathbf{V_0} = \mathbf{W_0}.
 \end{eqnarray}
@@
 
-\begin{eqnarray}
-  \begin{split}
-{arg\,min}_\theta \hspace{0.2cm} {\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
-+ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \ell_1(\theta) + \ell_2(\theta)
+% XXX Michael: In resolving conflict I opted for my change. Feel free to revert if that was intended.
+\begin{equation}
+{\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
++ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega(\theta).
   \label{eq:loss_equ}
-\end{split}
-\end{eqnarray}
+\end{equation}
 
@@
 loss terms to the same order of magnitude.
-For parameter shrinkage, we
-impose a combination of $\ell_1$ and $\ell_2$ penalties
-$\forall p \in \theta$ (i.e., elasticnet).
+\(\Omega(\theta)\) represents a regularization, for which we choose the
+ElasticNet type, i.e. a combination of $\ell_1$ and $\ell_2$ penalties
+$\forall p \in \theta$.
+
+% XXX Michael: To sum up the math in this paragraph: We need to decide what
+% XXX the \mathcal L are supposed to mean - sample wise loss or expected loss
+% XXX or some global deterministic loss (the latter would be bad, because we)
+% XXX are not optimizing it. Right now it looks like a sample-wise loss.
 
 \paragraph{Optimization.}
-The common objective was approximated 
+The common objective was \suggestremove{approximated} \suggestadd{optimized}
 % XXX why 'approximated' ?
 in the neuroimaging data
-by updating parameters by derivates 
+by \suggestremove{updating parameters by derivates}
+\suggestadd{gradient descent in the parameters}
 % XXX derivates ???

### 2015-05-30 paper_nips/paper.tex
done first pass

@@
 are given by
+% XXX Michael: It is crucial, but nowhere mentioned, that the optimization is
+% XXX done in an alternating manner. The whole structure relies on this, by
+% XXX construction it is vulnerable to optimizing too much towards labels or
+% XXX rest at a time. I also identify this as one of the major weak points
+% XXX of the method, so it needs to be stated in a way that is not easily
+% XXX attackable.
+% XXX Maybe we could also talk less about RMSprop.
 
@@
 \end{eqnarray}
+% XXX Michael: Why, in the formula, do we have two different notations for
+% XXX partial derivatives? \partial and \nabla
 
+% XXX Michael: Probably remove the explanations following and just state values
 where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
@@
 Its purpose is not to prevent overfitting but to introduce
+% XXX Michael: I use f earlier up, is this f here necessary?
 prior knowledge on known properties of the unknown target function $f$.
@@
 and
-\textit{theano} for automatic, numerically stable
+\textit{Theano} for automatic, numerically stable
 differentiation of symbolic computation graphs
@@
 \textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning
-}
+{\color{blue} Michael: Does this also happen when you double the task data?
+i.e. rest:task 3:1, but duplicate the task data, yielding essentially 3:2?
+I assume this will work, which again points to the instability of the network architecture}}
 }

### 2015-05-30 paper_nips/paper.tex
added 2 tables

--- 2015-05-30 paper_nips/paper.tex
@@
 \usepackage{booktabs}
-\usepackage{ulem}
 
@@
 
-\newcommand{\suggestadd}[1]{{\color{blue} #1}}
-\newcommand{\suggestremove}[1]{{\color{red} \sout{#1}}}
-
 \nipsfinalcopy % Uncomment for camera-ready version
@@
 %
-% XXX Michael: Added explicit outline of the model, description can be 
-% XXX improved
 We therefore propose to optimize
 representation and task classification in
-\suggestremove{an identical} \suggestadd{a unified} statistical learning 
-problem\suggestadd{, by introducing a multinomial logistic regression with
-joint rank constraint, which we regularize by forcing representational
-capacity in the form of an autoencoder}.
+an identical statistical learning problem.
 %
@@
 
-% XXX Michael: Suggesting removal of bioinformatics tag,
-% XXX see http://en.wikipedia.org/wiki/Bioinformatics
 \textbf{keywords}: dimensionality reduction; semi-supervised learning;
-\suggestremove{bioinformatics;} fMRI; systems neuroscience
+bioinformatics; fMRI; systems neuroscience
 
@@
 during ongoing random thought without controlled environmental stimulation.
-% XXX Michael: Added variation/variance, because otherwise that is trivial:
-In fact, a large proportion of the BOLD signal \suggestadd{variation} is 
-known
+In fact, a large proportion of the BOLD signal is known
 not to correlate with a particular behavior, stimulus, or experimental task
@@
 by experimental tasks \cite{friston97}.
-% XXX Michael more and more or ever more:
-As a last example, \suggestremove{always} \suggestadd{more and} more 
-neuroimaging studies model
+As a last example, always more neuroimaging studies model
 experimental tasks by training classification algorithms on brain signals
@@
 This ignores the fact that the BOLD
-signal exhibits spatially distributed \suggestremove{activity} patterns
-\suggestadd{of coherent activity}.
+signal exhibits spatially distributed activity patterns.
 %
 Second, existing supervised neuroimaging analyses
-\suggestremove{do not acknowledge} \suggestadd{cannot exploit} the abundance
+do not acknowledge the abundance
 of easily acquired resting-state data \cite{biswaldiscovery}.
-These may allow better \suggestremove{sampling} \suggestadd{discovery of} 
-the manifold of brain states
+These may allow better sampling the manifold of brain states
 due to the high task-rest similarities of neural activity patterns,
@@
 
-% XXX Michael: This parenthesis structure should probably be unravelled into
-% XXX a sentence.
 Both these biological properties can be conjointly exploited in a mixed
- (i.e., using task and rest data)
-low-rank (i.e., performing network decomposition)
+ (i.e., use task and rest data)
+low-rank (i.e., perform network decomposition)
 approach.
@@
 supervised classification formally yields a semi-supervised learning
-% XXX Michael: No, it doesn't necessarily. You can do that with labeled data
 framework. The most relevant
@@
 Autoencoders
-are \suggestremove{one-layered} learning models 
-that condense the input data to
-\suggestremove{local and global} \suggestadd{(typically lower-dimensional)} 
-representations
-by \suggestremove{improving} \suggestadd{optimizing}
-reconstruction from them.
+are one-layered learning models that condense the input data to
+local and global representations
+by improving reconstruction from them.
 %
-% XXX Michael: In fact, they behave like a non-centered PCA, 
-% XXX which amounts to just doing an SVD
-% XXX which btw is NOT a convex optimization problem, although easy to 
-% XXX calculate and with a global optimum.
-% XXX Autoencoders are also not one-layered by default. Let's make sure that
-% XXX we don't trip up by getting the technicalities wrong and getting 
-% XXX reviewers annoyed at us because of it.
-They behave like a \suggestadd{(truncated, non-centered)} PCA
+They behave like a PCA
 in case of one linear hidden layer and a squared error loss
 \cite{baldi1989neural}.
-This architecture yields a \suggestadd{non-}convex optimization objective
+This architecture yields a convex optimization objective
 with unique global minimum.
@@
 and adding a nonlinear convex transformation \cite{le2011ica}.
-% XXX What convex transformation? Why is that relevant? Convexity is only
-% XXX relevant for the loss
 These authors further demonstrated that ICA, sparse autoencoders, and 
@@
 Hence, autoencoders can flexibly project the neuroimaging data
-onto the main axes of variation and thus % XXX Michael <- that's PCA
+onto the main axes of variation and thus
 reverse-engineer properties of the underlying
 neural processes \cite{olshausen96}.
-% XXX Michael ^ I don't get that last part nor the citation.
 
 In the present investigation,
-a linear autoencoder will be \suggestremove{fed by} \suggestadd{fit to} 
-(unlabeled) rest data and
-integrated as \suggestadd{a rank-reducing} bottleneck
-into a \suggestremove{low-rank} \suggestadd{multinomial} 
-logistic regression \suggestremove{fed by} \suggestadd{fit to} 
-(labeled) task data.
+a linear autoencoder will be fed by (unlabeled) rest data and
+integrated as bottleneck
+into a low-rank logistic regression fed by (labeled) task data.
 We can then solve the unsupervised data representation and the
@@
 From the perspective of dictionary learning, the first layer represents
-% XXX Michael: This is the crucial difference: The first layer is an analysis
-% XXX operation: it extracts information in some way. It may be extracting
-% XXX activations for basis functions, in which case it would be the pseudo-
-% XXX inverse of those basis functions. But unless they are orthogonal, these
-% XXX weights definitely don't represent basis functions themselves. It is 
-% XXX the second layer of the autoencoder that represents basis functions. If
-% XXX weights are tied, they are one and the same thing, with interesting
-% XXX implications. But in general, only the second layer represents a basis.
-\suggestadd{projectors to} the discovered set of basis functions
+the discovered set of basis functions
 whose linear combinations are learned
 by the second layer \cite{olshausen96}.
-% XXX Michael: ^ This needs to be changed around: This first layer obtains
-% XXX activations, the second layer weights are basis functions which are
-% XXX multiplied to these activations
 %
 Neurobiologically, this allows 
-delineating a low-dimensional \suggestremove{manifold}
-\suggestadd{vector space} of brain network patterns and then 
+delineating a low-dimensional manifold of brain network patterns and then 
 distinguishing mental tasks
@@
 put probability mass on the most neurobiologically
-% XXX Michael: ^ MEH, we are not really in a probabilistic setting ...
 valid models in the hypothesis set.
@@
 have mostly been studied in isolation.
-% XXX Michael: Taken together, the important modes of variation in brain dynamics and the neural correlates subserving mental operations have mostly been studied in isolation?
-Are they taken together, or isolated?
-:)
 We provide a principled computational framework to link these previously
@@
 32 channel head coil (TR=720ms, TE=33.1ms, flip angle=52°, BW=2290Hz/Px,
-in-plane FOV=$280\textrm{mm}\times180\textrm{mm}$, 72 slices, 2.0mm 
-isotropic voxels).
+in-plane FOV=$280\times180$mm, 72 slices, 2.0mm isotropic voxels).
 The ``minimally preprocessed'' pipeline includes
@@
 during a period of 15min (TR=720 ms, TE=33.1 ms, flip angle=52°,
-FOV=$280\textrm{mm}\times180\textrm{mm}$, and 2.0mm isotropic voxels). 
-Besides run duration,
+FOV=$280\times180$mm, and 2.0mm isotropic voxels). Besides run duration,
 the task acquisitions were identical to the resting-state fMRI acquisitions
@@
 channel head coil (TR=2400ms, TE=30ms, flip angle=60°, in-plane
-FOV=$192\textrm{mm}\times192\textrm{mm}$, 40 slices, 3.0mm isotropic voxels).
+FOV=$19.2\times19.2$cm, 40 slices, 3.0mm isotropic voxels).
 Standard preprocessing was performed with Nipype \cite{gorgo11}, including
 slice timing, motion correction, alignment, and spatial normalization.
-Activity maps were spatially smoothed by
+activity maps were spatially smoothed by
 a Gaussian kernel of 5mm (FWHM).
@@
   \begin{split}
-    \mathbf{z} &= \mathbf{W_0} \mathbf{x} + \mathbf{b_0} \\
-    \mathbf{x'} &= \mathbf{W_1} \mathbf{z} + \mathbf{b_1}
+    \mathbf{z} &= \mathbf{W_0} \mathbf{x} + \mathbf{b_{W_0}} \\
+    \mathbf{x'} &= \mathbf{W_1} \mathbf{z} + \mathbf{b_{W_1}}
   \end{split}
@@
 
-where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79{,}941$
+where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79,941$
 voxel values from each
 rest map,
-$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$\suggestremove{ hidden dimensions 
-(i.e.,
-distributed neural activity patterns)}
-\suggestadd{-dimensional hidden state}, and 
+$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$ hidden dimensions (i.e.,
+distributed neural activity patterns), and 
 $\mathbf{x' \in \mathbb{R}^{d}}$
@@
 output space (decoder).
-$\mathbf{b_0}$ and $\mathbf{b_1}$ are bias vectors.
-The optimal model parameters $\mathbf{W_0, b_0, b_1}$ are found by
-minimizing the expected squared reconstruction error
+$\mathbf{b_{W_0}}$ and $\mathbf{b_{W_1}}$ are bias vectors.
+Note that $\mathbf{W_0}$ and $\mathbf{W_1}$ are tied,
+$\mathbf{W_0} = \mathbf{W_1^T}$.
+The optimal model parameters $\mathbf{W_0, b_{W_0}, b_{W_1}}$ are found by
+minimizing the imperfection of reconstruction according to squared error
 
-\begin{equation}
-  \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
-  \mathbb E\left[\| \mathbf{x} - f_{AE}(\mathbf{x}) \|^2\right],
-\end{equation}
-with \(f_{AE}(\mathbf x) = %
-\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}\)
-the autoencoder.
-% XXX Michael: I don't know to what extent the following makes sense or is
-% XXX correct, so I am commenting it
-%% This reconstruction error criterion equates with
-%% maximizing a lower bound on the mutual information between
-%% input and the learned representation.
-\suggestadd{
-Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
-$\mathbf{W_0} = \mathbf{W_1^T}$. In particular, this means that the weights
-learned are forced to take a two-fold function: That of signal 
-\textit{analysis} and that of signal \textit{synthesis}.
-Whereas the first layer \textit{analyzes}
-the data in order to obtain the cleanest latent representation, the second
-layer represents building blocks from which to \textit{synthesize} the data
-using the latent activations. Tying these functionalities makes the analysis
-side interpretable and pulls all non-zero singular values towards 1. 
-Nonlinearities were not applied to the
+\begin{eqnarray}
+  {\mathcal{L_{AE}}}(\mathbf{x}, \mathbf{x'}) = || \mathbf{x} - \mathbf{x'} ||^2
+\end{eqnarray}
+
+This reconstruction error criterion equates with
+maximizing a lower bound on the mutual information between
+input and the learned representation.
+Nonlinearities were not applied on the
 activations in the first layer.
-}
+
 \paragraph{Reduced-rank logistic regression.}
@@
   \begin{split}
-    P(Y=i|\mathbf{x, V_0,V_1,c_0, c_1}) &= softmax_i(\mathbf{V_1 (V_0 x + c_0) + c_1}) \\
+    P(Y=i|\mathbf{x, V_0,V_1,b_{V_0}, b_{V_1}}) &= softmax_i(\mathbf{V_1 (V_0 x + b_{V_0}) + b_{V_1}}) \\
     &= \frac {e^{\mathbf{V_{1i} (V_{0i} x + b_{V_{0i}}) + b_{V_{1i}}}}} {\sum_j e^{\mathbf{V_{1j} (V_{0j} x + b_{V_{0j}}) + b_{V_{1j}}}}}            
@@
 onto hyperplanes that reflect $c$ class probabilities.
-$\mathbf{c_0}$ and $\mathbf{c_1}$ are corresponding
+$\mathbf{b_{V_0}}$ and $\mathbf{b_{V_1}}$ are corresponding
 bias vectors.
@@
   \begin{split}
-    {\mathcal{L_{LR}}}(\mathbf{x, y}) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|\mathbf{x^{(i)}; V_1 (V_0 x + c_0) + c_1})
+    {\mathcal{L_{LR}}}(\mathbf{x, y}) = \frac{1}{N_{X_{task}}} \sum_{i=0}^{N_{X_{task}}} \log(P(Y=y^{(i)}|\mathbf{x^{(i)}; V_1 (V_0 x + b_{V_0}) + b_{V_1}})
 \end{split}
@@
 In so doing, we search for the combined model parameters
-$\theta=\{\mathbf{V_0,V_1,c_0, c_1, b_0, b_1}\}$
+$\theta=\{\mathbf{V_0,V_1,b_{V_0}, b_{V_1}, b_{W_0}, b_{W_1}}\}$
 with respect to the (unsupervised) reconstruction error and the
@@
 the HCP task data.
+
 \textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning

### 2015-05-30 nips3mm_h38.py

@@ -561,7 +561,7 @@ for n_comp in n_comps:  #
     plt.figure()
     for p in pkgs:
         lambda_param = np.float(re.search('lambda=(.{4})', p).group(1))
-        n_hidden = int(re.search('comp=(?P<comp>.{1,2})_', p).group('comp'))
+        n_hidden = int(re.search('comp=(?P<comp>.{1,3})_', p).group('comp'))
         if n_comp != n_hidden:
             continue
         
@@ -816,8 +816,9 @@ for n_comp in n_comps:
     plt.show()
     plt.savefig(op.join(WRITE_DIR, 'rec_inds_%icomps.png' % n_comp))
 
-# in-dataset f1 at lambda=0.5
-pkgs = glob.glob(RES_NAME + '/*lambda=1.00*dbg_prfs_.npy')
+# in-dataset f1
+# pkgs = glob.glob(RES_NAME + '/*lambda=1.00*dbg_prfs_.npy')
+pkgs = glob.glob(RES_NAME + '/*dbg_prfs_.npy')
 for n_comp in n_comps:
     plt.figure()
     p_ep = glob.glob(RES_NAME + '/*comp=%i*dbg_epochs_*.npy' % n_comp)
@@ -842,12 +843,15 @@ for n_comp in n_comps:
             cur_label += 'RSpca20'
         cur_label += '/'
         cur_label += 'separate decomp.' if 'decomp_separate' in p else 'joint decomp.'
+        cur_label = 'Logistic regression'
         for i in np.arange(38):
             plt.plot(
                 dbg_epochs_[:len(np.array(dbg_prfs_)[:, 2, i])],
                 np.array(dbg_prfs_)[:, 2, i],
                 label='task %i' % (i + 1))
-    plt.title('Classification performance for 38 tasks')
+    # plt.title('Low-rank logistic regression ($n=%i$, $\lambda=1.0$)' % n_comp)
+    plt.title('Logistic regression (voxel space)')
+    # plt.title('Classification performance for 38 tasks')
     # plt.legend(loc='lower right', fontsize=9)
     # plt.yticks(np.linspace(0., 1., 11))
     plt.ylabel('f1 score (test set)')

### 2015-05-30 nips3mm_serial.py

@@ -1,0 +1,343 @@
+"""
+HCP: semi-supervised network decomposition by low-rank logistic regression
+"""
+"""
+preliminary conclusion:
+- the combined AE and LR loss yields best classification results
+- Regarding rest data, 20-comp PCA work better than uncompressed data
+questions:
+- the AE cost is much bigger, is it ok to T.log() that?
+No because it renders the objective function non-convex.
+- compare to nonlinearity + PCA only
+- how to genreate examples from 100 last model paramters during the
+gradient descent?
+"""
+
+print __doc__
+
+import os
+import os.path as op
+import numpy as np
+import glob
+from scipy.linalg import norm
+import nibabel as nib
+from sklearn.grid_search import RandomizedSearchCV
+from sklearn.base import BaseEstimator
+from sklearn.preprocessing import StandardScaler
+from nilearn.input_data import NiftiMasker
+from sklearn.metrics import precision_recall_fscore_support
+from sklearn.metrics import confusion_matrix
+import theano
+import theano.tensor as T
+from matplotlib import pylab as plt
+print('Running THEANO on %s' % theano.config.device)
+from nilearn.image import concat_imgs
+import joblib
+import time
+
+RES_NAME = 'nips3mm_serial'
+WRITE_DIR = op.join(os.getcwd(), RES_NAME)
+if not op.exists(WRITE_DIR):
+    os.mkdir(WRITE_DIR)
+
+##############################################################################
+# load+preprocess data
+##############################################################################
+
+mask_img = 'grey10_icbm_3mm_bin.nii.gz'
+nifti_masker = NiftiMasker(mask_img=mask_img, smoothing_fwhm=False,
+                           standardize=False)
+nifti_masker.fit()
+mask_nvox = nifti_masker.mask_img_.get_data().sum()
+
+print('Loading data...')
+X_task, labels = joblib.load('preload_HT_3mm')
+
+labels = np.int32(labels)
+
+# contrasts are IN ORDER -> shuffle!
+new_inds = np.arange(0, X_task.shape[0])
+np.random.shuffle(new_inds)
+X_task = X_task[new_inds]
+y = labels[new_inds]
+
+from sklearn.cross_validation import StratifiedShuffleSplit
+folder = StratifiedShuffleSplit(y, n_iter=1, test_size=0.50)
+new_devs, inds_val = iter(folder).next()
+X_dev, X_val = X_task[new_devs], X_task[inds_val]
+y_dev, y_val = y[new_devs], y[inds_val]
+
+del X_task
+
[…]

### 2015-05-30 nips3mm_vanilla.py

@@ -272,10 +272,12 @@ class SSEncoder(BaseEstimator):
 # plot figures
 ##############################################################################
 
-def dump_comps(masker, compressor, components, threshold=2, fwhm=None):
+def dump_comps(masker, compressor, components, threshold=2, fwhm=None,
+               perc=None):
     from scipy.stats import zscore
     from nilearn.plotting import plot_stat_map
     from nilearn.image import smooth_img
+    from scipy.stats import scoreatpercentile
 
     if isinstance(compressor, basestring):
         comp_name = compressor
@@ -287,22 +289,32 @@ def dump_comps(masker, compressor, components, threshold=2, fwhm=None):
                                                      n_comp, i_c + 1))
         nii_raw = masker.inverse_transform(comp)
         nii_raw.to_filename(path_mask + '.nii.gz')
+        
+        comp_z = zscore(comp)
+        
+        if perc is not None:
+            cur_thresh = scoreatpercentile(np.abs(comp_z), per=perc)
+            path_mask += '_perc%i' % perc
+            print('Applying percentile %.2f (threshold: %.2f)' % (perc, cur_thresh))
+        else:
+            cur_thresh = threshold
+            path_mask += '_thr%.2f' % cur_thresh
+            print('Applying threshold: %.2f' % cur_thresh)
 
-        nii_z = masker.inverse_transform(zscore(comp))
+        nii_z = masker.inverse_transform(comp_z)
         gz_path = path_mask + '_zmap.nii.gz'
         nii_z.to_filename(gz_path)
-        plot_stat_map(gz_path, bg_img='colin.nii', threshold=threshold,
+        plot_stat_map(gz_path, bg_img='colin.nii', threshold=cur_thresh,
                       cut_coords=(0, -2, 0), draw_cross=False,
                       output_file=path_mask + 'zmap.png')
                       
         # optional: do smoothing
         if fwhm is not None:
             nii_z_fwhm = smooth_img(nii_z, fwhm=fwhm)
-            plot_stat_map(nii_z_fwhm, bg_img='colin.nii', threshold=threshold,
+            plot_stat_map(nii_z_fwhm, bg_img='colin.nii', threshold=cur_thresh,
                           cut_coords=(0, -2, 0), draw_cross=False,
                           output_file=path_mask +
[…]

### 2015-05-31 paper_nips/paper.tex
finished first results section

--- 2015-05-30 paper_nips/paper.tex
@@
 
-\bigskip
-\begin{figure}
-\begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/accuracies.png}
-\end{centering}
-\caption{\textbf{Generalization performance}
-Depicts in-sample (apply current model on training set, \textit{left column}),
-out-of-sample (current model on test set, \textit{middle column}),
-and out-of-dataset (compress other dataset using the
-learned decomposition for classification, \textit{right column}) accuracy.
-Curves for different choices of $\lambda$ as well as for ordinary
-logistic regression (\textit{yellowish})
-across iterations over the HCP task data (epochs).
-XXX: Message ??
-}
-\end{figure}
-
-\begin{figure}
-\begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/figure_weights_perc75.png}
-\end{centering}
-\caption{\textbf{Interpretability of classification weights}
-The voxel predictors corresponding to 5 exemplary
-(of 18 total) psychological tasks (\textit{rows})
-from the HCP dataset \cite{barch2013}.
-\textit{Left column:} ordinary logistic regression (same
-implementation but without bottleneck or autoencoder),
-\textit{middle column:} semi-supervised low-rank logistic regression
-($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
-\textit{right column:} voxel-wise average of whole-brain
-activity maps from each task.
-Low-rank logistic regression
-a) puts higher absolute weights on relevant structure
-b) lower ones on irrelevant structure,
-and
-c) yields BOLD-typical local contiguity.
-All values are z-scored and thresholded at the 75th percentile.
-}
-\end{figure}
-
-\begin{figure}
-% XXX: figure not readable
-\begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/semisup.png}
-\end{centering}
-\caption{\textbf{Relation between support recovery and out-of-dataset performance}
-\textit{Upper row:} the fit-for-purpose of the classification models
-is quantified by linear correlation between the
-z-scored average task image and the z-scored weights from
-ordinary (\textit{red}) and present (\textit{blue colors}) logistic
-regression ($\lambda=0.25, 0.50, 0.75, 1.00$).
-\textit{Lower row:} ordinary logistic regression is applied to the
-ARCHI dataset after projection into the latent dimensions learned from
-the HCP task data.
-\textbf{whenever I tried to have 3+ times more rest than task images results dropped dramatically...naive interpretation: if the rest information is too rich it overrules task patterns and hinders the supervised learning
-{\color{blue} Michael: Does this also happen when you double the task data?
-i.e. rest:task 3:1, but duplicate the task data, yielding essentially 3:2?
-I assume this will work, which again points to the instability of the network architecture}}
-}
-\end{figure}
-
 \section{Experimental Results}
@@
  \centering
- \resizebox{0.4\textwidth}{!}{%
+ \resizebox{0.5\textwidth}{!}{%
  \begin{tabular}{lccc}
@@
 \paragraph{Model performance.}
-\begin{itemize}
-\item{20 components: high bias/low variance
-100 compoentens: low bias/high variance}
-\end{itemize}
-\begin{itemize}
-  \item{This was evaluated by the prediction
-accuracy on a validation set (20\% of the training data) at each iteration
-over the ARCHI task data}
-  \item{the less lambda, the faster the convergence, with small differences
-  though; smaller differences with larger number of components}
-  \item{both 20 and 100 hidden components achieve 97\% out-of-sample
-  accuracy}
-  \item{ordinary logistic regression converges faster but with stagnation at
-  92\% out-of-sample accuracy}
-  \item{classification at chance 6\% in case of lambda==0 because 
-  only the unsupervised layer of the estimator is optimized}
-  \item{important: the thus only resting-state-trained decomposition achieves
-  the same usefulness in the ARCHI classification than the
-  decomposition derived in semi-supervised fashion}
-
-out of sample performance: extract features of general relevance
-beyond the supervised problem at hand
-\end{itemize}
+SS-LR LogReg was subsequently trained ($500$ epochs) across parameter choices
+for the hidden components ($n=5, 20, 100$) and
+the balance between autoencoder and logistic regression
+($\lambda=0.25, 0.5, 0.75, 1$).
+Assuming 5 latent directions of variation should yield models with
+higher bias and smaller variance then SS-LR LogReg with 100 latent directions.
+%
+Given the 18-class problem of HCP, setting $\lambda$ to $0$
+consistently yields generalization performance
+at chance-level (5,6\%) because
+only the unsupervised layer of the estimator is optimized.
+%
+At each epoch (i.e., iteration over the data),
+the out-of-sample performance of the trained classifier
+was assessed on 20\% of unseen HCP data.
+Additionally, the ``out-of-dataset performance'' of the learned decomposition
+was assessed by using it as dimensionality reduction of an
+independent labaled dataset (i.e., ARCHI) and conducting ordinary
+logistic regression on the ensuing component loadings.
 
@@
   \resizebox{1.00\textwidth}{!}{%
-\begin{tabular}{l|cccc|llllcccc|llll|}
+\begin{tabular}{l|cccc|cccc|cccc|cccc|}
   \hline
                             & \multicolumn{4}{c|}{\textit{n = 5}}                                                                                                            & \multicolumn{4}{c|}{\textit{n = 20}}                                                                                                                   & \multicolumn{4}{c}{\textit{n = 100}}                                                       \\ \hline
-  \textit{}                 & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.50$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1.00$} & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.50$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1.00$}   & $\lambda=0.25$   & $\lambda=0.50$   & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1.00$} \\ \hline
+  \textit{}                 & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
   Out-of-sample\\accuracy   & \textit{88,90\%}                   & 95,13\%                            & \textbf{96,49\%}                  & 95,72\%                          & 97,44\%                              & \textbf{97,80\%}                    & 97,33\%                              & \textit{97,32\%}                   & 97,21\%          & \textit{97,03\%} & \textbf{97,80\%} & 97,38\%                            \\
@@
   \end{tabular}
-
 }
+\vspace{-0.2cm}
+ \caption{Performance of SS-LR LogReg across model parameter choices}
+  \label{table_two}
 \end{table}
 
+We made several noteworthy observations.
+%
+First, the purely supervised estimator ($\lambda=1$) achieved in no
+instance the best accuracy, precision, recall or f1 scores on HCP data.
+Classification by low-rank logistic regression is therefore facilitated by
+imposing structure from the unlabeled rest data.
+%
+Second, the higher the number of latent components $n$,
+the higher the out-of-dataset performance with small choices for $\lambda$.
+This suggests that the presence of more rest-data-inspired hidden components
+yields to more effective feature representation in unrelated task data.
+%
+Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
+decomposition matrix ($\lambda=0$) resulted in
+noninferior out-of-dataset performance
+of X and Y, respectively (not shown in Table \ref{table_two}).
+This confirms that guiding model learning by task-unrelated structure
+extract features of general relevance
+beyond the supervised problem at hand.
+
+Depicts f1 scores for prediction of each of 38 classes across training epochs.
+Ordinary logistic regression operating in voxel space (\textit{left plot})
+converged faster but performed worse than
+low-rank logistic regression for few
+(\textit{middle plot}) and many (\textit{right plot}).
+Autoencoder or rest data were not used for these analyses
+($\lambda=1$).
+Hence, projecting the input data into a reduced space for classification
+yields higher class separability.
+
+\paragraph{Individual effects of dimensionality reduction and rest data.}
+We first quantified the gain of introducing a bottleneck layer
+disregaring the autoencoder.
+To this end, ordinary logstic regression was juxtaposed
+with LR LogRed at $\lambda=1$ (no autoencoder).
+For this experiment, we increased the difficulty of the classification problem
+by including data from all 38 HCP tasks.
+Indeed, increased class separability in component space,
+as compared to voxel space, entails differences in f1 score of
+$\approx{17\%}$ (Figure \ref{fig_dimred}).
+
+\begin{figure}
+\begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/accuracies.png}
+\end{centering}
+\vspace{-0.5cm}
+\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
+Depicts f1 scores for prediction of each of 38 classes across training epochs.
+Ordinary logistic regression operating in voxel space (\textit{left plot})
+converged faster but performed worse than
+low-rank logistic regression for few
+(\textit{middle plot}) and many (\textit{right plot}).
+Autoencoder or rest data were not used for these analyses
+($\lambda=1$).
+Hence, projecting the input data into a reduced space for classification
+yields higher class separability.
+}
+\label{fig_dimred}
+\end{figure}
+
+We then quantified the impact of structure from rest data keeping
+all model parameters constant
+($n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$).
+At the beginning of every epoch,
+2000 task and 2000 rest maps were drawn with replacement
+from same amounts of task maps but varying amounts of rest maps.
+In frequently encountered data-scarce settings ($\approx{100}$ samples),
+but not in data-rich settings ($\approx{1000}$ samples),
+the repertoire of rest structure modulates model performance
+(Figure \ref{fig_semisup}).
+
+\begin{figure}
+\begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/semisup_both.png}
+\end{centering}
+\vspace{-0.5cm}
+\caption{\textbf{Effect of richness in rest data}
+Gradient descent was performed on 2000 task and 2000 rest images.
+These were drawn with replacement from identical versus varying
+quantities of task versus rest maps, at the begining of each epoch.
+}
+\label{fig_semisup}
+\end{figure}
+
 \paragraph{Feature identification.}
-\begin{itemize}
-  \item{fit for purpose; the right model should have
-  high probability in the right places}
-  \item{comparing to ordinary logistic regression, low-rank logreg yielded class weights 
-  that were much more similar to feature of the respective training examples,
-  for any choice of $\lambda$}
-  \item{the higher $\lambda$, the better support recovery}
-  \item{conversely, the higher lambda (i.e., bigger emphasis on the
-  resting-state autoencoder), the more useful the low-dimensional transformations
-  for classification in an independent dataset (ARCHI) as measured by f1 score
-  for higher (i.e., 50 and 100) but not lower (i.e., 20) numbers
-  of latent components}
+We finally examined whether the models
+were fit for purpose
+(Figure \ref{fig_weights}).
+%
+To this end, we computed Pearson's correlation between the classifier weights
+and the averaged neural activity map for each of the 18 tasks.
+%
+Ordinary logistic regression thus yielded a mean feature identification
+of $\rho=0.28$.
+%
+For SS-LR LogReg ($\lambda=0.25, 0.5, 0.75, 1$),
+a per-class-weight map was computed by matrix
+multiplication of the two inner layers.
+Feature identification thus ranged
+between $\rho=0.35$ and $\rho=0.55$ for $n=5$,
+between $\rho=0.59$ and $\rho=0.69$ for $n=20$, and
+between $\rho=0.58$ and $\rho=0.69$ for $n=100$.
+%
+Consequently,
+SS-LR LogReg puts higher absolute weights on relevant structure.
+This reflect a higher signal-to-noise ratio, in part explained
+by the more BOLD-typical local contiguity.
+%
+Conversely, SS-LR LogReg puts lower probability mass on irrelevant structure.
+Despite lower interpretability, the
+salt-and-pepper-like weight maps obtained from
+ordinary logistic regression are sufficient for
+good classification performance.
+%
+Hence, SS-LR LogReg yielded class weights 
+that were much more similar to features of the respective training samples
+for any choice of $\lambda$.
+SS-LR LogReg captures genuine properties of neural activity patterns,
+rather than participant- or paradigm-specific artefacts.
 
-support recovery:
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.98\textwidth]{figures/figure_weights_perc75.png}
+\end{centering}
+\vspace{-0.1cm}
+\caption{\textbf{Classification weight maps}
+The voxel predictors corresponding to 5 exemplary
+(of 18 total) psychological tasks (\textit{rows})
+from the HCP dataset \cite{barch2013}.
+\textit{Left column:} ordinary logistic regression (same
+implementation but without bottleneck or autoencoder),
+\textit{middle column:} semi-supervised low-rank logistic regression
+($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
+\textit{right column:} voxel-wise average of whole-brain
+activity maps from each task.
 Low-rank logistic regression
@@
 c) yields BOLD-typical local contiguity.
-
-the features of the classes show that we captured activaiton properties,
-not particpant or paradigm properties
-\end{itemize}
+All values are z-scored and thresholded at the 75th percentile.
+}
+\label{fig_weights}
+\end{figure}
 

### 2015-05-31 nips3mm_recovery.py

@@ -27,6 +27,7 @@ print('Running THEANO on %s' % theano.config.device)
 from nilearn.image import concat_imgs
 import joblib
 import time
+from scipy.stats import zscore
 
 LR_AE_DIR = 'nips3mm'
 LR_DIR = 'nips3mm_vanilla'
@@ -274,7 +275,6 @@ for itask, task in enumerate(HCP_contrasts):
 mean_supp_z = zscore(mean_supp, axis=1)
     
 # get classification weights
-from scipy.stats import zscore
 lr_supp = np.load(op.join(LR_DIR, 'V0comps.npy'))
 lr_supp_z = zscore(lr_supp, axis=1)
 
@@ -283,7 +283,7 @@ WRITE_DIR = 'nips3mm_recovery'
 lambs = [0.25, 0.5, 0.75, 1]
 import re
 from scipy.stats import pearsonr
-for n_comp in [20, 50, 100]:
+for n_comp in [5]:
     corr_means_lr = np.zeros((len(lambs), 18))
     corr_means_lr_ae = np.zeros((len(lambs), 18))
     for ilamb, lamb in enumerate(lambs):
@@ -322,7 +322,7 @@ for n_comp in [20, 50, 100]:
     plt.ylabel('correlation r')
     plt.title('Support Recovery: normal versus low-rank logistic regression\n'
               '%i components' % n_comp)
-    tick_strs = [u'normal'] + [u'low-rank $λ=%.2f$' % val for val in lambs]
+    tick_strs = [u'normal'] + [u'low-rank lambda=%.2f' % val for val in lambs]
     plt.xticks(np.arange(6) + 1, tick_strs, rotation=320)
     plt.ylim(0, 1.0)
     plt.yticks(np.linspace(0, 1., 11), np.linspace(0, 1., 11))

### 2015-05-31 paper_nips/paper.tex
first complete manuscript version for NIPS

@@
 %
-% XXX Michael: Added explicit outline of the model, description can be 
-% XXX improved
 We therefore propose to optimize
 representation and task classification in
-\suggestremove{an identical} \suggestadd{a unified} statistical learning 
-problem\suggestadd{, by introducing a multinomial logistic regression with
-joint rank constraint, which we regularize by forcing representational
-capacity in the form of an autoencoder}.
+a unified statistical learning problem.
+%
+A multinomial logistic regression is introduced that is
+constrained by joint rank with a representation-enforcing
+autoencoder.
 %
@@
 
-% XXX Michael: Suggesting removal of bioinformatics tag,
-% XXX see http://en.wikipedia.org/wiki/Bioinformatics
 \textbf{keywords}: dimensionality reduction; semi-supervised learning;
-\suggestremove{bioinformatics;} fMRI; systems neuroscience
+fMRI; systems neuroscience
 
@@
 during ongoing random thought without controlled environmental stimulation.
-% XXX Michael: Added variation/variance, because otherwise that is trivial:
-In fact, a large proportion of the BOLD signal \suggestadd{variation} is 
-known
+In fact, a large proportion of the BOLD signal variation
+is known
 not to correlate with a particular behavior, stimulus, or experimental task
@@
 by experimental tasks \cite{friston97}.
-% XXX Michael more and more or ever more:
-As a last example, \suggestremove{always} \suggestadd{more and} more 
+As a last example, an increasing number of
 neuroimaging studies model
@@
 This ignores the fact that the BOLD
-signal exhibits spatially distributed \suggestremove{activity} patterns
-\suggestadd{of coherent activity}.
+signal exhibits spatially distributed patterns
+of coherent neural activity.
 %
 Second, existing supervised neuroimaging analyses
-\suggestremove{do not acknowledge} \suggestadd{cannot exploit} the abundance
+cannot exploit the abundance
 of easily acquired resting-state data \cite{biswaldiscovery}.
-These may allow better \suggestremove{sampling} \suggestadd{discovery of} 
+These may allow better discovery of
 the manifold of brain states
@@
 
-% XXX Michael: This parenthesis structure should probably be unravelled into
-% XXX a sentence.
-Both these biological properties can be conjointly exploited in a mixed
- (i.e., using task and rest data)
-low-rank (i.e., performing network decomposition)
-approach.
+Both these biological properties can be conjointly exploited in an
+approach that is
+mixed (i.e., using task and rest data)
+and 
+low-rank (i.e., performing network decomposition).
 %
 The integration of brain-network discovery into
-supervised classification formally yields a semi-supervised learning
-% XXX Michael: No, it doesn't necessarily. You can do that with labeled data
+supervised classification can yield a semi-supervised learning
 framework. The most relevant
@@
 Autoencoders
-are \suggestremove{one-layered} learning models 
+are often one-layered learning models 
 that condense the input data to
-\suggestremove{local and global} \suggestadd{(typically lower-dimensional)} 
+local and global 
 representations
-by \suggestremove{improving} \suggestadd{optimizing}
-reconstruction from them.
+by optimizing reconstruction from them.
 %
@@
 % XXX reviewers annoyed at us because of it.
-They behave like a \suggestadd{(truncated, non-centered)} PCA
+% DanilO: cool thanks. That is the stuff, we need.
+They behave like a (truncated, non-centered) PCA
 in case of one linear hidden layer and a squared error loss
 \cite{baldi1989neural}.
-This architecture yields a \suggestadd{non-}convex optimization objective
+This architecture yields an optimization objective
 with unique global minimum.
@@
 Moreover, they have the characteristics of an ICA in case of tied weights
-and adding a nonlinear convex transformation \cite{le2011ica}.
+and adding a nonlinear convex function
+at the first layer \cite{le2011ica}.
 % XXX What convex transformation? Why is that relevant? Convexity is only
 % XXX relevant for the loss
+% Danilo: it's about adding a neural-network-like nonconvex function for
+% first-layer activation
 These authors further demonstrated that ICA, sparse autoencoders, and 
@@
 onto the main axes of variation and thus % XXX Michael <- that's PCA
-reverse-engineer properties of the underlying
-neural processes \cite{olshausen96}.
+reverse-engineer properties of the data-generating
+processes in the brain \cite{olshausen96}.
 % XXX Michael ^ I don't get that last part nor the citation.
+% Danilo: that is the magic
 
 In the present investigation,
-a linear autoencoder will be \suggestremove{fed by} \suggestadd{fit to} 
+a linear autoencoder will be fit to
 (unlabeled) rest data and
-integrated as \suggestadd{a rank-reducing} bottleneck
-into a \suggestremove{low-rank} \suggestadd{multinomial} 
-logistic regression \suggestremove{fed by} \suggestadd{fit to} 
+integrated as a rank-reducing bottleneck
+into a multinomial
+logistic regression fit to
 (labeled) task data.
@@
 % XXX implications. But in general, only the second layer represents a basis.
-\suggestadd{projectors to} the discovered set of basis functions
+projectors to the discovered set of basis functions
 whose linear combinations are learned
@@
 Neurobiologically, this allows 
-delineating a low-dimensional \suggestremove{manifold}
-\suggestadd{vector space} of brain network patterns and then 
+delineating a low-dimensional manifold
+of brain network patterns and then 
 distinguishing mental tasks
@@
 resting-state autoencoders that
-put probability mass on the most neurobiologically
-% XXX Michael: ^ MEH, we are not really in a probabilistic setting ...
+favor the most neurobiologically
 valid models in the hypothesis set.
 %
-Taken together,
+In summary,
 the important modes of variation in brain dynamics and
@@
 contrasts from the task battery, which certifies excellent
-coverage \cite{barch2013}.
-%
-% No, this is not what is called 'coverage'
+activity patterns covering extended parts of the brain \cite{barch2013}.
 %
@@
 rest map,
-$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$\suggestremove{ hidden dimensions 
-(i.e.,
-distributed neural activity patterns)}
-\suggestadd{-dimensional hidden state}, and 
+$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$-dimensional hidden state
+(i.e., distributed neural activity patterns), and 
 $\mathbf{x' \in \mathbb{R}^{d}}$
@@
 % XXX correct, so I am commenting it
-%% This reconstruction error criterion equates with
-%% maximizing a lower bound on the mutual information between
-%% input and the learned representation.
-\suggestadd{
+% XXX Danilo: I found this in Bengio's papers somewhere
+This reconstruction error criterion equates with
+maximizing a lower bound on the mutual information between
+input and the learned representation.
 Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
@@
 \textit{analysis} and that of signal \textit{synthesis}.
-Whereas the first layer \textit{analyzes}
-the data in order to obtain the cleanest latent representation, the second
+The first layer \textit{analyzes}
+the data in order to obtain the cleanest latent representation,
+while the second
 layer represents building blocks from which to \textit{synthesize} the data
-using the latent activations. Tying these functionalities makes the analysis
-side interpretable and pulls all non-zero singular values towards 1. 
+using the latent activations.
+Tying these processes together makes the analysis
+layer interpretable and pulls all non-zero singular values towards 1. 
 Nonlinearities were not applied to the
 activations in the first layer.
-}
+
 \paragraph{Reduced-rank logistic regression.}
@@
 \(\Omega(\theta)\) represents a regularization, for which we choose the
-ElasticNet type, i.e. a combination of $\ell_1$ and $\ell_2$ penalties
+ElasticNet-type, i.e., a combination of $\ell_1$ and $\ell_2$ penalties
 $\forall p \in \theta$.
@@
 \paragraph{Optimization.}
-The common objective was \suggestremove{approximated} \suggestadd{optimized}
-% XXX why 'approximated' ?
+The common objective was optimized
 in the neuroimaging data
-by \suggestremove{updating parameters by derivates}
-\suggestadd{gradient descent in the parameters}
-% XXX derivates ???
+by gradient descent in the parameters
 of the 
-semi-supervised low-rank logistic regression.
+semi-supervised low-rank logistic regression (SS-LR LogReg).
 The required gradients are easily obtained by using the chain rule to
 backpropagate error derivatives through the linear network.
-As solver, we chose \textit{rmsprop} \cite{rmsprop},
-a mini-batch version of rprop.
+As solver, we chose \textit{rmsprop} \cite{rmsprop}.
+%, a mini-batch version of rprop.
 This procedure dictates an adaptive learning rate
@@
 % XXX Maybe we could also talk less about RMSprop.
+%
+% XXX Danilo: I think we are not alternating, Xrest and Xtask go in
+% simultaneously at every bach in every epoch
 
@@
 % XXX Michael: I use f earlier up, is this f here necessary?
-prior knowledge on known properties of the unknown target function $f$.
+prior knowledge on
+\textit{known} properties of the \textit{unknown} target function $f$.
 Rather than only relying on input-output pairs in the learning process,
@@
 %
-This 2-step approaches were compared against semi-supervised low-rank
-logistic regression
-(SS-LR LogReg, two hidden layers, $\ell_1=0.1$, $\ell_2=0.1$,
+This serail two-step approach was
+compared against SS-LR LogReg
+(two hidden layers, $\ell_1=0.1$, $\ell_2=0.1$,
 500 maximum iterations, $\lambda=0.75$).
@@
 was assessed by using it as dimensionality reduction of an
-independent labaled dataset (i.e., ARCHI) and conducting ordinary
+independent labeled dataset (i.e., ARCHI) and conducting ordinary
 logistic regression on the ensuing component loadings.
@@
 
-We made several noteworthy observations.
+We made several noteworthy observations (Table \ref{table_two}).
 %
@@
 This suggests that the presence of more rest-data-inspired hidden components
-yields to more effective feature representation in unrelated task data.
+results in more effective feature representations in unrelated task data.
 %
@@
 
-Depicts f1 scores for prediction of each of 38 classes across training epochs.
-Ordinary logistic regression operating in voxel space (\textit{left plot})
-converged faster but performed worse than
-low-rank logistic regression for few
-(\textit{middle plot}) and many (\textit{right plot}).
-Autoencoder or rest data were not used for these analyses
-($\lambda=1$).
-Hence, projecting the input data into a reduced space for classification
-yields higher class separability.
+% Depicts f1 scores for prediction of each of 38 classes across training epochs.
+% Ordinary logistic regression operating in voxel space (\textit{left plot})
+% converged faster but performed worse than
+% low-rank logistic regression for few
+% (\textit{middle plot}) and many (\textit{right plot}).
+% Autoencoder or rest data were not used for these analyses
+% ($\lambda=1$).
+% Hence, projecting the input data into a reduced space for classification
+% yields higher class separability.
 
@@
 We first quantified the gain of introducing a bottleneck layer
-disregaring the autoencoder.
+disregarding the autoencoder.
 To this end, ordinary logstic regression was juxtaposed
@@
 Indeed, increased class separability in component space,
-as compared to voxel space, entails differences in f1 score of
+as compared to voxel space, entails differences in
+generalization perforamnce of
 $\approx{17\%}$ (Figure \ref{fig_dimred}).
@@
 low-rank logistic regression for few
-(\textit{middle plot}) and many (\textit{right plot}).
+(\textit{middle plot}) and many (\textit{right plot}) latent modes.
 Autoencoder or rest data were not used for these analyses
@@
 but not in data-rich settings ($\approx{1000}$ samples),
-the repertoire of rest structure modulates model performance
+the repertoire of rest structure modulated model performance
 (Figure \ref{fig_semisup}).
@@
 for any choice of $\lambda$.
-SS-LR LogReg captures genuine properties of neural activity patterns,
-rather than participant- or paradigm-specific artefacts.
+SS-LR LogReg therefore
+captures genuine properties of neural activity patterns,
+rather than participant- or study-specific artefacts.
 
@@
 k) only $\ell_2$-penalty was disadvantageous in all tested cases,
-which favored elasticnet regularization chosen in the above analyses.
+which favored ElasticNet regularization chosen in the above analyses.
 
@@
 supervised learning question.
-The higher generalization accuracy and support recovery, comparing to
+The higher generalization accuracy and feature recovery, comparing to
 ordinary logistic regression, hold potential

### 2015-05-31 paper_nips/paper.tex
Comments on abstract and introduction

@@
 %
-However, testing hypotheses on the neural correlates of mental tasks
-depends on the representation of the observations.
+However, testing hypotheses on the neural correlates of an increasing number of mental tasks makes it crucial to represent the observations adequately.
 %
@@
 We show that this approach yields more accurate and more interpretable
-neural models of psychological tasks in a reference dataset.
+neural models of psychological tasks in a reference dataset, and better generalization to other datasets.
 %
@@
 separate BOLD signals into parsimonious network components.
-Thus extracted brain networks are probably
+The extracted brain networks are probably
 manifestations of electrophysiological oscillation frequencies \cite{hipp15}.
@@
 % XXX Michael ^ I don't get that last part nor the citation.
+\textbf{XXX Bertrand: neither do I. I think you exxagerate.}
 % Danilo: that is the magic
@@
 unconnected domains of systems neuroscience.
+
+\textbf{XXX Bertrand: I think that you could emphasize the problem that we are always data-poor, so that the set of representations that  can be extracted from SPMs taken from  a few tens of subjects is limited.
+What you propose it to extend it by i) co-analysing many problems simultaneously (multi-task) and using unlabled data that spana space of meaningful configurations
+}
+\textbf{XXX Bertrand: We need a sentence like: Our contribution is 1. XXX 2.YYY etc.}
 

### 2015-06-01 paper_nips/paper.tex
A pass on the end of the paper

@@
 The labeled and unlabeled data were fed into a linear statistical model
-composed of autoencoder and low-rank logistic regression.
+composed of an autoencoder and low-rank logistic regression.
+
+XXX (Bertrand): You can simplify data description if needed for the sake of place. It's OK as long as you can refer to a more complete publication describing the data.
 
@@
 latent representation $\mathbf{z}$
-by
+and reconstructs it back to $\mathbf{x'}$ by
 
@@
   \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
-  \mathbb E\left[\| \mathbf{x} - f_{AE}(\mathbf{x}) \|^2\right],
+  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right],
 \end{equation}
-with \(f_{AE}(\mathbf x) = %
-\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}\)
-the autoencoder.
+%with \(f_{AE}(\mathbf x) = %
+%\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}\)
+% the autoencoder.
 % XXX Michael: I don't know to what extent the following makes sense or is
@@
 scaled gradients from a running average.
-Gradient normalization by rmsprop
+Gradient normalization by \textit{rmsprop}
 is known to effectively exploit curvature information.
@@
 The matrix parameters were initalized by Gaussian random values multiplied
-by a gain of $0.004$. The bias parameters were initalized to zero.
+by a gain of $0.004$. 
+%
+
+XXX: Sorry, I don't like magic numbers. Please explain where this comes from
+ ? Maybe $\sqrt{1 / p}$ ?
+
+%
+The bias parameters were initalized to zero.
 With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
@@
 % XXX Michael: Probably remove the explanations following and just state values
+% XXX: Bertrand: what is f ??
 where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
@@
 (stochastic gradient descent, adadelta, and adagrad) but found that
-rmsprop converged faster and with higher generalization performance.
+\textit{rmsprop} converged faster and with higher generalization performance.
 
@@
 supervised learning based on ICA and SPCA loadings
-ranged from 37,53\% to 87,28\%
+ranged from 38\% to 87\%
 and
-32,19\% to 83,97\%, respectively (Table \ref{table_one}).
+32\% to 84\%, respectively (Table \ref{table_one}).
 %
@@
  \vspace{-0.2cm}
- \caption{Serial versus parallel dimensionality reduction and classification}
+ \caption{Serial versus parallel dimensionality reduction and classification
+XXX: only one decimal is meaningful}
  \label{table_one}
@@
 \vspace{-0.2cm}
- \caption{Performance of SS-LR LogReg across model parameter choices}
+ \caption{Performance of SS-LR LogReg across model parameter choices. XXX: recall what the chance-level performance is.XXX: please add lambda=0}
   \label{table_two}
@@
 noninferior out-of-dataset performance
-of X and Y, respectively (not shown in Table \ref{table_two}).
+of X and Y, respectively (not shown in Table \ref{table_two} XXX: please show it).
 This confirms that guiding model learning by task-unrelated structure
@@
 To this end, ordinary logstic regression was juxtaposed
-with LR LogRed at $\lambda=1$ (no autoencoder).
+with LR LogReg  at $\lambda=1$ (no autoencoder).
 For this experiment, we increased the difficulty of the classification problem
@@
 yields higher class separability.
+XXX: ugly figure: use larger font for labels and titles.
+XXX: includes pdf figures, not a png file.
 }
@@
 %
-Ordinary logistic regression thus yielded a mean feature identification
+Ordinary logistic regression thus yielded a mean correlation of
 of $\rho=0.28$.
@@
 rather than participant- or study-specific artefacts.
+XXX: But what you show here is mostly a denoising or regularization effect.
 
@@
 implementation but without bottleneck or autoencoder),
+XXX: sorry, but is this a multinomial or plain logistic ? Please clarify 
 \textit{middle column:} semi-supervised low-rank logistic regression
@@
 %
+\textbf{XXX: Please give explanations why this is the case. My feeling is that we're still not rich enough in the sample dimension to make these tricks worthwhile.}
 Further, g) ``pretraining'' of the bottleneck $\mathbf{W_0}$
@@
 fMRI data probably concentrate near
-a low-dimensional manifold of brain networks.
+a low-dimensional manifold that correspond to peculiar brain networks combinations.
 Extracting fundamental building blocks of brain organization might
@@
 by the German National Academic Foundation (D.B.).
+XXX: Note that this will be removed at submission for the sake of anonymization.
 }

### 2015-06-02 nips3mm_scarcity.py
before Oliviers comments

@@ -41,8 +41,8 @@ if not op.exists(WRITE_DIR):
     os.mkdir(WRITE_DIR)
 
 MAX_SAMPLES = 2000
-MAX_REST_SAMPLES = 1
-MAX_TASK_SAMPLES = 1000
+MAX_REST_SAMPLES = 750
+MAX_TASK_SAMPLES = 100
 
 ##############################################################################
 # load+preprocess data
@@ -466,7 +466,7 @@ n_comps = [20]
 for n_comp in n_comps:
     # for lambda_param in [0]:
     #for lambda_param in [0.5]:
-    for lambda_param in [1.0]:
+    for lambda_param in [0.5]:
         l1 = 0.1
         l2 = 0.1
         my_title = r'Low-rank LR + AE (combined loss, shared decomp): n_comp=%i L1=%.1f L2=%.1f lambda=%.2f res=3mm spca20RS max%i r%i t%i' % (
@@ -514,7 +514,6 @@ STOP_CALCULATION
 # equally scaled plots
 import re
 pkgs = glob.glob(RES_NAME + '/*dbg_epochs*.npy')
-dbg_epochs_ = np.load(pkgs[0])
 dbg_epochs_ = np.load(pkgs[0])
 
 d = {
@@ -587,10 +586,10 @@ for n_comp in n_comps:  #
         # n_hidden = int(re.search('comp=(?P<comp>.{1,2,3})_', p).group('comp'))
         n_hidden = int(re.search('comp=(.{1,3})_', p).group(1))
         max_samples = int(re.search('max(.{1,4})_', p).group(1))
-        rest_samples = int(re.search('_r(.{2,4})_', p).group(1))
+        rest_samples = int(re.search('_r(.{1,4})_', p).group(1))
         task_samples = int(re.search('_t(.{2,4})dbg', p).group(1))
         
-        if (task_samples != 1000):
+        if (task_samples != 100):
             continue
         
         if n_comp != n_hidden:

### 2015-06-02 nips3mm_serial.py

@@ -73,6 +73,157 @@ del X_task
 # class defintion
 ##############################################################################
 
+class AutoEncoder(BaseEstimator):
+    def __init__(self, n_hidden, gain1, learning_rate, max_epochs=500):
+        self.n_hidden = n_hidden
+        self.gain1 = gain1
+        self.max_epochs = max_epochs
+        self.learning_rate = np.float32(learning_rate)
+
+    # def rectify(X):
+    #     return T.maximum(0., X)
+
+    from theano.tensor.shared_randomstreams import RandomStreams
+
+    def RMSprop(self, cost, params, lr=0.001, rho=0.9, epsilon=1e-6):
+        grads = T.grad(cost=cost, wrt=params)
+        updates = []
+        for p, g in zip(params, grads):
+            acc = theano.shared(p.get_value() * 0.)
+            acc_new = rho * acc + (1 - rho) * g ** 2
+            gradient_scaling = T.sqrt(acc_new + epsilon)
+            g = g / gradient_scaling
+            updates.append((acc, acc_new))
+            updates.append((p, p - lr * g))
+        return updates
+        
+    def get_param_pool(self):
+        cur_params = (
+            self.W0s, self.W1s, self.bW0s, self.bW1s
+        )
+        return cur_params
+        
+
+    def fit(self, X_rest):
+        DEBUG_FLAG = True
+
+        # self.max_epochs = 333
+        self.batch_size = 100
+        n_input = X_rest.shape[1]  # sklearn-like structure
+        n_output = n_input
+        rng = np.random.RandomState(42)
+        self.params_from_last_iters = []
+        
+        self.input_restdata = T.matrix(dtype='float32', name='input_restdata')
+
+        index = T.iscalar(name='index')
+        
+        # prepare data for theano computation
+        if DEBUG_FLAG:
+            self.dbg_ae_cost_ = list()
+            self.dbg_ae_nonimprovesteps = list()
+        X_rest_s = theano.shared(value=np.float32(X_rest), name='X_rest_s')
+        ae_train_samples = len(X_rest)
+
+        # W -> unsupervised / auto-encoder
+
+        # computational graph: auto-encoder
+        W0_vals = rng.randn(n_input, self.n_hidden).astype(np.float32) * self.gain1
+
+        self.W0s = theano.shared(W0_vals)
[…]

### 2015-06-02 paper_nips/paper.tex
sneaked in BTs last comments

@@
   Bertrand Thirion,
-  Ga\"el Varoquaux \\\\\textbf{\textit{email:} }firstname.lastname@inria.fr}
+  Ga\"el Varoquaux \\\textbf{\textit{email:} }firstname.lastname@inria.fr}
 
@@
 %
-However, testing hypotheses on the neural correlates of an increasing number of mental tasks makes it crucial to represent the observations adequately.
+However, testing hypotheses on the neural correlates of an
+increasing number of mental tasks makes it crucial to
+represent the observations adequately.
 %
@@
 %
-We show that this approach yields more accurate and more interpretable
-neural models of psychological tasks in a reference dataset, and better generalization to other datasets.
+We show that this approach yields more accurate and interpretable
+neural models of psychological tasks in a reference dataset,
+as well as better generalization to other datasets.
 %
 
-\textbf{keywords}: dimensionality reduction; semi-supervised learning;
-fMRI; systems neuroscience
+% Danilo: NIPS does actually not seem to have keywords
+%\textbf{\\keywords}: curse of dimensionality; semi-supervised learning;
+%fMRI; systems neuroscience
 
@@
 Hence, autoencoders can flexibly project the neuroimaging data
-onto the main axes of variation and thus % XXX Michael <- that's PCA
-reverse-engineer properties of the data-generating
-processes in the brain \cite{olshausen96}.
-% XXX Michael ^ I don't get that last part nor the citation.
-\textbf{XXX Bertrand: neither do I. I think you exxagerate.}
-% Danilo: that is the magic
+onto the main directions of variation.
 
@@
 We can then solve the unsupervised data representation and the
-supervised classification in an identical statistical learning
-problem.
+supervised classification, previously studied in isolation,
+in an identical statistical learning problem.
 %
@@
 %
-In summary,
-the important modes of variation in brain dynamics and
-the neural correlates subserving mental operations
-have mostly been studied in isolation.
-We provide a principled computational framework to link these previously
-unconnected domains of systems neuroscience.
-
-\textbf{XXX Bertrand: I think that you could emphasize the problem that we are always data-poor, so that the set of representations that  can be extracted from SPMs taken from  a few tens of subjects is limited.
-What you propose it to extend it by i) co-analysing many problems simultaneously (multi-task) and using unlabled data that spana space of meaningful configurations
-}
-\textbf{XXX Bertrand: We need a sentence like: Our contribution is 1. XXX 2.YYY etc.}
+Practically, neuroimaging research is frequently suffering from
+data scarcity. This limits the set of representations that can be
+extracted from GLM analyses based on few participants.
+%
+We therefore contribute a computational framework that
+1.) co-analyzes many problems simultaneously
+(thus find shared representations by multi-task learning)
+and
+2.) exploits unlabeled data
+(that span a space of meaningful configurations).
 
@@
 composed of an autoencoder and low-rank logistic regression.
-
-XXX (Bertrand): You can simplify data description if needed for the sake of place. It's OK as long as you can refer to a more complete publication describing the data.
 
@@
 %
-
-XXX: Sorry, I don't like magic numbers. Please explain where this comes from
- ? Maybe $\sqrt{1 / p}$ ?
-
-%
 The bias parameters were initalized to zero.
@@
   \begin{split}
-    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial f}{\partial \theta}\right)^2
+    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial \mathcal{L}}{\partial \theta}\right)^2
 %v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
@@
 % XXX Michael: Probably remove the explanations following and just state values
-% XXX: Bertrand: what is f ??
+\textbf{XXX: Bertrand: what is f ?? Danilo: replaced f by loss L}
+
 where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
@@
 We used \textit{nilearn} to handle
-the high-dimensional neuroimaging data 
+the large quantities of neuroimaging data 
 \cite{abrah14}
@@
  \textit{n}               & ICA + LogReg & SPCA + LogReg & SS-LR LogReg \\ \hline
- \multicolumn{1}{l|}{5}   & 37,53 \%     & 32,19 \%      & 96,50 \%     \\
- \multicolumn{1}{l|}{20}  & 80,98 \%     & 78,15 \%      & 97,33 \%     \\
- \multicolumn{1}{l|}{50}  & 84,19 \%     & 83,97 \%      & 97,69 \%     \\
- \multicolumn{1}{l|}{100} & 87,28 \%     & 82,19 \%      & 97,80 \%     \\ \hline
+ \multicolumn{1}{l|}{5}   & 37,5 \%     & 32,2 \%      & 96,5 \%     \\
+ \multicolumn{1}{l|}{20}  & 81,0 \%     & 78,2 \%      & 97,3 \%     \\
+ \multicolumn{1}{l|}{50}  & 84,2 \%     & 84,0 \%      & 97,7 \%     \\
+ \multicolumn{1}{l|}{100} & 87,3 \%     & 82,2 \%      & 97,8 \%     \\ \hline
  \end{tabular}
@@
  \vspace{-0.2cm}
- \caption{Serial versus parallel dimensionality reduction and classification
-XXX: only one decimal is meaningful}
+ \caption{\textbf{Serial versus parallel dimensionality reduction and classification.}
+ Chance is at 5,6\%.
+ }
  \label{table_one}
@@
 the balance between autoencoder and logistic regression
-($\lambda=0.25, 0.5, 0.75, 1$).
+($\lambda=0, 0.25, 0.5, 0.75, 1$).
 Assuming 5 latent directions of variation should yield models with
@@
   \resizebox{1.00\textwidth}{!}{%
-\begin{tabular}{l|cccc|cccc|cccc|cccc|}
+\tabcolsep=0.05cm
+\begin{tabular}{l|ccccc|ccccc|ccccc|}
   \hline
-                            & \multicolumn{4}{c|}{\textit{n = 5}}                                                                                                            & \multicolumn{4}{c|}{\textit{n = 20}}                                                                                                                   & \multicolumn{4}{c}{\textit{n = 100}}                                                       \\ \hline
-  \textit{}                 & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
-  Out-of-sample\\accuracy   & \textit{88,90\%}                   & 95,13\%                            & \textbf{96,49\%}                  & 95,72\%                          & 97,44\%                              & \textbf{97,80\%}                    & 97,33\%                              & \textit{97,32\%}                   & 97,21\%          & \textit{97,03\%} & \textbf{97,80\%} & 97,38\%                            \\
-  Precision (mean)          & \textit{86,98\%}                   & 94,86\%                            & \textbf{96,28\%}                  & 95,36\%                          & \textbf{97,38\%}                     & 97,06\%                             & \textit{96,98\%}                     & 97,02\%                            & 96,90\%          & \textit{96,47\%} & \textbf{97,50\%} & 96,91\%                            \\
-  Recall (mean)             & \textit{88,27\%}                   & 95,19\%                            & \textbf{96,57\%}                  & 95,70\%                          & \textbf{97,51\%}                     & 97,50\%                             & \textit{97,35\%}                     & 97,36\%                            & 97,20\%          & \textit{97,15\%} & \textbf{97,88\%} & 97,37\%                            \\
-  F1 score (mean)           & \textit{86,64\%}                   & 94,89\%                            & \textbf{96,38\%}                  & 95,43\%                          & \textbf{97,42\%}                     & 97,22\%                             & \textit{97,11\%}                     & 97,13\%                            & 97,00\%          & \textit{96,71\%} & \textbf{97,65\%} & 97,07\%                            \\ \hline
-  Out-of-dataset \\accuracy & 60,83\%                            & \textit{54,30\%}                   & 60,69\%                           & \textbf{62,92\%}                 & 79,72\%                              & \textbf{81,94\%} & 79,72\%          & \textit{79,44\%}                     & \textbf{82,08\%} & 81,66\%         & 81,25\%          & \textit{75,83\%}                   \\ \hline
+                            & \multicolumn{5}{c|}{\textit{n = 5}}                                                                                                                                                 & \multicolumn{5}{c|}{\textit{n = 20}}                                                                                                                                                      & \multicolumn{5}{c}{\textit{n = 100}}                                                                           \\ \hline
+  \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
+  Out-of-sample\\accuracy   & \textit{6,0\%}                     & 88,9\%                             & 95,1\%                            & \textbf{96,5\%}                    & 95,7\%                           & \textit{5,5\%}                   & 97,4\%                               & \textbf{97,8\%}                     & 97,3\%                               & 97,3\%                             & \textit{6,1\%}   & 97,2\%           & 97,0\%           & \textbf{97,8\%}  & 97,4\%                             \\
+  Precision (mean)          & \textit{5,9\%}                     & 87,0\%                             & 94,9\%                            & \textbf{96,3\%}                    & 95,4\%                           & \textit{5,1\%}                   & \textbf{97,4\%}                      & 97,1\%                              & 97,0\%                               & 97,0\%                             & \textit{5,9\%}   & 96,9\%           & 96,5\%           & \textbf{97,5\%}  & 96,9\%                             \\
+  Recall (mean)             & \textit{5,6\%}                     & 88,3\%                             & 95,2\%                            & \textbf{96,6\%}                    & 95,7\%                           & \textit{4,6\%}                   & \textbf{97,5\%}                      & 97,5\%                              & 97,4\%                               & 97,4\%                             & \textit{7,2\%}   & 97,2\%           & 97,2\%           & \textbf{97,9\%}  & 97,4\%                             \\
+  F1 score (mean)           & \textit{4,1\%}                     & 86,6\%                             & 94,9\%                            & \textbf{96,4\%}                    & 95,4\%                           & \textit{3,8\%}                   & \textbf{97,4\%}                      & 97,2\%                              & 97,1\%                               & 97,1\%                             & \textit{5,3\%}   & 97,0\%           & 96,7\%           & \textbf{97,7\%}  & 97,2\%                             \\ \hline
+  Out-of-study\\accuracy    & \textit{39,4\%}                    & 60,8\%                             & 54,3\%                            & 60,7\%                             & \textbf{62,9\%}                  & \textit{77,0\%}                  & 79,7\%                               & \textbf{81,9\%}                     & 79,7\%                               & 79,4\%                             & 79,2\%           & \textbf{82,2\%}  & 81,7\%           & 81,3\%           & \textit{75,8\%}                    \\ \hline
   \end{tabular}
@@
 \vspace{-0.2cm}
- \caption{Performance of SS-LR LogReg across model parameter choices. XXX: recall what the chance-level performance is.XXX: please add lambda=0}
+ \caption{\textbf{Performance of SS-LR LogReg across model parameter choices.}
+ Chance is at 5,6\%.
+ }
   \label{table_two}
@@
 Second, the higher the number of latent components $n$,
-the higher the out-of-dataset performance with small choices for $\lambda$.
+the higher the out-of-study performance with small choices for $\lambda$.
 This suggests that the presence of more rest-data-inspired hidden components
@@
 decomposition matrix ($\lambda=0$) resulted in
-noninferior out-of-dataset performance
-of X and Y, respectively (not shown in Table \ref{table_two} XXX: please show it).
+noninferior out-of-study performance
+of 77,0\% and 79,2\%,
+respectively (Table \ref{table_two}).
 This confirms that guiding model learning by task-unrelated structure
-extract features of general relevance
+extracts features of general relevance
 beyond the supervised problem at hand.
@@
 yields higher class separability.
+Chance is at 2,6\%.
 XXX: ugly figure: use larger font for labels and titles.
@@
 quantities of task versus rest maps, at the begining of each epoch.
+Chance is at 5,6\%.
 }
@@
 {\small
-Data were provided by the Human Connectome Project. The study was supported
-by the German National Academic Foundation (D.B.).
-XXX: Note that this will be removed at submission for the sake of anonymization.
+Data were provided by the Human Connectome Project.
+% The study was supported by the German National Academic Foundation (D.B.).
 }

### 2015-06-03 paper_nips/paper.tex
intermediate

@@
 
-\title{Semi-supervised low-rank logistic regression for
+\title{Semi-supervised factored logistic regression for
 high-dimensional neuroimaging data}
@@
 %
-However, testing hypotheses on the neural correlates of an
-increasing number of mental tasks makes it crucial to
-represent the observations adequately.
+However, testing hypotheses on the neural correlates of
+larger mental task sets
+necessitates adequate representation of the observations.
 %
 We therefore propose to optimize
-representation and task classification in
+representation modelling and task classification in
 a unified statistical learning problem.
@@
 with mental tasks.
-To discover, on the one hand, spatial distributions of neural activity
+To \textit{discover}, on the one hand,
+spatial distributions of neural activity
 structure across time,
@@
 
-To test, on the other hand,
+To \textit{test}, on the other hand,
 the neural correlates underlying mental tasks,
@@
 
-Both these biological properties can be conjointly exploited in an
+Both these neurobiological properties can be conjointly exploited in an
 approach that is
-mixed (i.e., using task and rest data)
+mixed (i.e., using rest and task data)
 and 
-low-rank (i.e., performing network decomposition).
+factored (i.e., performing network decomposition).
 %
@@
   \vspace{-0.7cm}
-  \caption {\textbf{Architecture.}
+  \caption {\textbf{Model architecture.}
   Linear autoencoders find an optimal compression of 79,941 brain voxels
@@
 %
-Autoencoders
+Autoencoders (AE)
 are often one-layered learning models 
@@
 \cite{baldi1989neural}.
-This architecture yields an optimization objective
-with unique global minimum.
+% This architecture yields an optimization objective
+% with unique global minimum.
 Autoencoders behave like a SPCA if shrinkage terms are added to the
@@
 %
-Hence, autoencoders can flexibly project the neuroimaging data
+Thus, autoencoders can flexibly project the neuroimaging data
 onto the main directions of variation.
@@
 We therefore contribute a computational framework that
-1.) co-analyzes many problems simultaneously
-(thus find shared representations by multi-task learning)
+1) co-analyzes many problems simultaneously
+(thus find shared representations by ``multi-task learning'')
 and
-2.) exploits unlabeled data
+2) exploits unlabeled data
 (that span a space of meaningful configurations).
@@
 These labeled data were complemented by unlabeled activity maps
-from HCP acquisitions of unconstrained resting-state activity.
+from HCP acquisitions of unconstrained resting-state activity
+\cite{smith2013resting}.
 These reflect brain activity in the absence of controlled thought.
@@
 
-We were further interested in the utility of the optimal low-rank projection
+We were further interested in the utility of the
+optimized low-rank projection
 in one task dataset for dimensionality reduction in another task dataset.
@@
 The labeled and unlabeled data were fed into a linear statistical model
-composed of an autoencoder and low-rank logistic regression.
+composed of an autoencoder and factored logistic regression.
 
@@
 
-\paragraph{Reduced-rank logistic regression.}
+\paragraph{Factored logistic regression.}
 Lossy compression by a low-dimensional bottleneck
-is also imposed by the first layer of the low-rank
+is also imposed by the first layer of the factored
 multinomial logistic regression.
@@
 Importantly, the optimization problem of the linear autoencoder
-and the low-rank logistic regression
+and the factored logistic regression
 are linked on two levels. First, their transformation matrices mapping from
@@
 rest and task activity data.
-Second, the objectives of the autoencoder and the low-rank
+Second, the objectives of the autoencoder and the factored
 logistic regression are interpolated in the common loss function
@@
 of the 
-semi-supervised low-rank logistic regression (SS-LR LogReg).
+semi-supervised factored logistic regression (SS-F LogReg).
 The required gradients are easily obtained by using the chain rule to
@@
   \begin{split}
-    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\frac{\partial \mathcal{L}}{\partial \theta}\right)^2
+    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})\right)^2
 %v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
 \\
-\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla f(\theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
+\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
   \end{split}
@@
 % XXX Michael: Probably remove the explanations following and just state values
-\textbf{XXX: Bertrand: what is f ?? Danilo: replaced f by loss L}
+% XXX: Bertrand: what is f ?? Danilo: replaced f by more detailed notation
 
@@
 performing data reduction on the (unlabeled)
-first half of the HCP data by ICA and SPCA ($n=5, 20, 50, 100$ components)
+first half of the HCP task data by PCA, SPCA, ICA, and
+AE ($n=5, 20, 50, 100$ components)
 and learning classification models in the (labeled) second half
 by ordinary logistic regression.
+%
+PCA reduced the dimensionality of the task data by
+finding orthogonal network components using a change of basis
+(whitening).
+%
+SPCA separated the BOLD signals into
+network components with few regions by
+a regression-type optimization problem constrained by
+$\ell_1$-penalty
+(no orthogonality assumptions, 1000 maximum iterations,
+per-iteration tolerance of 1 * 10\textsuperscript{-8},
+$alpha=1$).
 %
@@
 initialized by random mixing matrix, whitening).
-SPCA separated the BOLD signals into
-network components with few regions by
-a regression-type optimization problem constrained by
-$\ell_1$-penalty
-(no orthogonality assumptions, 1000 maximum iterations,
-per-iteration tolerance of 1 * 10\textsuperscript{-8}, sparsity
-alpha=1).
 %
-The second half of the data was projected onto the
-latent components discovered in first data half.
-The ensuing component loadings were submitted to ordinary
+AE found a code of latent representations by optimizing projection
+into a bottleneck
+(500 iterations, same implementation as below for rest data).
+%
+The second half of the task data was projected onto the
+latent components discovered in its first half.
+Only the ensuing component loadings were submitted to ordinary
 logistic regression
@@
 %
-This serail two-step approach was
-compared against SS-LR LogReg
-(two hidden layers, $\ell_1=0.1$, $\ell_2=0.1$,
-500 maximum iterations, $\lambda=0.75$).
+These serial two-step approaches were
+compared against parallel decomposition and classification by FLogReg
+(two hidden layers, $\lambda=1.00$, $\ell_1=0.1$, $\ell_2=0.1$,
+500 iterations).
 Importantly, all trained classification models were tested
-on a large, unseen test set (20\% of data) in the presented analyses.
+on a large, unseen test set (20\% of data) in the present analyses.
 %
-Across choices for $n$, SS-LR LogReg
-achieved more than 96\% out-of-sample accuracy, whereas
-supervised learning based on ICA and SPCA loadings
-ranged from 38\% to 87\%
-and
-32\% to 84\%, respectively (Table \ref{table_one}).
+Across choices for $n$, FLogReg
+achieved more than 95\% out-of-sample accuracy, whereas
+supervised learning based on PCA, SPCA, ICA, and AE loadings
+ranged from 32\% to 87\%
+(Table \ref{table_one}).
 %
@@
 classification-relevant structure in the fMRI data,
-rather than solving supervised and unsupervised problems independently.
+rather than solving the supervised and unsupervised problems independently.
 This effect was particularly pronounced when assuming few hidden dimensions.
@@
  \centering
- \resizebox{0.5\textwidth}{!}{%
- \begin{tabular}{lccc}
+ \resizebox{0.8\textwidth}{!}{%
+ \begin{tabular}{lccccc}
  \hline
- \textit{n}               & ICA + LogReg & SPCA + LogReg & SS-LR LogReg \\ \hline
- \multicolumn{1}{l|}{5}   & 37,5 \%     & 32,2 \%      & 96,5 \%     \\
- \multicolumn{1}{l|}{20}  & 81,0 \%     & 78,2 \%      & 97,3 \%     \\
- \multicolumn{1}{l|}{50}  & 84,2 \%     & 84,0 \%      & 97,7 \%     \\
- \multicolumn{1}{l|}{100} & 87,3 \%     & 82,2 \%      & 97,8 \%     \\ \hline
+ \textit{n}               & PCA + LogReg   & SPCA + LogReg  & ICA + LogReg  & AE + LogReg    & FLogReg      \\ \hline
+ \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & 95,7 \%      \\
+ \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & 97,3 \%      \\
+ \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & 97,6 \%      \\
+ \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & 97,4 \%      \\ \hline
  \end{tabular}
@@
 \paragraph{Model performance.}
-SS-LR LogReg was subsequently trained ($500$ epochs) across parameter choices
+SS-F LogReg was subsequently trained ($500$ epochs) across parameter choices
 for the hidden components ($n=5, 20, 100$) and
@@
 Assuming 5 latent directions of variation should yield models with
-higher bias and smaller variance then SS-LR LogReg with 100 latent directions.
+higher bias and smaller variance then SS-F LogReg with 100 latent directions.
 %
@@
 \tabcolsep=0.05cm
-\begin{tabular}{l|ccccc|ccccc|ccccc|}
+\begin{tabular}{l|ccccc|ccccc|ccccc}
   \hline
@@
 \vspace{-0.2cm}
- \caption{\textbf{Performance of SS-LR LogReg across model parameter choices.}
+ \caption{\textbf{Performance of SS-F LogReg across model parameter choices.}
  Chance is at 5,6\%.
@@
 instance the best accuracy, precision, recall or f1 scores on HCP data.
-Classification by low-rank logistic regression is therefore facilitated by
+Classification by factored logistic regression is therefore facilitated by
 imposing structure from the unlabeled rest data.
@@
 % converged faster but performed worse than
-% low-rank logistic regression for few
+% factored logistic regression for few
 % (\textit{middle plot}) and many (\textit{right plot}).
@@
 converged faster but performed worse than
-low-rank logistic regression for few
+factored logistic regression for few
 (\textit{middle plot}) and many (\textit{right plot}) latent modes.
@@
 %
-For SS-LR LogReg ($\lambda=0.25, 0.5, 0.75, 1$),
+For SS-F LogReg ($\lambda=0.25, 0.5, 0.75, 1$),
 a per-class-weight map was computed by matrix
@@
 Consequently,
-SS-LR LogReg puts higher absolute weights on relevant structure.
+SS-F LogReg puts higher absolute weights on relevant structure.
 This reflect a higher signal-to-noise ratio, in part explained
@@
 %
-Conversely, SS-LR LogReg puts lower probability mass on irrelevant structure.
+Conversely, SS-F LogReg puts lower probability mass on irrelevant structure.
 Despite lower interpretability, the
@@
 %
-Hence, SS-LR LogReg yielded class weights 
+Hence, SS-F LogReg yielded class weights 
 that were much more similar to features of the respective training samples
 for any choice of $\lambda$.
-SS-LR LogReg therefore
+SS-F LogReg therefore
 captures genuine properties of neural activity patterns,
@@
 from the HCP dataset \cite{barch2013}.
-\textit{Left column:} ordinary logistic regression (same
+\textit{Left column:} multinomial logistic regression (same
 implementation but without bottleneck or autoencoder),
-XXX: sorry, but is this a multinomial or plain logistic ? Please clarify 
-\textit{middle column:} semi-supervised low-rank logistic regression
+\textit{middle column:} semi-supervised factored logistic regression
 ($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
@@
 activity maps from each task.
-Low-rank logistic regression
-a) puts higher absolute weights on relevant structure
+Semi-supervised factored logistic regression
+a) puts higher absolute weights on relevant structure,
 b) lower ones on irrelevant structure,
@@
 c) yields BOLD-typical local contiguity.
-All values are z-scored and thresholded at the 75th percentile.
+All values are z-scored and thresholded at the $75^{th}$ percentile.
 }

### 2015-06-03 nips3mm_scarcity.py

@@ -1243,3 +1243,52 @@ plt.grid(False)
 plt.show()
 plt.savefig(op.join(WRITE_DIR,
             k.replace(' ', '_') + '_%icomps_scarcities2.png' % n_comp))
+            
+# final scarcity plot 3
+n_comps = [20]
+nrests = []
+naccs = []
+pkgs = glob.glob(RES_NAME + '/*dbg_acc_val_*.npy')
+for n_comp in n_comps:  # 
+    plt.figure()
+    for p in pkgs:
+        lambda_param = np.float(re.search('lambda=(.{4})', p).group(1))
+        # n_hidden = int(re.search('comp=(?P<comp>.{1,2,3})_', p).group('comp'))
+        n_hidden = int(re.search('comp=(.{1,3})_', p).group(1))
+        max_samples = int(re.search('max(.{1,4})_', p).group(1))
+        rest_samples = int(re.search('_r(.{1,4})_', p).group(1))
+        task_samples = int(re.search('_t(.{2,4})dbg', p).group(1))
+        
+        if (task_samples != 100) or (rest_samples == 1) or (rest_samples == 200):
+            continue
+        
+        if n_comp != n_hidden:
+            continue
+
+        print rest_samples
+        print p
+
+        dbg_acc_val_ = np.load(p)
+
+        # naccs.append(np.max(dbg_acc_val_))
+        naccs.append(dbg_acc_val_[-1])
+        nrests.append(rest_samples)
+
+    nrests = np.array(nrests)
+    naccs = np.array(naccs)
+    inds = np.argsort(nrests)
+    plt.plot(
+        nrests[inds],
+        naccs[inds])
+
+    # plt.title('Low-rank LR+AE L1=0.1 L2=0.1 res=3mm combined-loss')
+    plt.legend(loc='lower right', fontsize=9)
+    plt.yticks(np.linspace(0., 1., 11))
+    plt.xticks(nrests[inds], nrests[inds])
+    plt.ylabel('validation set accuracy')
+    plt.ylim(0., 1.05)
+    plt.xlabel('nrest')
+    plt.grid(True)
+    plt.show()
+    plt.savefig(op.join(WRITE_DIR,
+                'scarcities_ncomp=%i_100.png' % n_comp))

### 2015-06-03 nips3mm_serial.py

@@ -493,16 +493,46 @@ for n_comp in n_comps:
     # joblib.dump(estimator, outpath + '_est', compress=9)
 
     # 2-step approach 3: AE + LR
-    print('Compressing by autoencoder...')
-    compressor = AutoEncoder(
-            n_hidden=n_comp,
-            gain1=0.004,  # empirically determined by CV
-            learning_rate = np.float32(0.00001),  # empirically determined by CV,
-            max_epochs=500)
+    # print('Compressing by autoencoder...')
+    # compressor = AutoEncoder(
+    #         n_hidden=n_comp,
+    #         gain1=0.004,  # empirically determined by CV
+    #         learning_rate = np.float32(0.00001),  # empirically determined by CV,
+    #         max_epochs=500)
+    # compressor.fit(X_dev)
+    # 
+    # half2compr = compressor.transform(X_val)
+    # 
+    # print('Classifiying...')
+    # l1 = 0.1
+    # l2 = 0.1
+    # my_title = r'LR: L1=%.1f L2=%.1f res=3mm' % (
+    #     l1, l2
+    # )
+    # print(my_title)
+    # estimator = SEncoder(
+    #     gain1=0.004,  # empirically determined by CV
+    #     learning_rate = np.float32(0.00001),  # empirically determined by CV,
+    #     max_epochs=500, l1=l1, l2=l2)
+    # 
+    # estimator.fit(half2compr, y_val)
+    # 
+    # acc_AE = estimator.dbg_acc_val_
+    # print(acc_AE)
+    # print('AE: %.4f' % acc_AE[-1])
+    # 
+    # outpath = op.join(WRITE_DIR, 'AE-LR_ncomp=%i' % n_comp)
+    # np.save(outpath + '_acc', acc_AE)
+    # joblib.dump(estimator, outpath + '_est', compress=9)
+
+    # 2-step approach 4: PCA + LR
+    from sklearn.decomposition import PCA
+    print('Compressing by whitened PCA...')
+    compressor = PCA(n_components=n_comp, whiten=True)
     compressor.fit(X_dev)
-    
+
     half2compr = compressor.transform(X_val)
-    
+
     print('Classifiying...')
     l1 = 0.1
     l2 = 0.1
@@ -517,13 +547,11 @@ for n_comp in n_comps:
 
[…]

### 2015-06-03 paper_nips/paper.tex
OGs comments integrated

@@
 
-\title{Semi-supervised factored logistic regression for
-high-dimensional neuroimaging data}
+\title{Semi-Supervised Factored Logistic Regression for
+High-Dimensional Neuroimaging Data}
 
@@
 
-\nipsfinalcopy % Uncomment for camera-ready version
+%\nipsfinalcopy % Uncomment for camera-ready version
+\nipsfinalfalse
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A multinomial logistic regression is introduced that is
-constrained by joint rank with a representation-enforcing
+constrained by factored coefficients with a representation-enforcing
 autoencoder.
@@
 
-% Danilo: NIPS does actually not seem to have keywords
+% OUR keywords
 %\textbf{\\keywords}: curse of dimensionality; semi-supervised learning;
 %fMRI; systems neuroscience
+
+% official NIPS keywords
+\textbf{\\keywords}: Brain Imaging, Cognitive Science, Semi-Supervised Learning, Systems Biology
 
@@
 structure across time,
-independent component analysis (ICA) \cite{beckmann2005} is often used.
+independent component analysis (ICA) is often used
+\cite{beckmann2005}.
 It decomposes the BOLD (blood-oxygen level-dependent) signals into the
@@
 functionally interacting regions \cite{smith2009}.
-Similarly, sparse principal component analysis (SPCA) \cite{varoqu2011}
+Similarly, sparse principal component analysis (SPCA)
 has been used to
-separate BOLD signals into parsimonious network components.
+separate BOLD signals into parsimonious network components
+\cite{varoqu2011}.
 The extracted brain networks are probably
@@
   \caption {\textbf{Model architecture.}
-  Linear autoencoders find an optimal compression of 79,941 brain voxels
+  Linear autoencoders find an optimized compression of 79,941 brain voxels
   into $n$ unknown activity patterns by improving reconstruction from them.
@@
   Supervised multi-class learning on task data ($X_{task}$)
-  is thus guided by
+  can thus be guided by
   unsupervised decomposition of rest data ($X_{rest}$).
@@
 representations
-by optimizing reconstruction from them.
+via reconstruction under compression prior.
 %
@@
 %
-Thus, autoencoders can flexibly project the neuroimaging data
+Thus, autoencoders may flexibly project the neuroimaging data
 onto the main directions of variation.
@@
 resting-state autoencoders that
-favor the most neurobiologically
+privilege the most neurobiologically
 valid models in the hypothesis set.
@@
 8000 unlabeled, noise-cleaned rest maps with
-40 brain images from each of 200 randomly selected participants.
+40 brain maps from each of 200 randomly selected participants.
 
@@
 $\mathbf{b_0}$ and $\mathbf{b_1}$ are bias vectors.
-The optimal model parameters $\mathbf{W_0, b_0, b_1}$ are found by
+The model parameters $\mathbf{W_0, b_0, b_1}$ are found by
 minimizing the expected squared reconstruction error
@@
   \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
-  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right],
+  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
 \end{equation}
@@
 We thus search for a compression of the 79,941 voxel values into $n$ latent
-components that represent an optimal latent code for both
+components that represent a latent code optimized for both
 rest and task activity data.
@@
 The common objective was optimized
-in the neuroimaging data
-by gradient descent in the parameters
-of the 
-semi-supervised factored logistic regression (SS-F LogReg).
-The required gradients are easily obtained by using the chain rule to
-backpropagate error derivatives through the linear network.
-As solver, we chose \textit{rmsprop} \cite{rmsprop}.
-%, a mini-batch version of rprop.
-This procedure dictates an adaptive learning rate
+by gradient descent in the SS-F LogReg parameters.
+%
+The required gradients were obtained by using the chain rule to
+backpropagate error derivatives.
+We chose the \textit{rmsprop} solver \cite{rmsprop},
+a refinement of stochastic gradient descent.
+\textit{Rmsprop} dictates an adaptive learning rate
 for each model parameter by
 scaled gradients from a running average.
-Gradient normalization by \textit{rmsprop}
-is known to effectively exploit curvature information.
-We opted for a small batch size of $100$, given the high degree of
-redundancy in $X_{rest}$ and $X_{task}$.
-The matrix parameters were initalized by Gaussian random values multiplied
-by a gain of $0.004$. 
+The batch size was set to $100$
+(much expected redundancy in $X_{rest}$ and $X_{task}$),
+matrix parameters were initalized by Gaussian random values multiplied
+by $0.004$ (i.e., gain), and
+bias parameters were initalized to zero.
 %
-The bias parameters were initalized to zero.
-With a slight abuse of notation, let $\theta$ denote a component of $\theta$.
 The normalization factor and the update rule for $\theta$
@@
 
-where $0 < \rho < 1$ constitutes the decay rate. $\rho$ was set to
-$0.9$ to deemphasize the magnitude of the gradient.
-Further, $\alpha$ is the learning rate and $\epsilon$ a global damping factor.
-The hyper-parameter $\alpha$ was set to $0.00001$ by prior studies and $\epsilon$ was set to $10^{-6}$.
+where $f$ is the loss function computed on a minibatch sample at timestep $t$,
+$\alpha$ is the learning rate ($0.00001$),
+$\epsilon$ a global damping factor ($10^{-6}$),
+and
+$\rho$ the decay rate ($0.9$ to deemphasize the magnitude of the gradient).
 %
@@
 (stochastic gradient descent, adadelta, and adagrad) but found that
-\textit{rmsprop} converged faster and with higher generalization performance.
+\textit{rmsprop} converged faster and with
+similar or higher generalization performance.
 
@@
 rather than regularization in a strict sense \cite{abu1994hints}.
-Its purpose is not to prevent overfitting but to introduce
+Its purpose is to introduce
 % XXX Michael: I use f earlier up, is this f here necessary?
@@
  \textit{n}               & PCA + LogReg   & SPCA + LogReg  & ICA + LogReg  & AE + LogReg    & FLogReg              \\ \hline
- \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & 95,7 \%      \\
- \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & 97,3 \%      \\
- \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & 97,6 \%      \\
- \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & 97,4 \%      \\ \hline
+ \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & \textbf{95,7\%}      \\
+ \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & \textbf{97,3\%}      \\
+ \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & \textbf{97,6\%}      \\
+ \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & \textbf{97,4\%}      \\ \hline
  \end{tabular}
@@
 was assessed on 20\% of unseen HCP data.
-Additionally, the ``out-of-dataset performance'' of the learned decomposition
+Additionally, the ``out-of-study performance'' of the learned decomposition
 was assessed by using it as dimensionality reduction of an
@@
 To this end, ordinary logstic regression was juxtaposed
-with LR LogReg  at $\lambda=1$ (no autoencoder).
+with FLogReg  at $\lambda=1$ (no autoencoder).
 For this experiment, we increased the difficulty of the classification problem
@@
 as compared to voxel space, entails differences in
-generalization perforamnce of
+generalization performance of
 $\approx{17\%}$ (Figure \ref{fig_dimred}).
@@
 \caption{\textbf{Effect of richness in rest data}
-Gradient descent was performed on 2000 task and 2000 rest images.
+Gradient descent was performed on 2000 task and 2000 rest maps.
 These were drawn with replacement from identical versus varying
@@
 %
-Ordinary logistic regression thus yielded a mean correlation of
+Ordinary logistic regression thus yielded a mean correlation
 of $\rho=0.28$.
@@
 SS-F LogReg puts higher absolute weights on relevant structure.
-This reflect a higher signal-to-noise ratio, in part explained
+This reflects a higher signal-to-noise ratio, in part explained
 by the more BOLD-typical local contiguity.
@@
 %
-Hence, SS-F LogReg yielded class weights 
-that were much more similar to features of the respective training samples
-for any choice of $\lambda$.
-SS-F LogReg therefore
-captures genuine properties of neural activity patterns,
-rather than participant- or study-specific artefacts.
-XXX: But what you show here is mostly a denoising or regularization effect.
-
 \begin{figure}
@@
 and
-c) yields BOLD-typical local contiguity.
+c) yields BOLD-typical local contiguity (without enforcing a spatial prior).
 All values are z-scored and thresholded at the $75^{th}$ percentile.
@@
 \end{figure}
+%
+Hence, SS-F LogReg yielded class weights 
+that were much more similar to features of the respective training samples
+for any choice of $\lambda$.
+SS-F LogReg therefore
+captures genuine properties of neural activity patterns,
+rather than participant- or study-specific artefacts.
 
@@
 other commonly used nonlinearities (d) sigmoid, e) softplus,
-f) hyperbolic tangent) all led to smaller classification accuracies.
+f) hyperbolic tangent) all led to smaller classification accuracies,
+probably due to sample size limits.
 %
-\textbf{XXX: Please give explanations why this is the case. My feeling is that we're still not rich enough in the sample dimension to make these tricks worthwhile.}
 Further, g) ``pretraining'' of the bottleneck $\mathbf{W_0}$
@@
 %
-Finally, imposing imposing either j) only $\ell_1$-penalty or 
+Finally, imposing either j) only $\ell_1$-penalty or 
 k) only $\ell_2$-penalty was disadvantageous in all tested cases,
@@
 supervised learning question.
+%
+From a machine-learning perspective,
+factorization of the logistic regression weights
+can be viewed as transforming a ``multi-class problem'' into
+``multi-task learning''.
+%
 The higher generalization accuracy and feature recovery, comparing to
@@
 automatically learning a mapping to and from a brain-network space.
+%
 This domain-specifc classification algorithms
@@
 Neurobiologically,
-neural activity underlying defined mental operations
+brain activity underlying defined mental operations
 can be explained by linear combinations of the main activity
-patterns. That is,
+patterns.
+%
+That is,
 fMRI data probably concentrate near
-a low-dimensional manifold that correspond to peculiar brain networks combinations.
+a low-dimensional manifold of
+characteristic brain networks combinations.
 Extracting fundamental building blocks of brain organization might
@@
 human thought.
+%
 We hope that these first steps stimulate development towards
@@
 % grand perspective
-In the future, automatic reduction of brain images to
+In the future, automatic reduction of brain maps to
 their neurobiological essence
@@
 These promise structured integration
-of accumulating neuroscientific knowledge from neuroimaging databases.
+of neuroscientific knowledge accumulating
+in databases.
 Tractability by

### 2015-06-04 paper_nips/paper.tex
fixed style of equations, figures and tables

@@
 and reconstructs it back to $\mathbf{x'}$ by
-
 \begin{eqnarray}
-  \begin{split}
-    \mathbf{z} &= \mathbf{W_0} \mathbf{x} + \mathbf{b_0} \\
-    \mathbf{x'} &= \mathbf{W_1} \mathbf{z} + \mathbf{b_1}
-  \end{split}
+    \mathbf{z} = \mathbf{W_0} \mathbf{x} + \mathbf{b_0} \hspace{2cm}
+    \mathbf{x'} = \mathbf{W_1} \mathbf{z} + \mathbf{b_1},
   \label{eq:autoenc}
@@
 minimizing the expected squared reconstruction error
-
 \begin{equation}
@@
 \end{equation}
+
 %with \(f_{AE}(\mathbf x) = %
@@
 to a class $i \in \{1, \dots, l\}$
-
 \begin{equation}
   \begin{split}
-    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_{RL}(\mathbf x))
+    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_{RL}(\mathbf x)),
   \end{split}
@@
 The loss function is given by
-
 \begin{equation}
     \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
-\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1})
+\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1}).
 \label{eq:lr_loss}
@@
 logistic regression are interpolated in the common loss function
-
 % XXX Michael: In resolving conflict I opted for my change. Feel free to revert if that was intended.
@@
 % simultaneously at every bach in every epoch
-
 \begin{eqnarray}
@@
 % XXX partial derivatives? \partial and \nabla
-
 % XXX Michael: Probably remove the explanations following and just state values
 % XXX: Bertrand: what is f ?? Danilo: replaced f by more detailed notation
-
 where $f$ is the loss function computed on a minibatch sample at timestep $t$,
@@
 per-iteration tolerance of 1 * 10\textsuperscript{-8},
-$alpha=1$).
+$\alpha=1$).
 %
@@
 This effect was particularly pronounced when assuming few hidden dimensions.
-
 \begin{table}[h]
@@
  \resizebox{0.8\textwidth}{!}{%
- \begin{tabular}{lccccc}
+ \begin{tabular}{l|c|c|c|c|c}
  \hline
@@
  \end{table}
-
 \paragraph{Model performance.}
@@
 $\approx{17\%}$ (Figure \ref{fig_dimred}).
-
 \begin{figure}
 \begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/accuracies.png}
+\includegraphics[width=1.00\textwidth]{figures/accuracies.pdf}
 \end{centering}
@@
 \caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
-Depicts f1 scores for prediction of each of 38 classes across training epochs.
-Ordinary logistic regression operating in voxel space (\textit{left plot})
-converged faster but performed worse than
-factored logistic regression for few
-(\textit{middle plot}) and many (\textit{right plot}) latent modes.
+Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$).
+Multinomial logistic regression operating in voxel space (\textit{blue bars})
+was compared to 
+factored logistic regression operating in $20$
+(\textit{left plot}) and $100$ (\textit{right plot})
+latent modes (\textit{grey bars}).
 Autoencoder or rest data were not used for these analyses
 ($\lambda=1$).
-Hence, projecting the input data into a reduced space for classification
-yields higher class separability.
-Chance is at 2,6\%.
-XXX: ugly figure: use larger font for labels and titles.
-XXX: includes pdf figures, not a png file.
+Ordinary logistic regression yielded 77,7\% accuracy out of sample, while
+factored logistic regreesion scored at 94,4\% ($n=20$) and 94,2\% ($n=100$).
+Hence, compressing the voxel data into a component space for classification
+achieves higher task separability.
 }

### 2015-06-04 nips3mm_h38.py

@@ -1010,6 +1010,53 @@ plt.xlabel('Plain Vanilla LogReg')
 plt.ylabel('Low-Rank LogReg (n=100)')
 plt.savefig(op.join(WRITE_DIR, 'scattern_againstLR_%icomps.png' % 100))
 
+# barplot against vanilla
+plt.close('all')
+newmeth1 = np.load(op.join(WRITE_DIR, 'Low-rank_LR_AE_(combined_loss,_shared_decomp)_n_comp=20_L1=0.1_L2=0.1_lambda=1.00_res=3mm_spca20RSdbg_prfs_.npy'))
+newmeth2 = np.load(op.join(WRITE_DIR, 'Low-rank_LR_AE_(combined_loss,_shared_decomp)_n_comp=100_L1=0.1_L2=0.1_lambda=1.00_res=3mm_spca20RSdbg_prfs_.npy'))
+newmeth = (newmeth1, newmeth2)
+ordinary = np.load(op.join('nips3mm_h38_vanilla', 'LR_L1=0.1_L2=0.1_res=3mmdbg_prfs_.npy'))
+n_comps = [20, 100]
+for n_comp, newmeth in zip(n_comps, newmeth):
+    plt.figure()
+    my_width = 0.2
+    x_pos = 0
+    pos_list = list()
+    c_sslr = '#404040'
+    c_normal = '#51A2E0'
+    my_fs = 16
+    for i in np.arange(38):
+        if (i == 0):
+            plt.bar(x_pos, newmeth[-1, 2, i], width=my_width, color=c_sslr,
+                    label='Factored Logistic Regression')
+            plt.bar(x_pos, ordinary[-1, 2, i], width=my_width, color=c_normal,
+                    label='Ordinary Logistic Regression')
+        else:
+            plt.bar(x_pos, newmeth[-1, 2, i], width=my_width, color=c_sslr)
+            plt.bar(x_pos, ordinary[-1, 2, i], width=my_width, color=c_normal)
+        pos_list.append(x_pos)
+        x_pos += my_width
+
+    plt.ylim(0, 1.05)
+    plt.xlim(0, x_pos)
+    plt.xlabel('mental task')
+    plt.title('$n=%i$' % n_comp, fontsize=my_fs)
+    plt.xticks(
+        np.array(pos_list[::2]) + my_width / 2,
+        np.arange(0, 38)[::2] + 1,
+        fontsize=my_fs)
[…]

### 2015-06-04 paper_nips/paper.tex
interm

@@
 The common objective was optimized
-by gradient descent in the SS-F LogReg parameters.
+by gradient descent in the SS-FLogReg parameters.
 %
@@
 \paragraph{Model performance.}
-SS-F LogReg was subsequently trained ($500$ epochs) across parameter choices
+SS-FLogReg was subsequently trained ($500$ epochs) across parameter choices
 for the hidden components ($n=5, 20, 100$) and
@@
 Assuming 5 latent directions of variation should yield models with
-higher bias and smaller variance then SS-F LogReg with 100 latent directions.
+higher bias and smaller variance then SS-FLogReg with 100 latent directions.
 %
@@
 \vspace{-0.2cm}
- \caption{\textbf{Performance of SS-F LogReg across model parameter choices.}
+ \caption{\textbf{Performance of SS-FLogReg across model parameter choices.}
  Chance is at 5,6\%.
@@
 %
-For SS-F LogReg ($\lambda=0.25, 0.5, 0.75, 1$),
+For SS-FLogReg ($\lambda=0.25, 0.5, 0.75, 1$),
 a per-class-weight map was computed by matrix
@@
 Consequently,
-SS-F LogReg puts higher absolute weights on relevant structure.
+SS-FLogReg puts higher absolute weights on relevant structure.
 This reflects a higher signal-to-noise ratio, in part explained
@@
 %
-Conversely, SS-F LogReg puts lower probability mass on irrelevant structure.
+Conversely, SS-FLogReg puts lower probability mass on irrelevant structure.
 Despite lower interpretability, the
@@
 %
-Hence, SS-F LogReg yielded class weights 
+Hence, SS-FLogReg yielded class weights 
 that were much more similar to features of the respective training samples
 for any choice of $\lambda$.
-SS-F LogReg therefore
+SS-FLogReg therefore
 captures genuine properties of neural activity patterns,

### 2015-06-04 paper_nips/paper.tex
results: added notice on reward/punishment cognitive tasks

@@
 $\approx{17\%}$ (Figure \ref{fig_dimred}).
+%
+Notably,
+the cognitive tasks on reward and punishment processing
+are among the least predicted with ordinary but well predicted with
+factored logistic regression
+(tasks 3 and 4 in Figure \ref{fig_dimred}).
+These experimental conditions have indeed been reported to exhibit
+highly similar neural activity patterns in
+GLM analyses of that dataset \cite{barch2013}.
+Consequently, also local activity differences
+(in the striatum and visual cortex in this case) can
+be successfully captured by brain-network modelling.
+%
 \begin{figure}
@@
 \caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
-Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$).
+Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$)
+sorted from lowest to highest.
 Multinomial logistic regression operating in voxel space (\textit{blue bars})
@@
 % Conceptual recap+gain of current paper
-Using the flexibility of autoencoders, we
-learn the optimal decomposition from high-dimensional
-voxel brain space into the most important activity patterns for a
-supervised learning question.
+Using the flexibility of factored models,
+we learn the low-dimensional representation from high-dimensional
+voxel brain space that is most important for
+prediction of cognitive task sets.
 %
@@
 factorization of the logistic regression weights
-can be viewed as transforming a ``multi-class problem'' into
-``multi-task learning''.
+can be viewed as transforming a
+``multi-class classification problem''
+into a ``multi-task learning problem''.
 %
@@
 %
-This domain-specifc classification algorithms
+This domain-specifc learning algorithm
 encourages departure from the artificial and statistically
@@
 We hope that these first steps stimulate development towards
-powerful semi-supervised classification methods in systems neuroscience.
+powerful semi-supervised representation extraction
+in systems neuroscience.
 

### 2015-06-04 nips3mm_h38.py

@@ -1025,15 +1025,16 @@ for n_comp, newmeth in zip(n_comps, newmeth):
     c_sslr = '#404040'
     c_normal = '#51A2E0'
     my_fs = 16
+    inds = np.argsort(newmeth[-1, 2, :])
     for i in np.arange(38):
         if (i == 0):
-            plt.bar(x_pos, newmeth[-1, 2, i], width=my_width, color=c_sslr,
+            plt.bar(x_pos, newmeth[-1, 2, inds[i]], width=my_width, color=c_sslr,
                     label='Factored Logistic Regression')
-            plt.bar(x_pos, ordinary[-1, 2, i], width=my_width, color=c_normal,
+            plt.bar(x_pos, ordinary[-1, 2, inds[i]], width=my_width, color=c_normal,
                     label='Ordinary Logistic Regression')
         else:
-            plt.bar(x_pos, newmeth[-1, 2, i], width=my_width, color=c_sslr)
-            plt.bar(x_pos, ordinary[-1, 2, i], width=my_width, color=c_normal)
+            plt.bar(x_pos, newmeth[-1, 2, inds[i]], width=my_width, color=c_sslr)
+            plt.bar(x_pos, ordinary[-1, 2, inds[i]], width=my_width, color=c_normal)
         pos_list.append(x_pos)
         x_pos += my_width
 
@@ -1049,9 +1050,10 @@ for n_comp, newmeth in zip(n_comps, newmeth):
     plt.ylabel('f1 score')
     plt.xlabel('task', fontsize=my_fs)
     plt.ylabel('f1 score', fontsize=my_fs)
-    plt.legend(loc='lower left', fontsize=my_fs - 2.5)
+    plt.legend(loc='lower right', fontsize=my_fs + 1)
+    plt.grid(True)
     plt.show()
-    plt.savefig(op.join(WRITE_DIR, 'bars_against_LR_%icomps.png' % n_comp))
+    plt.savefig(op.join(WRITE_DIR, 'bars_against_LR_%icomps_sorted.png' % n_comp))
 
 plt.xlim(0, 1.)
 plt.ylim(0, 1.)

### 2015-06-04 paper_nips/paper.tex
abstract

--- 2015-06-04 paper_nips/paper.tex
@@
 larger mental task sets
-necessitates adequate representation of the observations.
+necessitates adequate representations of the observations.
 %
-We therefore propose to optimize
-representation modelling and task classification in
+We therefore propose to blend
+representation modelling and task classification into
 a unified statistical learning problem.
@@
 A multinomial logistic regression is introduced that is
-constrained by factored coefficients with a representation-enforcing
-autoencoder.
+constrained by factored coefficients and coupled with an autoencoder.
 %

### 2015-06-04 paper_nips/paper.tex
introduction

@@
 during ongoing random thought without controlled environmental stimulation.
-In fact, a large proportion of the BOLD signal variation
+In fact, a large portion of the BOLD signal variation
 is known
@@
 as observed using ICA \cite{smith2009}.
+% XXX: you don't discuss the absence of information sharing across tasks / cognitions, i.e. the multi-task aspect of your work.
 
@@
 approach that is
-mixed (i.e., using rest and task data)
-and 
+mixed (i.e., using rest and task data),
+multitask and 
 factored (i.e., performing network decomposition).
@@
 projectors to the discovered set of basis functions
-whose linear combinations are learned
-by the second layer \cite{olshausen96}.
+which are linearly combined 
+by the second layer to perform predictions\cite{olshausen96}.
 % XXX Michael: ^ This needs to be changed around: This first layer obtains
@@
 1) co-analyzes many problems simultaneously
-(thus find shared representations by ``multi-task learning'')
+(thus finds shared representations by ``multi-task learning'')
 and
-2) exploits unlabeled data
-(that span a space of meaningful configurations).
+2) exploits unlabeled data,
+as they span a space of meaningful configurations.
 

### 2015-06-04 paper_nips/paper.tex
A pass on methods

@@
 To this end, the HCP-derived network decompositions were used as preliminary
-step in the classifcation problem of another large sample.
+step in the classification problem of another large sample.
 The ARCHI dataset \cite{pinel07} provides activity maps from
@@
 from 18 diverse tasks acquired in 78 participants.
+% XXX: details on data should be given as supplementary material. the section is boring for most NIPS readers
 
@@
 input and the learned representation.
+% XXX Give a ref ?
 Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
@@
   \begin{split}
-    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_{RL}(\mathbf x)),
+    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_\mathcal{LR}(\mathbf x)),
   \end{split}
@@
 \end{equation}
-where \(f_{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
+where \(f_\mathcal{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
 multinomial logits and \(\softmax_i(x) = \exp(x_i)/\sum_j\exp(x_j)\).
@@
 \(\Omega(\theta)\) represents a regularization, for which we choose the
-ElasticNet-type, i.e., a combination of $\ell_1$ and $\ell_2$ penalties
-$\forall p \in \theta$.
+ElasticNet-type, i.e., a combination of $\ell_1$ and $\ell_2$ penalties.
+%$\forall p \in \theta$.
 
@@
   \begin{split}
-    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t+1)}} + (1 - \rho)\left(\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})\right)^2
+    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t)}} + (1 - \rho)\left(\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})\right)^2
+% XXX: I have changed the first equation, please check
 %v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
@@
 $\alpha$ is the learning rate ($0.00001$),
-$\epsilon$ a global damping factor ($10^{-6}$),
+$\epsilon$ a global damping factor ($10^{-6}$)
 and

### 2015-06-04 paper_nips/paper.tex
Pass on the end

@@
 by gradient descent in the SS-FLogReg parameters.
+% XXX you haven't defined FLogReg
 %
@@
 \paragraph{Serial versus parallel structure discovery and classification.}
-We first tested for an advantage of combined unsupervised decomposition
+We first tested whether there is a substantial advantage in combining unsupervised decomposition
 and supervised classification learning.
-We benchmarked against
+We benchmarked our approach against
 performing data reduction on the (unlabeled)
@@
 %
-These explorations attest to the advantage of directly searching for
+This experiment establishes the advantage of directly searching for
 classification-relevant structure in the fMRI data,
@@
 Second, the higher the number of latent components $n$,
-the higher the out-of-study performance with small choices for $\lambda$.
+the higher the out-of-study performance with small values of $\lambda$.
 This suggests that the presence of more rest-data-inspired hidden components
@@
 \paragraph{Individual effects of dimensionality reduction and rest data.}
-We first quantified the gain of introducing a bottleneck layer
-disregarding the autoencoder.
+We first quantified the impact of introducing a bottleneck layer
+that disregards the autoencoder.
 To this end, ordinary logstic regression was juxtaposed
@@
 %
-This domain-specifc learning algorithm
+This domain-specific learning algorithm
 encourages departure from the artificial and statistically

### 2015-06-04 nips3mm_scarcity.py
blah

@@ -1,3 +1,4 @@
+
 """
 HCP: semi-supervised network decomposition by low-rank logistic regression
 """
@@ -518,8 +519,8 @@ dbg_epochs_ = np.load(pkgs[0])
 
 d = {
     'in-sample accuracy': '/*dbg_acc_train*.npy',
-    'out-of-sample accuracy': '/*dbg_acc_val_*.npy',
-    'accuracy other ds': '/*dbg_acc_other_ds_*.npy',
+    # 'out-of-sample accuracy': '/*dbg_acc_val_*.npy',
+    # 'accuracy other ds': '/*dbg_acc_other_ds_*.npy',
     # 'loss ae': '/*dbg_ae_cost_*.npy',
     # 'loss lr': '/*dbg_lr_cost_*.npy',
     # 'loss combined': '/*dbg_combined_cost_*.npy'
@@ -578,7 +579,10 @@ for k, v in d.iteritems():
                     k.replace(' ', '_') + '_%icomps.png' % n_comp))
 
 n_comps = [20]
-pkgs = glob.glob(RES_NAME + '/*dbg_acc_val_*.npy')
+import re
+pkgs = glob.glob(RES_NAME + '/*dbg_epochs*.npy')
+dbg_epochs_ = np.load(pkgs[0])
+pkgs = glob.glob(RES_NAME + '/*dbg_acc_train*.npy')
 for n_comp in n_comps:  # 
     plt.figure()
     for p in pkgs:
@@ -589,7 +593,7 @@ for n_comp in n_comps:  #
         rest_samples = int(re.search('_r(.{1,4})_', p).group(1))
         task_samples = int(re.search('_t(.{2,4})dbg', p).group(1))
         
-        if (task_samples != 100):
+        if (task_samples != 100) or (rest_samples != 1000):
             continue
         
         if n_comp != n_hidden:
@@ -606,9 +610,10 @@ for n_comp in n_comps:  #
         cur_label += 'nrest%i' % rest_samples
         cur_label += '/'
         cur_label += 'ntask=%i' % task_samples
+        npoints = min(len(dbg_acc_val_), len(dbg_epochs_))
         plt.plot(
-            dbg_epochs_,
-            dbg_acc_val_,
+            dbg_epochs_[:npoints],
+            dbg_acc_val_[:npoints],
             label=cur_label)
     plt.title('Low-rank LR+AE L1=0.1 L2=0.1 res=3mm combined-loss')
     plt.legend(loc='lower right', fontsize=9)

### 2015-06-05 paper_nips/paper.tex
minor typos fixed

@@
 \usepackage{ulem}
+\usepackage{multirow}
 
@@
 larger mental task sets
-necessitates adequate representations of the observations.
+necessitates adequate representations for the observations.
 %
@@
 approach that is
-mixed (i.e., using rest and task data),
-multitask and 
-factored (i.e., performing network decomposition).
+\textit{mixed} (i.e., using rest and task data),
+\textit{factored} (i.e., performing network decomposition), and
+\textit{multi-task} (i.e., capitalize on neural representations shared across
+mental operations).
 %
@@
 Autoencoders (AE)
-are often one-layered learning models 
+are layered learning models
 that condense the input data to
@@
 which are linearly combined 
-by the second layer to perform predictions\cite{olshausen96}.
+by the second layer to perform predictions \cite{olshausen96}.
 % XXX Michael: ^ This needs to be changed around: This first layer obtains
@@
 We therefore contribute a computational framework that
-1) co-analyzes many problems simultaneously
+1) analyzes many problems simultaneously
 (thus finds shared representations by ``multi-task learning'')
@@
 \section{Methods}
+%
+\paragraph{Data.}
+As the currently biggest openly-accessible reference dataset,
+we chose resources from the Human Connectome Project (HCP)
+\cite{barch2013}.
+Neuroimaging task data with labels of ongoing cognitive processes
+were drawn from 500 unrelated,
+healthy HCP participants (cf. Appendix for details on datasets).
+18 HCP tasks 
+were selected that are known to elicit reliable neural activity
+across participants (Table \ref{table_tasks}).
+In sum, the HCP task data incorporated 8650 first-level activity maps
+from 18 diverse paradigms administered to 498 participants (2 removed
+due to incomplete data).
+All maps were resampled to a common $60\times72\times60$ space of
+3mm isotropic voxels and gray-matter masked (at least 10\% tissue
+probability).
+The supervised analyses were based on labeled HCP task maps with
+79,941 voxels of interest representing z-values in gray matter.
+
+\begin{table}[h]
+  \resizebox{0.98\textwidth}{!}{%
+  \begin{tabular}{l|l|l}
+    \hline
+  {\bf Cognitive Task} & {\bf Stimuli}                         & {\bf Instruction for participants}                                                \\ \hline
+  1 Reward             & \multirow{2}{*}{Card game}            & \multirow{2}{*}{Guess the number of a mystery card for gain/loss of money}        \\ \cline{1-1}
+  2 Punish             &                                       &                                                                                   \\ \hline
+  3 Shapes             & Shape pictures                        & Decide which of two shapes matches another shape geometrically                    \\ \hline
+  4 Faces              & Face pictures                         & Decide which of two faces matches another face emotionally                        \\ \hline
+  5 Random             & \multirow{2}{*}{Videos with objects}  & \multirow{2}{*}{Decide whether the circles, squares, triangles act intentionally} \\ \cline{1-1}
+  6 Theory of mind     &                                       &                                                                                   \\ \hline
+  7 Mathematics        & Spoken numbers                        & Complete addition and subtraction problems                                        \\ \hline
+  8 Language           & Auditory stories                      & Choose answer about the topic of the story                                        \\ \hline
+  9 Tongue movement    & \multirow{3}{*}{Visual cues}          & Move tongue                                                                       \\ \cline{1-1} \cline{3-3} 
+  10 Food movement     &                                       & Squeezing of the left of right toe                                                \\ \cline{1-1} \cline{3-3} 
+  11 Hand movement     &                                       & Tapping of the left or right finger                                               \\ \hline
+  12 Matching          & \multirow{2}{*}{Shapes with textures} & Decide whether two objects match in shape or texture                             \\ \cline{1-1} \cline{3-3} 
+  13 Relations         &                                       & Decide whether object pairs differ both along either shape or texture             \\ \hline
+  14 View Bodies       & Pictures                              & Passive watching                                                                   \\ \hline
+  15 View Faces        & Pictures                              & Passive watching                                                                   \\ \hline
+  16 View Places       & Pictures                              & Passive watching                                                                   \\ \hline
+  17 View Tools        & Pictures                              & Passive watching                                                                   \\ \hline
+  18 Two-Back          & Various pictures                      & Indicate whether current stimulus is the same as two items earlier                \\ \hline
+  \end{tabular}
+}
+\vspace{-0.2cm}
+\caption{\textbf{Description of psychological tasks to predict.}}
+\label{table_tasks}
+\end{table}
+
+These labeled data were complemented by unlabeled activity maps
+from HCP acquisitions of unconstrained resting-state activity
+\cite{smith2013resting}.
+These reflect brain activity in the absence of controlled thought.
+In sum, the HCP rest data concatenated
+8000 unlabeled, noise-cleaned rest maps with
+40 brain maps from each of 200 randomly selected participants.
+
+We were further interested in the utility of the
+optimized low-rank projection
+in one task dataset for dimensionality reduction in another task dataset.
+To this end, the HCP-derived network decompositions were used as preliminary
+step in the classification problem of another large sample.
+The ARCHI dataset \cite{pinel07} provides activity maps from
+diverse experimental tasks, including auditory and visual perception, motor action,
+reading, language comprehension and mental calculation.
+Analogous to HCP data, the second task dataset incorporated 1404
+labeled, grey-matter masked, and z-scored activity maps
+from 18 diverse tasks acquired in 78 participants.
+
+The labeled and unlabeled data were fed into a linear statistical model
+composed of an autoencoder and dimensionality-reducing logistic regression.
+
+\paragraph{Linear autoencoder.}
+The affine autoencoder takes the input 
+$\mathbf{x}$ and projects it into a coordinate system of
+latent representations $\mathbf{z}$
+and reconstructs it back to $\mathbf{x'}$ by
+\begin{eqnarray}
+    \mathbf{z} = \mathbf{W_0} \mathbf{x} + \mathbf{b_0} \hspace{2cm}
+    \mathbf{x'} = \mathbf{W_1} \mathbf{z} + \mathbf{b_1},
+  \label{eq:autoenc}
+\end{eqnarray}
+
+where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79{,}941$
+voxel values from each
+rest map,
+$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$-dimensional hidden state
+(i.e., distributed neural activity patterns), and 
+$\mathbf{x' \in \mathbb{R}^{d}}$
+is the reconstruction vector of the original activity map
+from the hidden variables. 
+% XXX use bold letters for vectors and matrices
+Further, $\mathbf{W_0}$ denotes the weight matrix that
+transforms
+from input space into the hidden space (encoder),
+$\mathbf{W_1}$ is the weight matrix for back-projection
+from the hidden variables to the
+output space (decoder).
+$\mathbf{b_0}$ and $\mathbf{b_1}$ are corresponding bias vectors.
+The model parameters $\mathbf{W_0, b_0, b_1}$ are found by
+minimizing the expected squared reconstruction error
+\begin{equation}
+  \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
+  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
+\end{equation}
+
+%with \(f_{AE}(\mathbf x) = %
+%\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}\)
+% the autoencoder.
+% XXX Michael: I don't know to what extent the following makes sense or is
+% XXX correct, so I am commenting it
+% XXX Danilo: I found this in Bengio's papers somewhere
+This reconstruction error criterion equates with
+maximizing a lower bound on the mutual information between
+input and the learned representation.
+% XXX Give a ref ?
+Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
+$\mathbf{W_0} = \mathbf{W_1^T}$. In particular, this means that the weights
+learned are forced to take a two-fold function: That of signal 
+\textit{analysis} and that of signal \textit{synthesis}.
+The first layer \textit{analyzes}
+the data to obtain the cleanest latent representation,
+while the second
+layer represents building blocks from which to \textit{synthesize} the data
+using the latent activations.
+Tying these processes together makes the analysis
+layer interpretable and pulls all non-zero singular values towards 1. 
+Nonlinearities were not applied to the
+activations in the first layer.
+
+\paragraph{Factored logistic regression.}
+Lossy compression by a low-dimensional bottleneck
+is also imposed by the first layer of the factored
+multinomial logistic regression.
+It gives the probability of an input $\mathbf{x}$ to belong
+to a class $i \in \{1, \dots, l\}$
+\begin{equation}
+  \begin{split}
+    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_\mathcal{LR}(\mathbf x)),
+  \end{split}
+  \label{eq:lr}
+\end{equation}
+where \(f_\mathcal{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
+multinomial logits and \(\softmax_i(x) = \exp(x_i)/\sum_j\exp(x_j)\).
+The matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
+transforms the input $\mathbf{x \in \mathbb{R}^{d}}$
+into $n$ latent components
+and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxl}}$
+projects the latent components
+onto hyperplanes that reflect $l$ label probabilities.
+$\mathbf{c_0}$ and $\mathbf{c_1}$ are
+bias vectors.
+The loss function is given by
+\begin{equation}
+    \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
+\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1}).
+\label{eq:lr_loss}
+\end{equation}
+
+\paragraph{Layer combination.}
+Importantly, the optimization problem of the linear autoencoder
+and the factored logistic regression
+are linked on two levels. First, their transformation matrices mapping from
+input to the latent space are tied
+\begin{eqnarray}
+  \mathbf{V_0} = \mathbf{W_0}.
+\end{eqnarray}
+We thus search for a compression of the 79,941 voxel values into $n$ latent
+components that represent a latent code optimized for both
+rest and task activity data.
+Second, the objectives of the autoencoder and the factored
+logistic regression are interpolated in the common loss function
+% XXX Michael: In resolving conflict I opted for my change. Feel free to revert if that was intended.
+\begin{equation}
+{\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
++ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega(\theta).
+  \label{eq:loss_equ}
+\end{equation}
+
+In so doing, we search for the combined model parameters
+$\theta=\{\mathbf{V_0,V_1,c_0, c_1, b_0, b_1}\}$
+with respect to the (unsupervised) reconstruction error and the
+(supervised) task classification.
+${\mathcal{L_{AE}}}$ is devided by ${N_{X_{rest}}}$ to equilibrate both
+loss terms to the same order of magnitude.
+\(\Omega(\theta)\) represents an ElasticNet-type regularization
+that combines $\ell_1$ and $\ell_2$ penalty terms.
+%$\forall p \in \theta$.
+
+% XXX Michael: To sum up the math in this paragraph: We need to decide what
+% XXX the \mathcal L are supposed to mean - sample wise loss or expected loss
+% XXX or some global deterministic loss (the latter would be bad, because we)
+% XXX are not optimizing it. Right now it looks like a sample-wise loss.
+
+\paragraph{Optimization.}
+The common objective was optimized
+by gradient descent in the SSFLogReg parameters.
+%^
+The required gradients were obtained by using the chain rule to
+backpropagate error derivatives.
+We chose the \textit{rmsprop} solver \cite{rmsprop},
+a refinement of stochastic gradient descent.
+\textit{Rmsprop} dictates an adaptive learning rate
+for each model parameter by
+scaled gradients from a running average.
+The batch size was set to $100$
+(much expected redundancy in $X_{rest}$ and $X_{task}$),
+matrix parameters were initalized by Gaussian random values multiplied
+by $0.004$ (i.e., gain), and
+bias parameters were initalized to $0$.
+
+The normalization factor and the update rule for $\theta$
+are given by
+% XXX Michael: It is crucial, but nowhere mentioned, that the optimization is
+% XXX done in an alternating manner. The whole structure relies on this, by
+% XXX construction it is vulnerable to optimizing too much towards labels or
+% XXX rest at a time. I also identify this as one of the major weak points
+% XXX of the method, so it needs to be stated in a way that is not easily
+% XXX attackable.
+% XXX Maybe we could also talk less about RMSprop.
+%
+% XXX Danilo: I think we are not alternating, Xrest and Xtask go in
+% simultaneously at every bach in every epoch
+\begin{eqnarray}
+  \begin{split}
+    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t)}} + (1 - \rho)\left(\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})\right)^2
+% XXX: I have changed the first equation, please check
+%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
+\\
+\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
+  \end{split}
+\end{eqnarray}
+% XXX Michael: Why, in the formula, do we have two different notations for
+% XXX partial derivatives? \partial and \nabla
+% XXX Michael: Probably remove the explanations following and just state values
+% XXX: Bertrand: what is f ?? Danilo: replaced f by more detailed notation
+where $f$ is the loss function computed on a minibatch sample at timestep $t$,
+$\alpha$ is the learning rate ($0.00001$),
+$\epsilon$ a global damping factor ($10^{-6}$)
+and
+$\rho$ the decay rate ($0.9$ to deemphasize the magnitude of the gradient).
+%
+Note that we have also experimented with other solvers
+(stochastic gradient descent, adadelta, and adagrad) but found that
+\textit{rmsprop} converged faster and with
+similar or higher generalization performance.
+
+\paragraph{Hints.}
+In fact, the constraint by a rest-data autoencoder qualifies as a
+\textit{hint}
+rather than regularization in a strict sense \cite{abu1994hints}.
+Its purpose is to introduce
+% XXX Michael: I use f earlier up, is this f here necessary?
+prior information on
+\textit{known} properties of the \textit{unknown} target function $f$.
+Rather than only relying on input-output pairs in the learning process,
+we thus narrow our hypothesis set to the biologically most plausible solutions.
+That is, we reduce the search space in a way that
+is compatible with the expected representation of BOLD activity signals.
+
+\paragraph{Implementation.}
+The analyses were performed in Python.
+We used \textit{nilearn} to handle
+the large quantities of neuroimaging data 
+\cite{abrah14}
+and
+\textit{Theano} for automatic, numerically stable
+differentiation of symbolic computation graphs
+\cite{bastien2012theano, bergstra2010theano}.
+All Python scripts that generated the results are
+accessible online for reproducibility and reuse
+\url{http://github.com/anonymous/anonymous}.
+% url{http://github.com/banilo/nips2015}.
+
+\section{Experimental Results}
+\paragraph{Serial versus parallel structure discovery and classification.}
+We first tested whether there is a substantial advantage in combining unsupervised decomposition
+and supervised classification learning.
+We benchmarked our approach against
+performing data reduction on the (unlabeled)
+first half of the HCP task data by PCA, SPCA, ICA, and
+AE ($n=5, 20, 50, 100$ components)
+and learning classification models in the (labeled) second half
+by ordinary logistic regression.
+%
+PCA reduced the dimensionality of the task data by
+finding orthogonal network components using a change of basis
+(whitening).
+%
+SPCA separated the BOLD signals into
+network components with few regions by
+a regression-type optimization problem constrained by
+$\ell_1$ penalty
+(no orthogonality assumptions, 1000 maximum iterations,
+per-iteration tolerance of 1 * 10\textsuperscript{-8},
+$\alpha=1$).
+%
+ICA performed iterative blind source separation
+by a parallel FASTICA implementation (200 maximum iterations,
+per-iteration tolerance of 0.0001,
+initialized by random mixing matrix, whitening).
+%
+AE found a code of latent representations by optimizing projection
+into a bottleneck
+(500 iterations, same implementation as below for rest data).
+%
+The second half of the task data was projected onto the
+latent components discovered in its first half.
+Only the ensuing component loadings were submitted to ordinary
+logistic regression
+(one hidden layer, $\ell_1=0.1$, $\ell_2=0.1$, 500 iterations).
+%
+These serial two-step approaches were
+compared against parallel decomposition and classification by SSFLogReg
+(two hidden layers, $\lambda=1$, $\ell_1=0.1$, $\ell_2=0.1$,
+500 iterations).
+Importantly, all trained classification models were tested
+on a large, unseen test set (20\% of data) in the present analyses.
+%
+Across choices for $n$, SSFLogReg
+achieved more than 95\% out-of-sample accuracy, whereas
+supervised learning based on PCA, SPCA, ICA, and AE loadings
+ranged from 32\% to 87\%
+(Table \ref{table_one}).
+%
+This experiment establishes the advantage of directly searching for
+classification-relevant structure in the fMRI data,
+rather than solving the supervised and unsupervised problems independently.
+This effect was particularly pronounced when assuming few hidden dimensions.
+\begin{table}[h]
+ \centering
+ \resizebox{0.8\textwidth}{!}{%
+ \begin{tabular}{l|c|c|c|c|c}
+ \hline
+ \textit{n}               & PCA + LogReg   & SPCA + LogReg  & ICA + LogReg  & AE + LogReg    & SSFLogReg            \\ \hline
+ \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & \textbf{95,7\%}      \\
+ \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & \textbf{97,3\%}      \\
+ \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & \textbf{97,6\%}      \\
+ \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & \textbf{97,4\%}      \\ \hline
+ \end{tabular}
+ }
+ \vspace{-0.2cm}
+ \caption{\textbf{Serial versus parallel dimensionality reduction and classification.}
+ Chance is at 5,6\%.
+ }
+ \label{table_one}
+ \end{table}
+\paragraph{Model performance.}
+SSFLogReg was subsequently trained ($500$ epochs) across parameter choices
+for the hidden components ($n=5, 20, 100$) and
+the balance between autoencoder and logistic regression
+($\lambda=0, 0.25, 0.5, 0.75, 1$).
+Assuming 5 latent directions of variation should yield models with
+higher bias and smaller variance then SSFLogReg with 100 latent directions.
+%
+Given the 18-class problem of HCP, setting $\lambda$ to $0$
+consistently yields generalization performance
+at chance-level (5,6\%) because
+only the unsupervised layer of the estimator is optimized.
+%
+At each epoch (i.e., iteration over the data),
+the out-of-sample performance of the trained classifier
+was assessed on 20\% of unseen HCP data.
+Additionally, the ``out-of-study performance'' of
+the learned decomposition ($\mathbf{W_0}$)
+was assessed by using it as dimensionality reduction of an
+independent labeled dataset (i.e., ARCHI) and conducting ordinary
+logistic regression on the ensuing component loadings.
+
+\begin{table}[h]
+  \centering
+  \resizebox{1.00\textwidth}{!}{%
+\tabcolsep=0.05cm
+\begin{tabular}{l|ccccc|ccccc|ccccc}
+  \hline
+                            & \multicolumn{5}{c|}{\textit{n = 5}}                                                                                                                                                 & \multicolumn{5}{c|}{\textit{n = 20}}                                                                                                                                                      & \multicolumn{5}{c}{\textit{n = 100}}                                                                           \\ \hline
+  \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
+  Out-of-sample\\accuracy   & \textit{6,0\%}                     & 88,9\%                             & 95,1\%                            & \textbf{96,5\%}                    & 95,7\%                           & \textit{5,5\%}                   & 97,4\%                               & \textbf{97,8\%}                     & 97,3\%                               & 97,3\%                             & \textit{6,1\%}   & 97,2\%           & 97,0\%           & \textbf{97,8\%}  & 97,4\%                             \\
+  Precision (mean)          & \textit{5,9\%}                     & 87,0\%                             & 94,9\%                            & \textbf{96,3\%}                    & 95,4\%                           & \textit{5,1\%}                   & \textbf{97,4\%}                      & 97,1\%                              & 97,0\%                               & 97,0\%                             & \textit{5,9\%}   & 96,9\%           & 96,5\%           & \textbf{97,5\%}  & 96,9\%                             \\
+  Recall (mean)             & \textit{5,6\%}                     & 88,3\%                             & 95,2\%                            & \textbf{96,6\%}                    & 95,7\%                           & \textit{4,6\%}                   & \textbf{97,5\%}                      & 97,5\%                              & 97,4\%                               & 97,4\%                             & \textit{7,2\%}   & 97,2\%           & 97,2\%           & \textbf{97,9\%}  & 97,4\%                             \\
+  F1 score (mean)           & \textit{4,1\%}                     & 86,6\%                             & 94,9\%                            & \textbf{96,4\%}                    & 95,4\%                           & \textit{3,8\%}                   & \textbf{97,4\%}                      & 97,2\%                              & 97,1\%                               & 97,1\%                             & \textit{5,3\%}   & 97,0\%           & 96,7\%           & \textbf{97,7\%}  & 97,2\%                             \\ \hline
+  Out-of-study\\accuracy    & \textit{39,4\%}                    & 60,8\%                             & 54,3\%                            & 60,7\%                             & \textbf{62,9\%}                  & \textit{77,0\%}                  & 79,7\%                               & \textbf{81,9\%}                     & 79,7\%                               & 79,4\%                             & 79,2\%           & \textbf{82,2\%}  & 81,7\%           & 81,3\%           & \textit{75,8\%}                    \\ \hline
+  \end{tabular}
+}
+\vspace{-0.2cm}
+ \caption{\textbf{Performance of SSFLogReg across model parameter choices.}
+ Chance is at 5,6\%.
+ }
+  \label{table_two}
+\end{table}
+
+We made several noteworthy observations (Table \ref{table_two}).
+%
+First, the most supervised estimator ($\lambda=1$) achieved in no
+instance the best accuracy, precision, recall or f1 scores on HCP data.
+Classification by SSFLogReg is therefore facilitated by
+imposing structure from the unlabeled rest data.
+%
+Second, the higher the number of latent components $n$,
+the higher the out-of-study performance with small values of $\lambda$.
+This suggests that the presence of more rest-data-inspired hidden components
+results in more effective feature representations in unrelated task data.
+%
+Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
+decomposition matrix ($\lambda=0$) resulted in
+noninferior out-of-study performance
+of 77,0\% and 79,2\%,
+respectively (Table \ref{table_two}).
+This confirms that guiding model learning by task-unrelated structure
+extracts features of general relevance
+beyond the supervised problem at hand.
+
+% Depicts f1 scores for prediction of each of 38 classes across training epochs.
+% Ordinary logistic regression operating in voxel space (\textit{left plot})
+% converged faster but performed worse than
+% factored logistic regression for few
+% (\textit{middle plot}) and many (\textit{right plot}).
+% Autoencoder or rest data were not used for these analyses
+% ($\lambda=1$).
+% Hence, projecting the input data into a reduced space for classification
+% yields higher class separability.
+
+\paragraph{Individual effects of dimensionality reduction and rest data.}
+We first quantified the impact of introducing a bottleneck layer
+disregarding the autoencoder.
+To this end, ordinary logstic regression was juxtaposed
+with SSFLogReg  at $\lambda=1$ (no autoencoder).
+For this experiment, we increased the difficulty of the classification problem
+by including data from all 38 HCP tasks.
+Indeed, increased class separability in component space,
+as compared to voxel space, entails differences in
+generalization performance of
+$\approx{17\%}$ (Figure \ref{fig_dimred}).
+%
+Notably,
+the cognitive tasks on reward and punishment processing
+are among the least predicted with ordinary but well predicted with
+SSFLogReg
+(tasks 3 and 4 in Figure \ref{fig_dimred}).
+These experimental conditions have been reported to exhibit
+highly similar neural activity patterns in
+GLM analyses of that dataset \cite{barch2013}.
+Consequently, also local activity differences
+(in the striatum and visual cortex in this case) can
+be successfully captured by brain-network modelling.
+%
+\begin{figure}
+\begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/accuracies.pdf}
+\end{centering}
+\vspace{-0.5cm}
+\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
+Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$)
+sorted from lowest to highest.
+Multinomial logistic regression operating in voxel space (\textit{blue bars})
+was compared to 
+SSFLogReg operating in $20$
+(\textit{left plot}) and $100$ (\textit{right plot})
+latent modes (\textit{grey bars}).
+Autoencoder or rest data were not used for these analyses
+($\lambda=1$).
+Ordinary logistic regression yielded 77,7\% accuracy out of sample, while
+SSFLogReg scored at 94,4\% ($n=20$) and 94,2\% ($n=100$).
+Hence, compressing the voxel data into a component space for classification
+achieves higher task separability.
+}
+\label{fig_dimred}
+\end{figure}
+
+We then quantified the impact of structure from rest data keeping
+all model parameters constant
+($n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$).
+At the beginning of every epoch,
+2000 task and 2000 rest maps were drawn with replacement
+from same amounts of task maps but varying amounts of rest maps.
+In frequently encountered data-scarce settings
+($\approx{100}$ labeled samples),
+but not in data-rich settings ($\approx{1000}$ labeled samples),
+the repertoire of rest structure modulated model performance
+(Figure \ref{fig_semisup}).
+
+\begin{figure}
+\begin{centering}
+\includegraphics[width=1.00\textwidth]{figures/semisup_both.png}
+\end{centering}
+\vspace{-0.5cm}
+\caption{\textbf{Effect of richness in rest data}
+Gradient descent was performed on 2000 task and 2000 rest maps.
+These were drawn with replacement from identical versus varying
+quantities of task versus rest maps, at the begining of each epoch.
+Chance is at 5,6\%.
+}
+\label{fig_semisup}
+\end{figure}
+
+\paragraph{Feature identification.}
+We finally examined whether the models
+were fit for purpose
+(Figure \ref{fig_weights}).
+%
+To this end, we computed Pearson's correlation between the classifier weights
+and the averaged neural activity map for each of the 18 tasks.
+%
+Ordinary logistic regression thus yielded a mean correlation
+of $\rho=0.28$ across tasks.
+%
+For SSFLogReg ($\lambda=0.25, 0.5, 0.75, 1$),
+a per-class-weight map was computed by matrix
+multiplication of the two inner layers.
+Feature identification performance thus ranged
+between $\rho=0.35$ and $\rho=0.55$ for $n=5$,
+between $\rho=0.59$ and $\rho=0.69$ for $n=20$, and
+between $\rho=0.58$ and $\rho=0.69$ for $n=100$.
+%
+Consequently,
+SSFLogReg puts higher absolute weights on relevant structure.
+This reflects an increased signal-to-noise ratio, in part explained
+by the more BOLD-typical local contiguity.
+%
+Conversely, SSFLogReg puts lower probability mass on irrelevant structure.
+Despite lower interpretability of the results from
+ordinary logistic regression, the
+salt-and-pepper-like weight maps were sufficient for
+good classification performance.
+%
+\begin{figure}
+\begin{centering}
+\includegraphics[width=0.98\textwidth]{figures/figure_weights_perc75.png}
+\end{centering}
+\vspace{-0.1cm}
+\caption{\textbf{Classification weight maps}
+The voxel predictors corresponding to 5 exemplary
+(of 18 total) psychological tasks (\textit{rows})
+from the HCP dataset \cite{barch2013}.
+\textit{Left column:} multinomial logistic regression (same
+implementation but without bottleneck or autoencoder),
+\textit{middle column:} semi-supervised factored logistic regression
+($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
+\textit{right column:} voxel-wise average across all samples of whole-brain
+activity maps from each task.
+Semi-supervised factored logistic regression
+a) puts higher absolute weights on relevant structure,
+b) lower ones on irrelevant structure,
+and
+c) yields BOLD-typical local contiguity (without enforcing a spatial prior).
+All values are z-scored and thresholded at the $75^{th}$ percentile.
+}
+\label{fig_weights}
+\end{figure}
+%
+Hence, SSFLogReg yielded class weights 
+that were much more similar to features of the respective training samples
+for all choices of $n$ and $\lambda$.
+SSFLogReg therefore
+captures genuine properties of neural activity patterns,
+rather than participant- or study-specific artefacts.
+
+\paragraph{Miscellaneous observations.}
+For the sake of completeness,
+we informally report modifications of the statistical model
+that did not improve generalization performance.
+%
+Introducing stochasticity into model learning by 
+$a)$ input corruption of $\mathbf{X_{task}}$ or
+$b)$ drop-out at the bottleneck $\mathbf{W_0}$ using binomial
+distributions ($p=0.1, 0.3, 0.5$) deteriorated model performance
+in all scenarios.
+%
+Adding $c)$ rectified linear units (ReLU) to $\mathbf{W_0}$ or
+other commonly used nonlinearities ($d)$ sigmoid, $e)$ softplus,
+$f)$ hyperbolic tangent) all led to decreased classification accuracies,
+probably due to sample size limits.
+%
+Further, $g)$ ``pretraining'' of the bottleneck $\mathbf{W_0}$
+(i.e., non-random initialization) by 
+either corresponding SPCA or ICA loadings did not exhibit improved accuracies,
+neither did $h)$ autoencoder pretraining.
+%
+Moreover,
+introducing an additional $i)$ overcomplete layer (100 units)
+after the bottleneck was not advantageous.
+%
+Finally, imposing either $j)$ only $\ell_1$ or 
+$k)$ only $\ell_2$ penalty terms
+was disadvantageous in all tested cases,
+which favored ElasticNet regularization chosen in the above analyses.
+
+\section{Discussion and Conclusion}
+% Conceptual recap+gain of current paper
+Using the flexibility of factored models,
+we learn the low-dimensional representation from high-dimensional
+voxel brain space that is most important for
+prediction of cognitive task sets.
+%
+From a machine-learning perspective,
+factorization of the logistic regression weights
+can be viewed as transforming a
+``multi-class classification problem''
+into a ``multi-task learning problem''.
+%
+The higher generalization accuracy and feature recovery, comparing to
+ordinary logistic regression, hold potential
+for adoption in various neuroimaging analyses.
+Besides increased performance, these models are more interpretable by
+automatically learning a mapping to and from a brain-network space.
+%
+This domain-specific learning algorithm
+encourages departure from the artificial and statistically
+less attractive voxel space.
+Neurobiologically,
+brain activity underlying defined mental operations
+can be explained by linear combinations of the main activity
+patterns.
+%
+That is,
+fMRI data probably concentrate near
+a low-dimensional manifold of
+characteristic brain network combinations.
+Extracting fundamental building blocks of brain organization might
+facilitate the quest for the cognitive primitives of
+human thought.
+%
+We hope that these first steps stimulate development towards
+powerful semi-supervised representation extraction
+in systems neuroscience.
+
+% grand perspective
+In the future, automatic reduction of brain maps to
+their neurobiological essence
+may leverage data-intense neuroimaging investigations.
+Initiatives for data collection are rapidly increasing
+in neuroscience \cite{poldrack2014data}.
+These promise structured integration
+of neuroscientific knowledge accumulating
+in databases.
+Tractability by
+condensed feature representations can avoid the ill-posed problem of
+learning the full distribution of activity patterns.
+%
+This is not only relevant to the 
+multi-class challenges spanning the human cognitive space
+\cite{schwartz2013mapping}
+but also the
+multi-modal combination with
+high-resolution 3D models of brain anatomy \cite{amunts2013bigbrain}
+and
+high-throughput genomics \cite{need2010gwas}.
+%
+The biggest socioeconomic potential may
+lie in across-hospital clinical studies that
+predict disease trajectories and drug responses
+in psychiatric/neurological populations
+\cite{frackowiak2015future, gustav2011cost}.
+
+\paragraph{Acknowledgment}
+{\small
+Data were provided by the Human Connectome Project.
+% The study was supported by the German National Academic Foundation (D.B.).
+}
+
+\small
+\bibliographystyle{splncs03}
+\bibliography{nips_refs}
+
+\newpage
+\section{Appendix}
 %
@@
 from 18 diverse tasks acquired in 78 participants.
-% XXX: details on data should be given as supplementary material. the section is boring for most NIPS readers
-
-The labeled and unlabeled data were fed into a linear statistical model
-composed of an autoencoder and factored logistic regression.
-
-\paragraph{Linear autoencoder.}
-The affine autoencoder takes the input 
-$\mathbf{x}$ and projects it into coordinates of a
-latent representation $\mathbf{z}$
-and reconstructs it back to $\mathbf{x'}$ by
-\begin{eqnarray}
-    \mathbf{z} = \mathbf{W_0} \mathbf{x} + \mathbf{b_0} \hspace{2cm}
-    \mathbf{x'} = \mathbf{W_1} \mathbf{z} + \mathbf{b_1},
-  \label{eq:autoenc}
-\end{eqnarray}
-
-where $\mathbf{x \in \mathbb{R}^{d}}$ denotes the vector of $d=79{,}941$
-voxel values from each
-rest map,
-$\mathbf{z \in \mathbb{R}^{n}}$ is the $n$-dimensional hidden state
-(i.e., distributed neural activity patterns), and 
-$\mathbf{x' \in \mathbb{R}^{d}}$
-is the reconstruction vector of the original activity map
-from the hidden variables. 
-% XXX use bold letters for vectors and matrices
-Further, $\mathbf{W_0}$ denotes the weight matrix that
-transforms
-from input space into the hidden space (encoder),
-$\mathbf{W_1}$ is the weight matrix for back-projection
-from the hidden variables to the
-output space (decoder).
-$\mathbf{b_0}$ and $\mathbf{b_1}$ are bias vectors.
-The model parameters $\mathbf{W_0, b_0, b_1}$ are found by
-minimizing the expected squared reconstruction error
-\begin{equation}
-  \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
-  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
-\end{equation}
-
-%with \(f_{AE}(\mathbf x) = %
-%\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}\)
-% the autoencoder.
-% XXX Michael: I don't know to what extent the following makes sense or is
-% XXX correct, so I am commenting it
-% XXX Danilo: I found this in Bengio's papers somewhere
-This reconstruction error criterion equates with
-maximizing a lower bound on the mutual information between
-input and the learned representation.
-% XXX Give a ref ?
-Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
-$\mathbf{W_0} = \mathbf{W_1^T}$. In particular, this means that the weights
-learned are forced to take a two-fold function: That of signal 
-\textit{analysis} and that of signal \textit{synthesis}.
-The first layer \textit{analyzes}
-the data in order to obtain the cleanest latent representation,
-while the second
-layer represents building blocks from which to \textit{synthesize} the data
-using the latent activations.
-Tying these processes together makes the analysis
-layer interpretable and pulls all non-zero singular values towards 1. 
-Nonlinearities were not applied to the
-activations in the first layer.
-
-\paragraph{Factored logistic regression.}
-Lossy compression by a low-dimensional bottleneck
-is also imposed by the first layer of the factored
-multinomial logistic regression.
-It gives the probability of an input $\mathbf{x}$ to belong
-to a class $i \in \{1, \dots, l\}$
-\begin{equation}
-  \begin{split}
-    P(Y=i|\mathbf{x; V_0,V_1,c_0, c_1}) &= \softmax_i(f_\mathcal{LR}(\mathbf x)),
-  \end{split}
-  \label{eq:lr}
-\end{equation}
-where \(f_\mathcal{LR}(\mathbf x) = \mathbf{V_1 (V_0 x + c_0) + c_1}\) computes 
-multinomial logits and \(\softmax_i(x) = \exp(x_i)/\sum_j\exp(x_j)\).
-The matrix $\mathbf{V_0 \in \mathbb{R}^{dxn}}$
-transforms the input $\mathbf{x \in \mathbb{R}^{d}}$
-into $n$ latent components
-and the matrix $\mathbf{V_1 \in \mathbb{R}^{nxl}}$
-projects the latent components
-onto hyperplanes that reflect $l$ label probabilities.
-$\mathbf{c_0}$ and $\mathbf{c_1}$ are corresponding
-bias vectors.
-The loss function is given by
-\begin{equation}
-    \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
-\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1}).
-\label{eq:lr_loss}
-\end{equation}
-
-\paragraph{Layer combination.}
-Importantly, the optimization problem of the linear autoencoder
-and the factored logistic regression
-are linked on two levels. First, their transformation matrices mapping from
-input to the latent space are tied
-\begin{eqnarray}
-  \mathbf{V_0} = \mathbf{W_0}.
-\end{eqnarray}
-We thus search for a compression of the 79,941 voxel values into $n$ latent
-components that represent a latent code optimized for both
-rest and task activity data.
-Second, the objectives of the autoencoder and the factored
-logistic regression are interpolated in the common loss function
-% XXX Michael: In resolving conflict I opted for my change. Feel free to revert if that was intended.
-\begin{equation}
-{\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
-+ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega(\theta).
-  \label{eq:loss_equ}
-\end{equation}
-
-In so doing, we search for the combined model parameters
-$\theta=\{\mathbf{V_0,V_1,c_0, c_1, b_0, b_1}\}$
-with respect to the (unsupervised) reconstruction error and the
-(supervised) task classification.
-${\mathcal{L_{AE}}}$ is devided by ${N_{X_{rest}}}$ to equilibrate both
-loss terms to the same order of magnitude.
-\(\Omega(\theta)\) represents a regularization, for which we choose the
-ElasticNet-type, i.e., a combination of $\ell_1$ and $\ell_2$ penalties.
-%$\forall p \in \theta$.
-
-% XXX Michael: To sum up the math in this paragraph: We need to decide what
-% XXX the \mathcal L are supposed to mean - sample wise loss or expected loss
-% XXX or some global deterministic loss (the latter would be bad, because we)
-% XXX are not optimizing it. Right now it looks like a sample-wise loss.
-
-\paragraph{Optimization.}
-The common objective was optimized
-by gradient descent in the SS-FLogReg parameters.
-% XXX you haven't defined FLogReg
-%
-The required gradients were obtained by using the chain rule to
-backpropagate error derivatives.
-We chose the \textit{rmsprop} solver \cite{rmsprop},
-a refinement of stochastic gradient descent.
-\textit{Rmsprop} dictates an adaptive learning rate
-for each model parameter by
-scaled gradients from a running average.
-The batch size was set to $100$
-(much expected redundancy in $X_{rest}$ and $X_{task}$),
-matrix parameters were initalized by Gaussian random values multiplied
-by $0.004$ (i.e., gain), and
-bias parameters were initalized to zero.
-%
-The normalization factor and the update rule for $\theta$
-are given by
-% XXX Michael: It is crucial, but nowhere mentioned, that the optimization is
-% XXX done in an alternating manner. The whole structure relies on this, by
-% XXX construction it is vulnerable to optimizing too much towards labels or
-% XXX rest at a time. I also identify this as one of the major weak points
-% XXX of the method, so it needs to be stated in a way that is not easily
-% XXX attackable.
-% XXX Maybe we could also talk less about RMSprop.
-%
-% XXX Danilo: I think we are not alternating, Xrest and Xtask go in
-% simultaneously at every bach in every epoch
-\begin{eqnarray}
-  \begin{split}
-    \mathbf{v^{(t+1)}} &= \rho \mathbf{v^{(t)}} + (1 - \rho)\left(\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})\right)^2
-% XXX: I have changed the first equation, please check
-%v^{(t+1)} = \rho v^{(t+1)} + (1 - \rho)˜|\nabla f
-\\
-\theta^{(t+1)} &= \theta^{(t)} + \alpha \frac{\nabla_{\theta} f(x^{(t)}, y^{(t)}, \theta^{(t)})}{\sqrt{\mathbf{v^{(t+1)}} + \epsilon}},
-  \end{split}
-\end{eqnarray}
-% XXX Michael: Why, in the formula, do we have two different notations for
-% XXX partial derivatives? \partial and \nabla
-% XXX Michael: Probably remove the explanations following and just state values
-% XXX: Bertrand: what is f ?? Danilo: replaced f by more detailed notation
-where $f$ is the loss function computed on a minibatch sample at timestep $t$,
-$\alpha$ is the learning rate ($0.00001$),
-$\epsilon$ a global damping factor ($10^{-6}$)
-and
-$\rho$ the decay rate ($0.9$ to deemphasize the magnitude of the gradient).
-%
-Note that we have also experimented with other solvers
-(stochastic gradient descent, adadelta, and adagrad) but found that
-\textit{rmsprop} converged faster and with
-similar or higher generalization performance.
-
-\paragraph{Hints.}
-In fact, the constraint by a rest-data autoencoder qualifies as a
-\textit{hint}
-rather than regularization in a strict sense \cite{abu1994hints}.
-Its purpose is to introduce
-% XXX Michael: I use f earlier up, is this f here necessary?
-prior knowledge on
-\textit{known} properties of the \textit{unknown} target function $f$.
-Rather than only relying on input-output pairs in the learning process,
-we thus narrow our hypothesis set to the biologically most plausible solutions.
-That is, we reduce the search space in a way that
-is compatible with the expected representation of BOLD activity signals.
-
-\paragraph{Implementation.}
-The analyses were performed in Python.
-We used \textit{nilearn} to handle
-the large quantities of neuroimaging data 
-\cite{abrah14}
-and
-\textit{Theano} for automatic, numerically stable
-differentiation of symbolic computation graphs
-\cite{bastien2012theano, bergstra2010theano}.
-All Python scripts that generated the results are
-accessible online for reproducibility and reuse
-\url{http://github.com/anonymous/anonymous}.
-% url{http://github.com/banilo/nips2015}.
-
-\section{Experimental Results}
-\paragraph{Serial versus parallel structure discovery and classification.}
-We first tested whether there is a substantial advantage in combining unsupervised decomposition
-and supervised classification learning.
-We benchmarked our approach against
-performing data reduction on the (unlabeled)
-first half of the HCP task data by PCA, SPCA, ICA, and
-AE ($n=5, 20, 50, 100$ components)
-and learning classification models in the (labeled) second half
-by ordinary logistic regression.
-%
-PCA reduced the dimensionality of the task data by
-finding orthogonal network components using a change of basis
-(whitening).
-%
-SPCA separated the BOLD signals into
-network components with few regions by
-a regression-type optimization problem constrained by
-$\ell_1$-penalty
-(no orthogonality assumptions, 1000 maximum iterations,
-per-iteration tolerance of 1 * 10\textsuperscript{-8},
-$\alpha=1$).
-%
-ICA performed iterative blind source separation
-by a parallel FASTICA implementation (200 maximum iterations,
-per-iteration tolerance of 0.0001,
-initialized by random mixing matrix, whitening).
-%
-AE found a code of latent representations by optimizing projection
-into a bottleneck
-(500 iterations, same implementation as below for rest data).
-%
-The second half of the task data was projected onto the
-latent components discovered in its first half.
-Only the ensuing component loadings were submitted to ordinary
-logistic regression
-(one hidden layer, $\ell_1=0.1$, $\ell_2=0.1$, 500 maximum iterations).
-%
-These serial two-step approaches were
-compared against parallel decomposition and classification by FLogReg
-(two hidden layers, $\lambda=1.00$, $\ell_1=0.1$, $\ell_2=0.1$,
-500 iterations).
-Importantly, all trained classification models were tested
-on a large, unseen test set (20\% of data) in the present analyses.
-%
-Across choices for $n$, FLogReg
-achieved more than 95\% out-of-sample accuracy, whereas
-supervised learning based on PCA, SPCA, ICA, and AE loadings
-ranged from 32\% to 87\%
-(Table \ref{table_one}).
-%
-This experiment establishes the advantage of directly searching for
-classification-relevant structure in the fMRI data,
-rather than solving the supervised and unsupervised problems independently.
-This effect was particularly pronounced when assuming few hidden dimensions.
-\begin{table}[h]
- \centering
- \resizebox{0.8\textwidth}{!}{%
- \begin{tabular}{l|c|c|c|c|c}
- \hline
- \textit{n}               & PCA + LogReg   & SPCA + LogReg  & ICA + LogReg  & AE + LogReg    & FLogReg              \\ \hline
- \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & \textbf{95,7\%}      \\
- \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & \textbf{97,3\%}      \\
- \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & \textbf{97,6\%}      \\
- \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & \textbf{97,4\%}      \\ \hline
- \end{tabular}
- }
- \vspace{-0.2cm}
- \caption{\textbf{Serial versus parallel dimensionality reduction and classification.}
- Chance is at 5,6\%.
- }
- \label{table_one}
- \end{table}
-\paragraph{Model performance.}
-SS-FLogReg was subsequently trained ($500$ epochs) across parameter choices
-for the hidden components ($n=5, 20, 100$) and
-the balance between autoencoder and logistic regression
-($\lambda=0, 0.25, 0.5, 0.75, 1$).
-Assuming 5 latent directions of variation should yield models with
-higher bias and smaller variance then SS-FLogReg with 100 latent directions.
-%
-Given the 18-class problem of HCP, setting $\lambda$ to $0$
-consistently yields generalization performance
-at chance-level (5,6\%) because
-only the unsupervised layer of the estimator is optimized.
-%
-At each epoch (i.e., iteration over the data),
-the out-of-sample performance of the trained classifier
-was assessed on 20\% of unseen HCP data.
-Additionally, the ``out-of-study performance'' of the learned decomposition
-was assessed by using it as dimensionality reduction of an
-independent labeled dataset (i.e., ARCHI) and conducting ordinary
-logistic regression on the ensuing component loadings.
-
-\begin{table}[h]
-  \centering
-  \resizebox{1.00\textwidth}{!}{%
-\tabcolsep=0.05cm
-\begin{tabular}{l|ccccc|ccccc|ccccc}
-  \hline
-                            & \multicolumn{5}{c|}{\textit{n = 5}}                                                                                                                                                 & \multicolumn{5}{c|}{\textit{n = 20}}                                                                                                                                                      & \multicolumn{5}{c}{\textit{n = 100}}                                                                           \\ \hline
-  \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
-  Out-of-sample\\accuracy   & \textit{6,0\%}                     & 88,9\%                             & 95,1\%                            & \textbf{96,5\%}                    & 95,7\%                           & \textit{5,5\%}                   & 97,4\%                               & \textbf{97,8\%}                     & 97,3\%                               & 97,3\%                             & \textit{6,1\%}   & 97,2\%           & 97,0\%           & \textbf{97,8\%}  & 97,4\%                             \\
-  Precision (mean)          & \textit{5,9\%}                     & 87,0\%                             & 94,9\%                            & \textbf{96,3\%}                    & 95,4\%                           & \textit{5,1\%}                   & \textbf{97,4\%}                      & 97,1\%                              & 97,0\%                               & 97,0\%                             & \textit{5,9\%}   & 96,9\%           & 96,5\%           & \textbf{97,5\%}  & 96,9\%                             \\
-  Recall (mean)             & \textit{5,6\%}                     & 88,3\%                             & 95,2\%                            & \textbf{96,6\%}                    & 95,7\%                           & \textit{4,6\%}                   & \textbf{97,5\%}                      & 97,5\%                              & 97,4\%                               & 97,4\%                             & \textit{7,2\%}   & 97,2\%           & 97,2\%           & \textbf{97,9\%}  & 97,4\%                             \\
-  F1 score (mean)           & \textit{4,1\%}                     & 86,6\%                             & 94,9\%                            & \textbf{96,4\%}                    & 95,4\%                           & \textit{3,8\%}                   & \textbf{97,4\%}                      & 97,2\%                              & 97,1\%                               & 97,1\%                             & \textit{5,3\%}   & 97,0\%           & 96,7\%           & \textbf{97,7\%}  & 97,2\%                             \\ \hline
-  Out-of-study\\accuracy    & \textit{39,4\%}                    & 60,8\%                             & 54,3\%                            & 60,7\%                             & \textbf{62,9\%}                  & \textit{77,0\%}                  & 79,7\%                               & \textbf{81,9\%}                     & 79,7\%                               & 79,4\%                             & 79,2\%           & \textbf{82,2\%}  & 81,7\%           & 81,3\%           & \textit{75,8\%}                    \\ \hline
-  \end{tabular}
-}
-\vspace{-0.2cm}
- \caption{\textbf{Performance of SS-FLogReg across model parameter choices.}
- Chance is at 5,6\%.
- }
-  \label{table_two}
-\end{table}
-
-We made several noteworthy observations (Table \ref{table_two}).
-%
-First, the purely supervised estimator ($\lambda=1$) achieved in no
-instance the best accuracy, precision, recall or f1 scores on HCP data.
-Classification by factored logistic regression is therefore facilitated by
-imposing structure from the unlabeled rest data.
-%
-Second, the higher the number of latent components $n$,
-the higher the out-of-study performance with small values of $\lambda$.
-This suggests that the presence of more rest-data-inspired hidden components
-results in more effective feature representations in unrelated task data.
-%
-Third, for $n=20$ and $100$ (but not $5$) the purely rest-data-trained
-decomposition matrix ($\lambda=0$) resulted in
-noninferior out-of-study performance
-of 77,0\% and 79,2\%,
-respectively (Table \ref{table_two}).
-This confirms that guiding model learning by task-unrelated structure
-extracts features of general relevance
-beyond the supervised problem at hand.
-
-% Depicts f1 scores for prediction of each of 38 classes across training epochs.
-% Ordinary logistic regression operating in voxel space (\textit{left plot})
-% converged faster but performed worse than
-% factored logistic regression for few
-% (\textit{middle plot}) and many (\textit{right plot}).
-% Autoencoder or rest data were not used for these analyses
-% ($\lambda=1$).
-% Hence, projecting the input data into a reduced space for classification
-% yields higher class separability.
-
-\paragraph{Individual effects of dimensionality reduction and rest data.}
-We first quantified the impact of introducing a bottleneck layer
-that disregards the autoencoder.
-To this end, ordinary logstic regression was juxtaposed
-with FLogReg  at $\lambda=1$ (no autoencoder).
-For this experiment, we increased the difficulty of the classification problem
-by including data from all 38 HCP tasks.
-Indeed, increased class separability in component space,
-as compared to voxel space, entails differences in
-generalization performance of
-$\approx{17\%}$ (Figure \ref{fig_dimred}).
-%
-Notably,
-the cognitive tasks on reward and punishment processing
-are among the least predicted with ordinary but well predicted with
-factored logistic regression
-(tasks 3 and 4 in Figure \ref{fig_dimred}).
-These experimental conditions have indeed been reported to exhibit
-highly similar neural activity patterns in
-GLM analyses of that dataset \cite{barch2013}.
-Consequently, also local activity differences
-(in the striatum and visual cortex in this case) can
-be successfully captured by brain-network modelling.
-%
-\begin{figure}
-\begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/accuracies.pdf}
-\end{centering}
-\vspace{-0.5cm}
-\caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
-Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$)
-sorted from lowest to highest.
-Multinomial logistic regression operating in voxel space (\textit{blue bars})
-was compared to 
-factored logistic regression operating in $20$
-(\textit{left plot}) and $100$ (\textit{right plot})
-latent modes (\textit{grey bars}).
-Autoencoder or rest data were not used for these analyses
-($\lambda=1$).
-Ordinary logistic regression yielded 77,7\% accuracy out of sample, while
-factored logistic regreesion scored at 94,4\% ($n=20$) and 94,2\% ($n=100$).
-Hence, compressing the voxel data into a component space for classification
-achieves higher task separability.
-}
-\label{fig_dimred}
-\end{figure}
-
-We then quantified the impact of structure from rest data keeping
-all model parameters constant
-($n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$).
-At the beginning of every epoch,
-2000 task and 2000 rest maps were drawn with replacement
-from same amounts of task maps but varying amounts of rest maps.
-In frequently encountered data-scarce settings ($\approx{100}$ samples),
-but not in data-rich settings ($\approx{1000}$ samples),
-the repertoire of rest structure modulated model performance
-(Figure \ref{fig_semisup}).
-
-\begin{figure}
-\begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/semisup_both.png}
-\end{centering}
-\vspace{-0.5cm}
-\caption{\textbf{Effect of richness in rest data}
-Gradient descent was performed on 2000 task and 2000 rest maps.
-These were drawn with replacement from identical versus varying
-quantities of task versus rest maps, at the begining of each epoch.
-Chance is at 5,6\%.
-}
-\label{fig_semisup}
-\end{figure}
-
-\paragraph{Feature identification.}
-We finally examined whether the models
-were fit for purpose
-(Figure \ref{fig_weights}).
-%
-To this end, we computed Pearson's correlation between the classifier weights
-and the averaged neural activity map for each of the 18 tasks.
-%
-Ordinary logistic regression thus yielded a mean correlation
-of $\rho=0.28$.
-%
-For SS-FLogReg ($\lambda=0.25, 0.5, 0.75, 1$),
-a per-class-weight map was computed by matrix
-multiplication of the two inner layers.
-Feature identification thus ranged
-between $\rho=0.35$ and $\rho=0.55$ for $n=5$,
-between $\rho=0.59$ and $\rho=0.69$ for $n=20$, and
-between $\rho=0.58$ and $\rho=0.69$ for $n=100$.
-%
-Consequently,
-SS-FLogReg puts higher absolute weights on relevant structure.
-This reflects a higher signal-to-noise ratio, in part explained
-by the more BOLD-typical local contiguity.
-%
-Conversely, SS-FLogReg puts lower probability mass on irrelevant structure.
-Despite lower interpretability, the
-salt-and-pepper-like weight maps obtained from
-ordinary logistic regression are sufficient for
-good classification performance.
-%
-\begin{figure}
-\begin{centering}
-\includegraphics[width=0.98\textwidth]{figures/figure_weights_perc75.png}
-\end{centering}
-\vspace{-0.1cm}
-\caption{\textbf{Classification weight maps}
-The voxel predictors corresponding to 5 exemplary
-(of 18 total) psychological tasks (\textit{rows})
-from the HCP dataset \cite{barch2013}.
-\textit{Left column:} multinomial logistic regression (same
-implementation but without bottleneck or autoencoder),
-\textit{middle column:} semi-supervised factored logistic regression
-($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
-\textit{right column:} voxel-wise average of whole-brain
-activity maps from each task.
-Semi-supervised factored logistic regression
-a) puts higher absolute weights on relevant structure,
-b) lower ones on irrelevant structure,
-and
-c) yields BOLD-typical local contiguity (without enforcing a spatial prior).
-All values are z-scored and thresholded at the $75^{th}$ percentile.
-}
-\label{fig_weights}
-\end{figure}
-%
-Hence, SS-FLogReg yielded class weights 
-that were much more similar to features of the respective training samples
-for any choice of $\lambda$.
-SS-FLogReg therefore
-captures genuine properties of neural activity patterns,
-rather than participant- or study-specific artefacts.
-
-\paragraph{Miscellaneous observations.}
-For the sake of completeness,
-we informally report modifications of the model
-that did not improve the generalization performance.
-%
-Introducing stochasticity into model learning by 
-a) input corruption of $\mathbf{X_{task}}$ or
-b) drop-out at the bottleneck $\mathbf{W_0}$ using binomial
-distributions ($p=0.1, 0.3, 0.5$) deteriorated model performance
-in all instances. 
-%
-Adding c) rectified linear units (ReLU) to $\mathbf{W_0}$ or
-other commonly used nonlinearities (d) sigmoid, e) softplus,
-f) hyperbolic tangent) all led to smaller classification accuracies,
-probably due to sample size limits.
-%
-Further, g) ``pretraining'' of the bottleneck $\mathbf{W_0}$
-(i.e., non-random initialization) by 
-either corresponding SPCA or ICA loadings did not exhibit improved accuracies,
-neither did h) autoencoder pretraining.
-%
-Moreover,
-introducing an additional i) overcomplete layer (100 units)
-after the bottleneck was not advantageous.
-%
-Finally, imposing either j) only $\ell_1$-penalty or 
-k) only $\ell_2$-penalty was disadvantageous in all tested cases,
-which favored ElasticNet regularization chosen in the above analyses.
-
-\section{Discussion and Conclusion}
-% Conceptual recap+gain of current paper
-Using the flexibility of factored models,
-we learn the low-dimensional representation from high-dimensional
-voxel brain space that is most important for
-prediction of cognitive task sets.
-%
-From a machine-learning perspective,
-factorization of the logistic regression weights
-can be viewed as transforming a
-``multi-class classification problem''
-into a ``multi-task learning problem''.
-%
-The higher generalization accuracy and feature recovery, comparing to
-ordinary logistic regression, hold potential
-for adoption in various neuroimaging analyses.
-Besides increased performance, these models are more interpretable by
-automatically learning a mapping to and from a brain-network space.
-%
-This domain-specific learning algorithm
-encourages departure from the artificial and statistically
-less attractive voxel space.
-Neurobiologically,
-brain activity underlying defined mental operations
-can be explained by linear combinations of the main activity
-patterns.
-%
-That is,
-fMRI data probably concentrate near
-a low-dimensional manifold of
-characteristic brain networks combinations.
-Extracting fundamental building blocks of brain organization might
-facilitate the quest for the cognitive primitives of
-human thought.
-%
-We hope that these first steps stimulate development towards
-powerful semi-supervised representation extraction
-in systems neuroscience.
-
-% grand perspective
-In the future, automatic reduction of brain maps to
-their neurobiological essence
-may leverage data-intense neuroimaging studies.
-Initiatives for data collection are rapidly increasing
-in neuroscience \cite{poldrack2014data}.
-These promise structured integration
-of neuroscientific knowledge accumulating
-in databases.
-Tractability by
-condensed feature representations can avoid the ill-posed problem of
-learning the full distribution of activity patterns.
-%
-This is not only relevant to the 
-multi-class challenges spanning the human cognitive space
-\cite{schwartz2013mapping}
-but also the
-multi-modal combination with
-high-resolution 3D models of brain anatomy \cite{amunts2013bigbrain}
-and
-high-throughput genome analyses \cite{need2010gwas}.
-%
-The biggest socioeconomic potential may
-lie in across-hospital clinical studies that
-predict disease trajectories and drug responses
-in psychiatric/neurological populations
-\cite{frackowiak2015future, gustav2011cost}.
-
-\paragraph{Acknowledgment}
-{\small
-Data were provided by the Human Connectome Project.
-% The study was supported by the German National Academic Foundation (D.B.).
-}
-
-\small
-\bibliographystyle{splncs03}
-\bibliography{nips_refs}
 

### 2015-06-05 paper_nips/paper.tex
removed non-centered from PCA-autoencoder description
grammar

@@
 % DanilO: cool thanks. That is the stuff, we need.
-They behave like a (truncated, non-centered) PCA
+They behave like a (truncated) PCA
 in case of one linear hidden layer and a squared error loss
@@
 %
-Practically, neuroimaging research is frequently suffering from
+Practically, neuroimaging research frequently suffers from
 data scarcity. This limits the set of representations that can be

### 2015-06-05 paper_nips/paper.tex
typo

@@
   9 Tongue movement    & \multirow{3}{*}{Visual cues}          & Move tongue                                                                       \\ \cline{1-1} \cline{3-3} 
-  10 Food movement     &                                       & Squeezing of the left of right toe                                                \\ \cline{1-1} \cline{3-3} 
+  10 Food movement     &                                       & Squeezing of the left or right toe                                                \\ \cline{1-1} \cline{3-3} 
   11 Hand movement     &                                       & Tapping of the left or right finger                                               \\ \hline

### 2015-06-05 paper_nips/paper.tex
minor
standard deviation

@@
 are linked on two levels. First, their transformation matrices mapping from
-input to the latent space are tied
+input to the latent space are tied, i.e.
 \begin{eqnarray}
@@
 (much expected redundancy in $X_{rest}$ and $X_{task}$),
-matrix parameters were initalized by Gaussian random values multiplied
-by $0.004$ (i.e., gain), and
+matrix parameters were initalized by Gaussian random values with
+standard deviation $0.004$ (i.e., gain), and
 bias parameters were initalized to $0$.

### 2015-06-05 paper_nips/paper.tex
remove 1 *
typo
remove dropout

@@
 (no orthogonality assumptions, 1000 maximum iterations,
-per-iteration tolerance of 1 * 10\textsuperscript{-8},
+per-iteration tolerance of 10\textsuperscript{-8},
 $\alpha=1$).
@@
 Assuming 5 latent directions of variation should yield models with
-higher bias and smaller variance then SSFLogReg with 100 latent directions.
+higher bias and smaller variance than SSFLogReg with 100 latent directions.
 %
@@
 %
+$a)$
 Introducing stochasticity into model learning by 
-$a)$ input corruption of $\mathbf{X_{task}}$ or
-$b)$ drop-out at the bottleneck $\mathbf{W_0}$ using binomial
-distributions ($p=0.1, 0.3, 0.5$) deteriorated model performance
+%$a)$ 
+input corruption of $\mathbf{X_{task}}$ %or
+%% $b)$ drop-out at the bottleneck $\mathbf{W_0}$ using binomial
+%% distributions ($p=0.1, 0.3, 0.5$) 
+deteriorated model performance
 in all scenarios.
 %
-Adding $c)$ rectified linear units (ReLU) to $\mathbf{W_0}$ or
-other commonly used nonlinearities ($d)$ sigmoid, $e)$ softplus,
-$f)$ hyperbolic tangent) all led to decreased classification accuracies,
+Adding $b)$ rectified linear units (ReLU) to $\mathbf{W_0}$ or
+other commonly used nonlinearities ($c)$ sigmoid, $d)$ softplus,
+$e)$ hyperbolic tangent) all led to decreased classification accuracies,
 probably due to sample size limits.
 %
-Further, $g)$ ``pretraining'' of the bottleneck $\mathbf{W_0}$
+Further, $f)$ ``pretraining'' of the bottleneck $\mathbf{W_0}$
 (i.e., non-random initialization) by 
 either corresponding SPCA or ICA loadings did not exhibit improved accuracies,
-neither did $h)$ autoencoder pretraining.
+neither did $g)$ autoencoder pretraining.
 %
 Moreover,
-introducing an additional $i)$ overcomplete layer (100 units)
+introducing an additional $h)$ overcomplete layer (100 units)
 after the bottleneck was not advantageous.
 %
-Finally, imposing either $j)$ only $\ell_1$ or 
+Finally, imposing either $i)$ only $\ell_1$ or 
 $k)$ only $\ell_2$ penalty terms

### 2015-06-05 paper_nips/paper.tex
included OGs comments

@@
   \vspace{-0.7cm}
-  \caption {\textbf{Model architecture.}
+  \caption {\textbf{Model architecture}
   Linear autoencoders find an optimized compression of 79,941 brain voxels
@@
 (labeled) task data.
-We can then solve the unsupervised data representation and the
-supervised classification, previously studied in isolation,
-in an identical statistical learning problem.
+We can then solve the compound statistical problem of
+unsupervised data representation and
+supervised classification, previously studied in isolation.
 %
@@
 probability).
-The supervised analyses were based on labeled HCP task maps with
+The supervised analyses were thus based on labeled HCP task maps with
 79,941 voxels of interest representing z-values in gray matter.
@@
   4 Faces              & Face pictures                         & Decide which of two faces matches another face emotionally                        \\ \hline
-  5 Random             & \multirow{2}{*}{Videos with objects}  & \multirow{2}{*}{Decide whether the circles, squares, triangles act intentionally} \\ \cline{1-1}
+  5 Random             & \multirow{2}{*}{Videos with objects}  & \multirow{2}{*}{Decide whether the objects act randomly or intentionally} \\ \cline{1-1}
   6 Theory of mind     &                                       &                                                                                   \\ \hline
@@
 reading, language comprehension and mental calculation.
-Analogous to HCP data, the second task dataset incorporated 1404
+Analogous to HCP data, the second task dataset thus incorporated 1404
 labeled, grey-matter masked, and z-scored activity maps
@@
 
+\paragraph{Linear autoencoder.}
 The labeled and unlabeled data were fed into a linear statistical model
 composed of an autoencoder and dimensionality-reducing logistic regression.
-
-\paragraph{Linear autoencoder.}
 The affine autoencoder takes the input 
-$\mathbf{x}$ and projects it into a coordinate system of
+$\mathbf{x}$, projects it into a coordinate system of
 latent representations $\mathbf{z}$
@@
 % XXX Danilo: I found this in Bengio's papers somewhere
-This reconstruction error criterion equates with
-maximizing a lower bound on the mutual information between
-input and the learned representation.
+% This reconstruction error criterion equates with
+% maximizing a lower bound on the mutual information between
+% input and the learned representation.
 % XXX Give a ref ?
 Here we choose $\mathbf{W_0}$ and $\mathbf{W_1}$ to be tied, i.e.
-$\mathbf{W_0} = \mathbf{W_1^T}$. In particular, this means that the weights
-learned are forced to take a two-fold function: That of signal 
+$\mathbf{W_0} = \mathbf{W_1^T}$.
+Consequently, the learned weights
+ are forced to take a two-fold function: That of signal
 \textit{analysis} and that of signal \textit{synthesis}.
@@
 \paragraph{Factored logistic regression.}
-Lossy compression by a low-dimensional bottleneck
-is also imposed by the first layer of the factored
-multinomial logistic regression.
-It gives the probability of an input $\mathbf{x}$ to belong
-to a class $i \in \{1, \dots, l\}$
+Our factored logistic regression model is best described as a
+variant of a multinomial logistic regression.
+Specifically, the weight
+matrix is replaced by the product of two weight matrices with a common
+latent dimension.
+The later is typically much lower than the dimension of the data.
+%
+Alternatively, this model can be viewed as a single-hidden-layer
+feed-forward neural network with a linear activation function for the
+hidden layer and a softmax function on the output layer.
+As the
+dimension of the hidden layer is much lower than the input layer, this
+architecture is sometimes referred to as a ``linear bottleneck'' in the
+literature.
+%
+The probability of an input $\mathbf{x}$ to belong
+to a class $i \in \{1, \dots, l\}$ is given by
 \begin{equation}
@@
 and the factored logistic regression
-are linked on two levels. First, their transformation matrices mapping from
-input to the latent space are tied, i.e.
+are linked in two ways. First, their transformation matrices mapping from
+input to the latent space are tied
 \begin{eqnarray}
@@
 \end{eqnarray}
-We thus search for a compression of the 79,941 voxel values into $n$ latent
+We hence search for a compression of the 79,941 voxel values into $n$ unknown
 components that represent a latent code optimized for both
@@
 The batch size was set to $100$
-(much expected redundancy in $X_{rest}$ and $X_{task}$),
-matrix parameters were initalized by Gaussian random values with
-standard deviation $0.004$ (i.e., gain), and
+(given much expected redundancy in $X_{rest}$ and $X_{task}$),
+matrix parameters were initalized by Gaussian random values multiplied
+by $0.004$ (i.e., gain), and
 bias parameters were initalized to $0$.
@@
 $\alpha$ is the learning rate ($0.00001$),
-$\epsilon$ a global damping factor ($10^{-6}$)
+$\epsilon$ a global damping factor ($10^{-6}$),
 and
@@
 similar or higher generalization performance.
-
-\paragraph{Hints.}
-In fact, the constraint by a rest-data autoencoder qualifies as a
-\textit{hint}
-rather than regularization in a strict sense \cite{abu1994hints}.
-Its purpose is to introduce
-% XXX Michael: I use f earlier up, is this f here necessary?
-prior information on
-\textit{known} properties of the \textit{unknown} target function $f$.
-Rather than only relying on input-output pairs in the learning process,
-we thus narrow our hypothesis set to the biologically most plausible solutions.
-That is, we reduce the search space in a way that
-is compatible with the expected representation of BOLD activity signals.
+% \paragraph{Hints.}
+% In fact, the constraint by a rest-data autoencoder qualifies as a
+% \textit{hint}
+% rather than regularization in a strict sense \cite{abu1994hints}.
+% Its purpose is to introduce
+% % XXX Michael: I use f earlier up, is this f here necessary?
+% prior information on
+% \textit{known} properties of the \textit{unknown} target function $f$.
+% Rather than only relying on input-output pairs in the learning process,
+% we thus narrow our hypothesis set to the biologically most plausible solutions.
+% That is, we reduce the search space in a way that
+% is compatible with the expected representation of BOLD activity signals.
 
@@
 accessible online for reproducibility and reuse
-\url{http://github.com/anonymous/anonymous}.
-% url{http://github.com/banilo/nips2015}.
+(\url{http://github.com/anonymous/anonymous}).
+% (url{http://github.com/banilo/nips2015}).
 
@@
 %
-SPCA separated the BOLD signals into
+SPCA separated the task-related BOLD signals into
 network components with few regions by
@@
 
-We made several noteworthy observations (Table \ref{table_two}).
+We made three noteworthy observations (Table \ref{table_two}).
 %
 First, the most supervised estimator ($\lambda=1$) achieved in no
-instance the best accuracy, precision, recall or f1 scores on HCP data.
+instance the best accuracy, precision, recall, or f1 scores on HCP data.
 Classification by SSFLogReg is therefore facilitated by
@@
 To this end, ordinary logstic regression was juxtaposed
-with SSFLogReg  at $\lambda=1$ (no autoencoder).
+with SSFLogReg with $\lambda=1$.
 For this experiment, we increased the difficulty of the classification problem
@@
 \caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
-Depicts the f1 score for each of 38 class predictions (chance is  $2,6\%$)
+Depicts the f1 score for each of 38 class predictions (chance is at $2,6\%$)
 sorted from lowest to highest.
@@
 
-We then quantified the impact of structure from rest data keeping
-all model parameters constant
-($n=20$, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$).
+% rest effect
+We then contemplated the impact of rest structure by modulating
+its influence ($\lambda=0.25, 0.5, 0.75$) in data-scarce
+and data-rich settings ($n=20$, $\ell_1=0.1$, $\ell_2=0.1$).
 At the beginning of every epoch,
 2000 task and 2000 rest maps were drawn with replacement
-from same amounts of task maps but varying amounts of rest maps.
-In frequently encountered data-scarce settings
-($\approx{100}$ labeled samples),
+from same amounts of task and rest maps.
+In frequently encountered data-scarce scenarios,
+both training and validation
+scores improved as we depart from the most supervised model ($lambda=1$).
+This suggests that the autoencoder of SSFLogReg does not
+behave like a classical regularization but more
+like transfer learning.
+In data-rich scenarios, was also present, although less pronounced.
+
 but not in data-rich settings ($\approx{1000}$ labeled samples),
@@
 
-\begin{figure}
+\begin{figure}[h]
 \begin{centering}
-\includegraphics[width=1.00\textwidth]{figures/semisup_both.png}
-\end{centering}
+\includegraphics[width=0.55\textwidth]{figures/semisup.png}
 \vspace{-0.5cm}
-\caption{\textbf{Effect of richness in rest data}
+\caption{\textbf{Effect of rest structure}
+Model performance of SSFLogReg ($n=20$, $\ell_1=0.1$, $\ell_2=0.1$)
+for different choices of $\lambda$ in
+data-scarce (100 task and 100 rest maps, \textit{hot color}) and
+data-rich (1000 task and 1000 task maps, \textit{cold color}) scenarios.
 Gradient descent was performed on 2000 task and 2000 rest maps.
-These were drawn with replacement from identical versus varying
-quantities of task versus rest maps, at the begining of each epoch.
+At the begining of each epoch, these were drawn with replacement
+from a pool of 100 or 1000 different task and rest maps, respectively.
+Especially the \textit{cold} curve
+suggests a ``transfer learning''-type effect.
 Chance is at 5,6\%.
 }
+\end{centering}
 \label{fig_semisup}
@@
 implementation but without bottleneck or autoencoder),
-\textit{middle column:} semi-supervised factored logistic regression
+\textit{middle column:} SSFLogReg
 ($n=20$ latent components, $\lambda=0.5$, $\ell_1=0.1$, $\ell_2=0.1$),
@@
 activity maps from each task.
-Semi-supervised factored logistic regression
+SSFLogReg
 a) puts higher absolute weights on relevant structure,
@@
 and
-c) yields BOLD-typical local contiguity (without enforcing a spatial prior).
+c) yields BOLD-typical local contiguity
+(without enforcing an explicit spatial prior).
 All values are z-scored and thresholded at the $75^{th}$ percentile.
@@
 SSFLogReg therefore
-captures genuine properties of neural activity patterns,
+captures genuine properties of task activity patterns,
 rather than participant- or study-specific artefacts.
@@
 (i.e., non-random initialization) by 
-either corresponding SPCA or ICA loadings did not exhibit improved accuracies,
-neither did $g)$ autoencoder pretraining.
+either corresponding PCA, SPCA or ICA loadings did not exhibit improved accuracies,
+neither did $h)$ autoencoder pretraining.
 %
@@
 $k)$ only $\ell_2$ penalty terms
-was disadvantageous in all tested cases,
-which favored ElasticNet regularization chosen in the above analyses.
+was disadvantageous in all tested cases.
+Thiss favored ElasticNet regularization chosen in the above analyses.
 
@@
 
-We were further interested in the utility of the
-optimized low-rank projection
-in one task dataset for dimensionality reduction in another task dataset.
+We further evaluated whether the low-dimensional space
+learned in HCP task/rest data can be re-used as a feature
+extraction step for learning classification models
+in an independent task dataset.
+These experiments therefore probe
+the generality of the learned representation
+by assessing transfer learning effects.
 To this end, the HCP-derived network decompositions were used as preliminary

### 2015-06-05 nips3mm_scarcity.py

@@ -582,7 +582,7 @@ n_comps = [20]
 import re
 pkgs = glob.glob(RES_NAME + '/*dbg_epochs*.npy')
 dbg_epochs_ = np.load(pkgs[0])
-pkgs = glob.glob(RES_NAME + '/*dbg_acc_train*.npy')
+pkgs = glob.glob(RES_NAME + '/*dbg_acc_val*.npy')
 for n_comp in n_comps:  # 
     plt.figure()
     for p in pkgs:
@@ -593,7 +593,7 @@ for n_comp in n_comps:  #
         rest_samples = int(re.search('_r(.{1,4})_', p).group(1))
         task_samples = int(re.search('_t(.{2,4})dbg', p).group(1))
         
-        if (task_samples != 100) or (rest_samples != 1000):
+        if (task_samples != 100) or (rest_samples != 100):
             continue
         
         if n_comp != n_hidden:
@@ -1167,8 +1167,10 @@ plt.savefig(op.join(WRITE_DIR,
 # n_comp = 20
 # path_main = 'nips3mm_scarcity'
 
-# bars
-plt.close('all')
+# bars (varying nrest)
+my_fs = 13
+k, v = ('out-of-sample performance', '/*_n_comp=%i*_lambda=0.50*r%i_t%i*dbg_acc_val_*.npy')
+# plt.close('all')
 cols = ('#EFC94C', '#E27A3F', '#DF4949',
         '#2998FF', '#0068C4', '#004B8D')
 width = 0.25
@@ -1194,18 +1196,93 @@ for (ntask, nrest), col in zip([[100, 50], [100, 100], [100, 150],
         cur_values[-1], width=width,
         color=col)
 
-plt.title('Semi-supervision in data-scarce and data-rich scenarios')
+plt.title('Varying amount of rest structure', fontsize=my_fs + 2)
 # plt.legend(loc='lower right', fontsize=10,
 #            ncol=1, shadow=True, title="task-maps : rest-maps", fancybox=True)
-plt.yticks(np.linspace(0., 1., 11))
-plt.ylabel(k)
-plt.xlabel('task-maps : rest-maps')
-plt.xticks(np.array(x_pos) + (width / 2), x_labels, rotation=35)
+plt.yticks(np.linspace(0., 1., 11), fontsize=my_fs)
+plt.ylabel(k, fontsize=my_fs)
[…]

### 2015-09-28 paper_nips/paper.tex
start of camera-ready

@@
 
-%\nipsfinalcopy % Uncomment for camera-ready version
-\nipsfinalfalse
+% \nipsfinalcopy % Uncomment for camera-ready version
+\nipsfinaltrue
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-However, testing hypotheses on the neural correlates of
-larger mental task sets
+However, testing hypotheses on the neural correlates underlying
+larger sets of mental tasks
 necessitates adequate representations for the observations.
@@
 with respect to the (unsupervised) reconstruction error and the
-(supervised) task classification.
+(supervised) task detection.
 ${\mathcal{L_{AE}}}$ is devided by ${N_{X_{rest}}}$ to equilibrate both
@@
 PCA reduced the dimensionality of the task data by
-finding orthogonal network components using a change of basis
-(whitening).
+finding orthogonal network components (whitening of the data).
 %
@@
 per-iteration tolerance of 0.0001,
-initialized by random mixing matrix, whitening).
+initialized by random mixing matrix, whitening of the data).
 %
@@
 logistic regression
-(one hidden layer, $\ell_1=0.1$, $\ell_2=0.1$, 500 iterations).
+(no hidden layer, $\ell_1=0.1$, $\ell_2=0.1$, 500 iterations).
 %
@@
 compared against parallel decomposition and classification by SSFLogReg
-(two hidden layers, $\lambda=1$, $\ell_1=0.1$, $\ell_2=0.1$,
+(one hidden layers, $\lambda=1$, $\ell_1=0.1$, $\ell_2=0.1$,
 500 iterations).
@@
 disregarding the autoencoder.
-To this end, ordinary logstic regression was juxtaposed
+To this end, ordinary logistic regression was juxtaposed
 with SSFLogReg with $\lambda=1$.
@@
 from same amounts of task and rest maps.
-In frequently encountered data-scarce scenarios,
-both training and validation
-scores improved as we depart from the most supervised model ($lambda=1$).
+In data-scarce scenarios (frequently encountered by neuroimaging practitioners),
+the out-of-sample scores improve
+as we depart from the most supervised model ($lambda=1$).
 This suggests that the autoencoder of SSFLogReg does not
@@
 like transfer learning.
-In data-rich scenarios, was also present, although less pronounced.
-
-but not in data-rich settings ($\approx{1000}$ labeled samples),
-the repertoire of rest structure modulated model performance
-(Figure \ref{fig_semisup}).
+In data-rich scenarios, we observed the same trend to be apparent.
 
@@
 from a pool of 100 or 1000 different task and rest maps, respectively.
-Especially the \textit{cold} curve
+Especially the \textit{hot} curve
 suggests a ``transfer learning''-type effect.
@@
 was disadvantageous in all tested cases.
-Thiss favored ElasticNet regularization chosen in the above analyses.
+This favored ElasticNet regularization chosen in the above analyses.
 

### 2015-09-28 nips3mm_scarcity.py

@@ -1,4 +1,3 @@
-
 """
 HCP: semi-supervised network decomposition by low-rank logistic regression
 """
@@ -35,6 +34,7 @@ print('Running THEANO on %s' % theano.config.device)
 from nilearn.image import concat_imgs
 import joblib
 import time
+from nilearn.image import smooth_img
 
 RES_NAME = 'nips3mm_scarcity'
 WRITE_DIR = op.join(os.getcwd(), RES_NAME)
@@ -42,7 +42,7 @@ if not op.exists(WRITE_DIR):
     os.mkdir(WRITE_DIR)
 
 MAX_SAMPLES = 2000
-MAX_REST_SAMPLES = 1000
+MAX_REST_SAMPLES = 100
 MAX_TASK_SAMPLES = 100
 
 ##############################################################################
@@ -87,6 +87,45 @@ AT_niis, AT_labels, AT_subs = joblib.load('preload_AT_3mm')
 AT_X = nifti_masker.transform(AT_niis)
 AT_X = StandardScaler().fit_transform(AT_X)
 print('done :)')
+
+def make_new_noise(masker):
+    # CONST
+    n_random_niis = 1000
+    n_random_foci = 200, 300  # random amount of foci per study
+    fwhm_range = (4, 20)
+
+    print "Inventing NOISE images from Gaussians..."
+    # generate some noise niftis
+    inds = np.where(masker.mask_img_.get_data())  # grid of locations
+    x_inds, y_inds, z_inds = inds  # unpack tuple
+    list_niis = []
+    for inii in xrange(n_random_niis):
+        # if inii % 24 == 0:
+        #     print "%i/%i" % (inii + 1, n_random_niis)
+        nfoci = np.random.randint(*n_random_foci)
+        cur_img = np.zeros(masker.mask_img_.shape)
+        for ifocus in xrange(nfoci):
+            # find random point within mask
+            i = np.random.randint(0, len(x_inds) - 1)
+            x, y, z = x_inds[i], y_inds[i], z_inds[i]
+            # put a dot there
+            if np.random.randint(0, 2):
+                cur_img[x, y, z] = 150
+            else:
+                cur_img[x, y, z] = -150
+
+        # smooth current image of random foci
[…]

### 2015-10-04 paper_nips/paper.tex
CAMERA READY

@@
   Bertrand Thirion,
-  Ga\"el Varoquaux \\\textbf{\textit{email:} }firstname.lastname@inria.fr}
+  Ga\"el Varoquaux\\
+  INRIA, Parietal team, Saclay, France\\
+  CEA, Neurospin, Gif-sur-Yvette, France\\
+  firstname.lastname@inria.fr}
 
@@
 due to the high task-rest similarities of neural activity patterns,
-as observed using ICA \cite{smith2009}.
+as observed using ICA \cite{smith2009}
+and linear correlation \cite{cole14}.
 % XXX: you don't discuss the absence of information sharing across tasks / cognitions, i.e. the multi-task aspect of your work.
@@
 variants of most unsupervised learning algorithms,
-including PCA, SPCA, and ICA \cite{hinton06}.
+including PCA, SPCA, and ICA \cite{hertz1991introduction, hinton06}.
 \begin{wrapfigure}{r}{0.40\textwidth}
@@
 and
-2) exploits unlabeled data,
-as they span a space of meaningful configurations.
+2) exploits unlabeled data
+(since they span a space of meaningful configurations).
 
@@
 Neuroimaging task data with labels of ongoing cognitive processes
-were drawn from 500 unrelated,
+were drawn from 500
 healthy HCP participants (cf. Appendix for details on datasets).
@@
   \mathbb E\left[{\mathcal{L_{AE}}}(\mathbf{x})\right] = %
-  \mathbb E\left[\| \mathbf{x} - \mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
+  \mathbb E\left[\| \mathbf{x} - (\mathbf{W_1}(\mathbf{W_0}\mathbf{x} + \mathbf{b_0}) + \mathbf{b_1}) \|^2\right].
 \end{equation}
@@
     \mathbb E\left[{\mathcal{L_{LR}}}(\mathbf{x, y})\right] \approx %
-\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1}).
+\frac{1}{N_{X_{task}}} \sum_{k=0}^{N_{X_{task}}} \log(P(Y=y^{(k)}|\mathbf{x^{(k)}; V_0, V_1, c_0, c_1})).
 \label{eq:lr_loss}
@@
 \paragraph{Layer combination.}
-Importantly, the optimization problem of the linear autoencoder
+The optimization problem of the linear autoencoder
 and the factored logistic regression
@@
 {\mathcal{L}}(\theta, \lambda) = \lambda {\mathcal{L_{LR}}}
-+ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega(\theta).
++ (1-\lambda)\frac{1}{N_{X_{rest}}} {\mathcal{L_{AE}}} + \Omega.
   \label{eq:loss_equ}
@@
 loss terms to the same order of magnitude.
-\(\Omega(\theta)\) represents an ElasticNet-type regularization
+\(\Omega\) represents an ElasticNet-type regularization
 that combines $\ell_1$ and $\ell_2$ penalty terms.
@@
 accessible online for reproducibility and reuse
-(\url{http://github.com/anonymous/anonymous}).
-% (url{http://github.com/banilo/nips2015}).
+% (\url{http://github.com/anonymous/anonymous}).
+(\url{http://github.com/banilo/nips2015}).
 
@@
  \textit{n}               & PCA + LogReg   & SPCA + LogReg  & ICA + LogReg  & AE + LogReg    & SSFLogReg            \\ \hline
- \multicolumn{1}{l|}{5}   & 45,1 \%        & 32,2 \%        & 37,5 \%       & 44,2 \%        & \textbf{95,7\%}      \\
- \multicolumn{1}{l|}{20}  & 78,1 \%        & 78,2 \%        & 81,0 \%       & 63,2 \%        & \textbf{97,3\%}      \\
- \multicolumn{1}{l|}{50}  & 81,7 \%        & 84,0 \%        & 84,2 \%       & 77,0 \%        & \textbf{97,6\%}      \\
- \multicolumn{1}{l|}{100} & 81,3 \%        & 82,2 \%        & 87,3 \%       & 76,6 \%        & \textbf{97,4\%}      \\ \hline
+ \multicolumn{1}{l|}{5}   & 45.1 \%        & 32.2 \%        & 37.5 \%       & 44.2 \%        & \textbf{95.7\%}      \\
+ \multicolumn{1}{l|}{20}  & 78.1 \%        & 78.2 \%        & 81.0 \%       & 63.2 \%        & \textbf{97.3\%}      \\
+ \multicolumn{1}{l|}{50}  & 81.7 \%        & 84.0 \%        & 84.2 \%       & 77.0 \%        & \textbf{97.6\%}      \\
+ \multicolumn{1}{l|}{100} & 81.3 \%        & 82.2 \%        & 87.3 \%       & 76.6 \%        & \textbf{97.4\%}      \\ \hline
  \end{tabular}
@@
 was assessed on 20\% of unseen HCP data.
-Additionally, the ``out-of-study performance'' of
+Additionally, the ``out-of-study'' performance of
 the learned decomposition ($\mathbf{W_0}$)
@@
   \textit{}                 & \multicolumn{1}{l}{$\lambda=0$}    & \multicolumn{1}{l}{$\lambda=0.25$} & \multicolumn{1}{l}{$\lambda=0.5$} & \multicolumn{1}{l}{$\lambda=0.75$} & \multicolumn{1}{l|}{$\lambda=1$} & \multicolumn{1}{l}{$\lambda=0$}  & \multicolumn{1}{l}{$\lambda=0.25$}   & \multicolumn{1}{l}{$\lambda=0.5$}   & \multicolumn{1}{l}{$\lambda=0.75$}   & \multicolumn{1}{l|}{$\lambda=1$}   & $\lambda=0$      & $\lambda=0.25$   & $\lambda=0.5$    & $\lambda=0.75$   & \multicolumn{1}{l}{$\lambda=1$}    \\ \hline
-  Out-of-sample\\accuracy   & \textit{6,0\%}                     & 88,9\%                             & 95,1\%                            & \textbf{96,5\%}                    & 95,7\%                           & \textit{5,5\%}                   & 97,4\%                               & \textbf{97,8\%}                     & 97,3\%                               & 97,3\%                             & \textit{6,1\%}   & 97,2\%           & 97,0\%           & \textbf{97,8\%}  & 97,4\%                             \\
-  Precision (mean)          & \textit{5,9\%}                     & 87,0\%                             & 94,9\%                            & \textbf{96,3\%}                    & 95,4\%                           & \textit{5,1\%}                   & \textbf{97,4\%}                      & 97,1\%                              & 97,0\%                               & 97,0\%                             & \textit{5,9\%}   & 96,9\%           & 96,5\%           & \textbf{97,5\%}  & 96,9\%                             \\
-  Recall (mean)             & \textit{5,6\%}                     & 88,3\%                             & 95,2\%                            & \textbf{96,6\%}                    & 95,7\%                           & \textit{4,6\%}                   & \textbf{97,5\%}                      & 97,5\%                              & 97,4\%                               & 97,4\%                             & \textit{7,2\%}   & 97,2\%           & 97,2\%           & \textbf{97,9\%}  & 97,4\%                             \\
-  F1 score (mean)           & \textit{4,1\%}                     & 86,6\%                             & 94,9\%                            & \textbf{96,4\%}                    & 95,4\%                           & \textit{3,8\%}                   & \textbf{97,4\%}                      & 97,2\%                              & 97,1\%                               & 97,1\%                             & \textit{5,3\%}   & 97,0\%           & 96,7\%           & \textbf{97,7\%}  & 97,2\%                             \\ \hline
-  Out-of-study\\accuracy    & \textit{39,4\%}                    & 60,8\%                             & 54,3\%                            & 60,7\%                             & \textbf{62,9\%}                  & \textit{77,0\%}                  & 79,7\%                               & \textbf{81,9\%}                     & 79,7\%                               & 79,4\%                             & 79,2\%           & \textbf{82,2\%}  & 81,7\%           & 81,3\%           & \textit{75,8\%}                    \\ \hline
+  Out-of-sample\\accuracy   & \textit{6.0\%}                     & 88.9\%                             & 95.1\%                            & \textbf{96.5\%}                    & 95.7\%                           & \textit{5.5\%}                   & 97.4\%                               & \textbf{97.8\%}                     & 97.3\%                               & 97.3\%                             & \textit{6.1\%}   & 97.2\%           & 97.0\%           & \textbf{97.8\%}  & 97.4\%                             \\
+  Precision (mean)          & \textit{5.9\%}                     & 87.0\%                             & 94.9\%                            & \textbf{96.3\%}                    & 95.4\%                           & \textit{5.1\%}                   & \textbf{97.4\%}                      & 97.1\%                              & 97.0\%                               & 97.0\%                             & \textit{5.9\%}   & 96.9\%           & 96.5\%           & \textbf{97.5\%}  & 96.9\%                             \\
+  Recall (mean)             & \textit{5.6\%}                     & 88.3\%                             & 95.2\%                            & \textbf{96.6\%}                    & 95.7\%                           & \textit{4.6\%}                   & \textbf{97.5\%}                      & 97.5\%                              & 97.4\%                               & 97.4\%                             & \textit{7.2\%}   & 97.2\%           & 97.2\%           & \textbf{97.9\%}  & 97.4\%                             \\
+  F1 score (mean)           & \textit{4.1\%}                     & 86.6\%                             & 94.9\%                            & \textbf{96.4\%}                    & 95.4\%                           & \textit{3.8\%}                   & \textbf{97.4\%}                      & 97.2\%                              & 97.1\%                               & 97.1\%                             & \textit{5.3\%}   & 97.0\%           & 96.7\%           & \textbf{97.7\%}  & 97.2\%                             \\ \hline
+  Reconstr.
+error (norm.)   & 0.76                               & 0.85                               & 0.87                              & 1.01                               & 1.79                             & 0.64                             & 0.67                                 & 0.69                                & 0.77                                 & 1.22                               & 0.60             & 0.65             & 0.68             & 0.73             & 1.08                               \\ \hline
+  Out-of-study\\accuracy    & \textit{39.4\%}                    & 60.8\%                             & 54.3\%                            & 60.7\%                             & \textbf{62.9\%}                  & \textit{77.0\%}                  & 79.7\%                               & \textbf{81.9\%}                     & 79.7\%                               & 79.4\%                             & 79.2\%           & \textbf{82.2\%}  & 81.7\%           & 81.3\%           & \textit{75.8\%}                    \\ \hline
   \end{tabular}
@@
  \caption{\textbf{Performance of SSFLogReg across model parameter choices.}
- Chance is at 5,6\%.
+ Chance is at 5.6\%.
  }
@@
 imposing structure from the unlabeled rest data.
+Confirmed by the normalized reconstruction error
+($\mathbf{E} = \|x-\hat{x}\| / \|x\|$),
+little weight on the supervised term is sufficient for good model performance
+while keeping $\mathbf{E}$ low and task-map decomposition rest-like. 
 %
@@
 noninferior out-of-study performance
-of 77,0\% and 79,2\%,
+of 77.0\% and 79.2\%,
 respectively (Table \ref{table_two}).
@@
 To this end, ordinary logistic regression was juxtaposed
-with SSFLogReg with $\lambda=1$.
+with SSFLogReg at $\lambda=1$.
 For this experiment, we increased the difficulty of the classification problem
@@
 are among the least predicted with ordinary but well predicted with
-SSFLogReg
-(tasks 3 and 4 in Figure \ref{fig_dimred}).
+low-rank logistic regression
+(tasks 1 and 2 in Figure \ref{fig_dimred}).
 These experimental conditions have been reported to exhibit
@@
 \end{centering}
-\vspace{-0.5cm}
+\vspace{-0.6cm}
 \caption{\textbf{Effect of bottleneck in a 38-task classificaton problem}
-Depicts the f1 score for each of 38 class predictions (chance is at $2,6\%$)
-sorted from lowest to highest.
+Depicts the f1 prediction scores for each of 38 psychological tasks.
 Multinomial logistic regression operating in voxel space (\textit{blue bars})
@@
 ($\lambda=1$).
-Ordinary logistic regression yielded 77,7\% accuracy out of sample, while
-SSFLogReg scored at 94,4\% ($n=20$) and 94,2\% ($n=100$).
+Ordinary logistic regression yielded 77.7\% accuracy out of sample, while
+SSFLogReg scored at 94.4\% ($n=20$) and 94.2\% ($n=100$).
 Hence, compressing the voxel data into a component space for classification
 achieves higher task separability.
+Chance is at $2,6\%$.
 }
@@
 % rest effect
-We then contemplated the impact of rest structure by modulating
+We then contemplated the impact of rest structure (Figure \ref{fig_semisup})
+by modulating
 its influence ($\lambda=0.25, 0.5, 0.75$) in data-scarce
@@
 the out-of-sample scores improve
-as we depart from the most supervised model ($lambda=1$).
-This suggests that the autoencoder of SSFLogReg does not
-behave like a classical regularization but more
-like transfer learning.
+as we depart from the most supervised model ($\lambda=1$).
 In data-rich scenarios, we observed the same trend to be apparent.
@@
 \includegraphics[width=0.55\textwidth]{figures/semisup.png}
-\vspace{-0.5cm}
+\vspace{-0.55cm}
 \caption{\textbf{Effect of rest structure}
@@
 data-scarce (100 task and 100 rest maps, \textit{hot color}) and
-data-rich (1000 task and 1000 task maps, \textit{cold color}) scenarios.
+data-rich (1000 task and 1000 rest maps, \textit{cold color}) scenarios.
 Gradient descent was performed on 2000 task and 2000 rest maps.
@@
 from a pool of 100 or 1000 different task and rest maps, respectively.
-Especially the \textit{hot} curve
-suggests a ``transfer learning''-type effect.
-Chance is at 5,6\%.
+Chance is at 5.6\%.
 }
@@
 \begin{centering}
-\includegraphics[width=0.98\textwidth]{figures/figure_weights_perc75.png}
+\includegraphics[width=0.98\textwidth]{figures/figure_weights_perc75.jpg}
 \end{centering}
-\vspace{-0.1cm}
+\vspace{-0.25cm}
 \caption{\textbf{Classification weight maps}
@@
 either corresponding PCA, SPCA or ICA loadings did not exhibit improved accuracies,
-neither did $h)$ autoencoder pretraining.
+neither did $g)$ autoencoder pretraining.
 %
@@
 Finally, imposing either $i)$ only $\ell_1$ or 
-$k)$ only $\ell_2$ penalty terms
+$j)$ only $\ell_2$ penalty terms
 was disadvantageous in all tested cases.
@@
 %
-The higher generalization accuracy and feature recovery, comparing to
+The higher generalization accuracy and support recovery, comparing to
 ordinary logistic regression, hold potential
@@
 predict disease trajectories and drug responses
-in psychiatric/neurological populations
-\cite{frackowiak2015future, gustav2011cost}.
+in psychiatric and neurological populations
+\cite{frackowiak2015future}.
 
 \paragraph{Acknowledgment}
-{\small
+{\small The research leading to these results has received funding from the
+European Union Seventh Framework Programme (FP7/2007-2013)
+under grant agreement no. 604102 (Human Brain Project).
 Data were provided by the Human Connectome Project.
-% The study was supported by the German National Academic Foundation (D.B.).
+Further support was received from
+the German National Academic Foundation (D.B.).
 }
@@
 Neuroimaging task data with labels of ongoing cognitive processes
-were drawn from 500 unrelated,
+were drawn from 500
 healthy HCP participants.
@@
 We here drew on ``minimally preprocessed'' rest data
-from 200 randomly selected healthy, unrelated participants.
+from 200 randomly selected healthy participants.
 PCA was applied to each set of 1,200 rest maps for
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=0.65\textwidth]{figures/semisup_conversion.png}
+  \vspace{-0.1cm}
+\caption{\textbf{Weight maps of a same hidden factor ranging from
+unsupervised to supervised regime}
+One of the $n$ factors from the hidden layer ($\mathbf{W_0}$)
+was plotted for the same data (full HCP dataset)
+and
+the same model choices
+($n=20, \ell_1=0.1$, $\ell_2=0.1$) along
+a $\lambda$-grid between
+purely unsupervised ($\lambda=0.0$, \textit{top row}) and
+purely supervised ($\lambda=1.0$, \textit{bottom row}) settings.
+%
+As qualitative evidence,
+a slow transition from rest- to task-typical brain networks was observed
+in brain space.
+Although difficult to quantify,
+rest network elements appear to get 'reassembled' to
+latent factors of the LR.
+This increased confidence that the improved model performance
+of rest-informed fLR is not only an arbitrary
+effect of spatially smooth noise.
+%
+All values are z-scored.
+}
+\label{fig_weights_conversion}
+\end{figure}
+
 \end{document}
